### 2019-12-28 Paper/paper.tex
add initial structure for tex paper

%----------------------------------------------------------------------------------------
%	HEADER
%----------------------------------------------------------------------------------------

\documentclass[12pt, a4paper, oneside]{article}

\usepackage[english]{babel} % Language hyphenation and typographical rules

\usepackage[top=3cm, left=3cm, right=2cm, bottom=2.5cm]{geometry} % Document margins
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text

\usepackage{enumitem} % Customized lists
\setlist[itemize]{noitemsep} % Make itemize lists more compact

% TODO do we need a abstract?
\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
%\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
\renewcommand\thesubsection{\roman{subsection}} % roman numerals for subsections
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Water Management $\bullet$ Seminar Computational Economics $\bullet$ Winter Semester 2019/2020} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

\usepackage{titling} % Customizing the title section

\usepackage{hyperref} % For hyperlinks in the PDF

% TODO delete the lorem ipsum
\usepackage{blindtext} % Package to generate dummy text throughout this template

\title{Water Management}
%----------------------------------------------------------------------------------------

\begin{document}

	%----------------------------------------------------------------------------------------
	%	COVER SHEET
	%----------------------------------------------------------------------------------------

	\begin{titlepage}
		\begin{center}
			\vspace*{1cm}

			\textbf{Seminar \textit{Computational Economics}}

			\vspace{0.5cm}
			{\makeatletter{\@title}\makeatother}

			\vspace{1.5cm}

			\textbf{Felix Fink} \\
			stud. 2nd semester M.Sc. Mechanical Engineering \\
			Matr. Nr. 345971 \\
			\href{mailto:felix.fink@rwth-aachen.de}{felix.fink@rwth-aachen.de}

			\vspace{1cm}

			% TODO input personal details
			\textbf{Niklas Sayer} \\
			\href{mailto:niklas.sayer@rwth-aachen.de}{niklas.sayer@rwth-aachen.de}

			\vspace{1cm}

			% TODO input personal details
			\textbf{Martin Wicke} \\
			\href{mailto:martin.wicke@rwth-aachen.de}{martin.wicke:@rwth-aachen.de}

			\vfill

			Supervisor\\
			Thomas Gehrmann, M.Sc.

			\vspace{0.8cm}

			%\includegraphics[width=0.4\textwidth]{university}

			Computational Economics\\
			RWTH Aachen University\\
			Winter Semester 2019/2020

		\end{center}
	\end{titlepage}

	%----------------------------------------------------------------------------------------
	%	TABLE OF CONTENTS
	%----------------------------------------------------------------------------------------

	\tableofcontents
	\clearpage

	%----------------------------------------------------------------------------------------
	%	ARTICLE CONTENTS
	%----------------------------------------------------------------------------------------

	\section{Introduction}

	\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
	\blindtext % Dummy text

	%------------------------------------------------

	\section{Main Part}
	\subsection{bla}
	\blindtext
	\blindtext
	\blindtext
	\blindtext

	%------------------------------------------------

	\section{Conclusion}
	\blindtext

	\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template

		\bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}
		Figueredo, A.~J. and Wolf, P. S.~A. (2009).
		\newblock Assortative pairing and life history strategy - a cross-cultural
		study.
		\newblock {\em Human Nature}, 20:317--330.

	\end{thebibliography}

	\clearpage

	%----------------------------------------------------------------------------------------
	%	STATUTORY DECLARATION IN LIEU OF AN OATH
	%----------------------------------------------------------------------------------------

	\begingroup
	\begin{center}
		\large
		\textbf{Eidesstattliche Versicherung}
	\end{center}

	\vspace{0.3cm}

	\begin{tabular}{@{}p{8cm}p{5.8cm}}
		\underline{\hspace{6cm}} & \underline{\hspace{5.8cm}} \\
		\vspace{0.02cm}Name, Vorname & \vspace{0.02cm}Matrikelnummer \\
	\end{tabular}
	\vspace{0.3cm}

	Ich versichere hiermit an Eides Statt, dass ich die vorliegende Arbeit mit dem Titel

{\makeatletter{\emph{{\@title}}}\makeatother}

	selbständig und ohne unzulässige fremde Hilfe erbracht habe. Ich habe keine anderen als die angegebenen Quellen und Hilfsmittel benutzt. Für den Fall, dass die Arbeit zusätzlich auf einem Datenträger eingereicht wird, erkläre ich, dass die schriftliche und die elektronische Form vollständig übereinstimmen. Die Arbeit hat in gleicher oder ähnlicher Form noch keiner Prüfungsbehörde vorgelegen.

	\begin{tabular}{@{}p{8cm}p{5.8cm}}
		\underline{\smash{Aachen, den \today}} & \underline{\hspace{5.8cm}}\\
		Ort, Datum & Unterschrift \\
	\end{tabular}

	\vspace{0.6cm}
	\begin{small}
		\textbf{Belehrung:}\\
		Wer vorsätzlich gegen eine die Täuschung über Prüfungsleistungen betreffende Regelung einer Hochschulprüfungsordnung verstößt, handelt ordnungswidrig. Die Ordnungswidrigkeit kann mit einer Geldbuße bis zu 50 000 Euro geahndet werden. Zuständige Verwaltungsbehörde für die Verfolgung und Ahndung von Ordnungswidrigkeiten ist die Kanzlerin oder der Kanzler der RWTH Aachen. Im Falle eines mehrfachen oder sonstigen schwerwiegenden Täuschungsversuches kann der Prüfling zudem exmatrikuliert werden (§ 63 Abs. 5 HG NRW).

		\textbf{§ 156 StGB: Falsche Versicherung an Eides Statt}\\
		Wer vor einer zur Abnahme einer Versicherung an Eides Statt zuständigen Behörde eine solche Versicherung falsch abgibt oder unter Berufung auf eine solche Versicherung falsch aussagt, wird mit Freiheitsstrafe bis zu drei Jahren oder mit Geldstrafe bestraft.

		\textbf{§ 161 StGB: Fahrlässiger Falscheid; fahrlässige falsche Versicherung an Eides Statt}\\
		(1) Wenn eine der in den §§ 154 bis 156 bezeichneten Handlungen aus Fahrlässigkeit begangen worden ist, so tritt Freiheitsstrafe bis zu einem Jahr oder Geldstrafe ein.\\
		(2) Straflosigkeit tritt ein, wenn der Täter die falsche Angabe rechtzeitig berichtigt. Die Vorschriften des § 158 Abs. 2 und 3 gelten entsprechend
	\end{small}

	Die vorstehende Belehrung habe ich zur Kenntnis genommen:

	\begin{tabular}{@{}p{8cm}p{5.8cm}}
		\underline{\smash{Aachen, den \today}} & \underline{\hspace{5.8cm}}\\
		Ort, Datum & Unterschrift \\
	\end{tabular}

	\endgroup
	\clearpage

\end{document}

### 2020-01-02 Code/code.m
Update code.m

@@ -1,0 +1,50 @@
+%% parameters
+% preference parameters
+a1 = 1;
+a2 = 2;
+b1 = -2;
+b2 = -3;
+
+% maximum capacity of the reservoir
+totalReservoirCapacity = 7;
+
+% initial fill level of the reservoir
+initialReservoirCapacity = totalReservoirCapacity/2;
+
+% discount factor
+beta = 0.9;
+
+% rainfall distribution
+mu = 0; 
+sigma = 1;
+
+% simulation periods
+periods = 365;
+
+
+%% utility functions
+% farmers utility
+uFarmer = @(x) a1/(1+b1)*x.^(1+b1);
+
+% recreational users utility
+uRecr = @(s,x) a2/(1+b2)*(s-x).^(1+b2);
+
+% total utility
+uTotal = @(s,x) uFarmer(x) + uRecr(s,x);
+
+
+%% other functions
+% rainfall
+r = lognrnd(mu,sigma, 1, periods);
+
+%% simulation
+
+reservoirFillLevels = zeros(periods,1);
+reservoirFillLevels(1) = initialReservoirCapacity;
+
+for p=2:periods
+    reservoirFillLevels(p) = reservoirFillLevels(p-1)+r(p);
+    reservoirFillLevels(p) = min(totalReservoirCapacity, reservoirFillLevels(p));
+end
+reservoirFillLevels
+

### 2020-01-02 Paper/paper.tex
further formatting
added personal details
begin introduction

@@
 \fancyfoot[RO,LE]{\thepage} % Custom footer text
+
+\setlength{\parindent}{0em}
 
@@
 
-			% TODO input personal details
 			\textbf{Martin Wicke} \\
+			stud. M.Sc. Business Administration and Engineering: Electrical Power Engineering\\
+			Matr. Nr. 336301 \\
 			\href{mailto:martin.wicke@rwth-aachen.de}{martin.wicke:@rwth-aachen.de}
@@
 
-	\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
-	\blindtext % Dummy text
+	\subsection{Background}
+In this work, we want to examine the variations in use of a (hypothetical) water reservoir.
+Water management of reservoir lakes poses a non trivial task.
+Different stakeholders have different interests in how to use the water and as lakes usually are public goods, claims of all parties shall be considered.
+As climate change proceeds, rainfall as the natural influx of lakes becomes more volatile and prolonged dry periods make water an even more precious resource than today.
+
+\subsection{Assumptions and Conditions}
+
+The water reservoir to be considered has to stakeholders: farmers, who want to use the water for irrigation purposes and ordinary people, who want to use the lake for recreational activities. \\\\
+The farmers utility function follows the form
+
+\begin{equation}
+%	F(x) = \frac{a_1}{1+b_1} * x^{1+b_1}
+	F(x) = -\frac{1}{x}
+\end{equation}
+
+where $x$ is the amount of water to be used for irrigation.\\\\
+
+Analogously the recreational users utility function corresponds to
+
+\begin{equation}
+%  G(s, x) = \frac{a_2}{1+b_2} * (s-x)^{1+b_2}
+  G(s, x) = -\frac{1}{(s-x)^2}
+\end{equation}
+
+with the additional paramter $s$ which indicates the water level.
+
+The rainfall shall be lognormal distributed: 
+\begin{equation}
+	r \sim lognormal(\mu, \sigma^2)
+\end{equation}
 
@@
 
-	\clearpage
+	\newpage
 
@@
 
+	\begin{center}
 		{\makeatletter{\emph{{\@title}}}\makeatother}
+	\end{center}
 
@@
 Die Arbeit hat in gleicher oder ähnlicher Form noch keiner Prüfungsbehörde vorgelegen.
+
+	\vspace{0.6cm}
 
@@
 	\vspace{0.6cm}
+
 	\begin{small}
@@
 
+	\vspace{0.6cm}
+
 	Die vorstehende Belehrung habe ich zur Kenntnis genommen:
+
+	\vspace{0.6cm}
 

### 2020-01-02 Code/code.m
Code: All given tasks

@@ -29,7 +29,7 @@ maxIter = 10000;
 tol = 1e-6;
 
 % Number of periods for forward simulation
-T = 100;
+T = 1000;
 
 %% Rainfall simulation
 % rainfall: lognormal distribution
@@ -96,9 +96,9 @@ hold off
 
 %% forward iteration
 
-% Index for current water level; set to dimWL in period 1
+% Index for current water level; set to 1 (=empty) in period 1
 waterInd = zeros(1, T+1);
-waterInd(1) = dimWL;
+waterInd(1) = 1;
 
 % Index for water that is used for irrigation
 irrigationInd = zeros(1, T);
@@ -111,14 +111,27 @@ for i=1:T
     
 end
 
+steadyStateLvl = mean(waterLevel(waterInd));
+
+% Plot water level, amount of water used for irrigation and steady state
+% value
 figure(2)
 hold on
 plot(waterLevel(waterInd));    
 plot(waterLevel(irrigationInd));
+plot([1 T],[steadyStateLvl steadyStateLvl],'--g');
 xlim([1 T]);
-legend('water left','water used for irrigation');
+legend('water left','water used for irrigation','steady state level');
 title('Optimal Extraction Policy');
 xlabel('period');
 ylabel('water level in reservoir');
 hold off
-    
+
+% plot histogram
+figure(3)
+hold on
+histogram(waterLevel(waterInd),30, 'Normalization','probability');
+title('Steady-State distribution of water level');
+ylabel('probability');
+xlabel('water level in reservoir');
+hold off

### 2020-01-03 Paper/paper.tex
Added personal details
minor stuff in paper

@@
 			\textbf{Felix Fink} \\
-			stud. 2nd semester M.Sc. Mechanical Engineering \\
+			stud. 2nd semester M.Sc. Business Administration and Engineering: Mechanical Engineering \\
 			Matr. Nr. 345971 \\
@@
 
-			% TODO input personal details
 			\textbf{Niklas Sayer} \\
+			stud. M.Sc. Business Administration and Engineering: Materials- and Process Engineering\\
+			Matr. Nr. 332404
 			\href{mailto:niklas.sayer@rwth-aachen.de}{niklas.sayer@rwth-aachen.de}
@@
 			Matr. Nr. 336301 \\
-			\href{mailto:martin.wicke@rwth-aachen.de}{martin.wicke:@rwth-aachen.de}
+			\href{mailto:martin.wicke@rwth-aachen.de}{martin.wicke@rwth-aachen.de}
 
@@
 Different stakeholders have different interests in how to use the water and as lakes usually are public goods, claims of all parties shall be considered.
-As climate change proceeds, rainfall as the natural influx of lakes becomes more volatile and prolonged dry periods make water an even more precious resource than today.
+As climate change proceeds, rainfall as the natural influx of lakes becomes more volatile and prolonged dry periods will make water an even more precious resource than it is today.
 
@@
 
-The water reservoir to be considered has to stakeholders: farmers, who want to use the water for irrigation purposes and ordinary people, who want to use the lake for recreational activities. \\\\
+The water reservoir to be considered has to stakeholders: farmers, who want to use the water for irrigation purposes and citzens, who want to use the lake for recreational activities. \\\\
 The farmers utility function follows the form

### 2020-01-03 Code/code.m
Delete code.asv

@@ -72,7 +72,7 @@ for j = 1:maxIter
         for iIrrigation = 1:iWL
             aux(iWL, iIrrigation) = utilFar(waterLevel(iIrrigation)) +...
                 utilRec(waterLevel(iWL), waterLevel(iIrrigation)) +...
-                beta.*V_old(min(max(iWL-iIrrigation+1+dimE_Rain,1),dimWL));
+                beta*V_old(min(max(iWL-iIrrigation+1+dimE_Rain,1),dimWL));
         end
     end
     [V, optIrrigation_ind]= max(aux,[],2);  

### 2020-01-03 Paper/paper.tex
Paper: Some changes to introduction

@@
 As climate change proceeds, rainfall as the natural influx of lakes becomes more volatile and prolonged dry periods will make water an even more precious resource than it is today.
+Therefore, the optimal allocation of the available water will be of increasing importance in the future.
 
@@
 
-The water reservoir to be considered has to stakeholders: farmers, who want to use the water for irrigation purposes and citzens, who want to use the lake for recreational activities. \\\\
+The water reservoir to be considered has two stakeholders: farmers, who want to use the water for irrigation purposes and citizens, who want to use the lake for recreational activities.
+At the beginning of every year we have to decide how much of the water from the reservoir is used for the irrigation of fields, which benefits the farmers' harvest.\\\\
 The farmers utility function follows the form
@@
 \begin{equation}
-%	F(x) = \frac{a_1}{1+b_1} * x^{1+b_1}
-	F(x) = -\frac{1}{x}
+	F(x) = \frac{a_1}{1+b_1} * x^{1+b_1}
+%	F(x) = -\frac{1}{x}
 \end{equation}
 
-where $x$ is the amount of water to be used for irrigation.\\\\
+where $x$ represents the amount of water to be used for irrigation in a year. The preference parameters are $a_1 = 1$ and $b_1 = -2$.\\\\
 
-Analogously the recreational users utility function corresponds to
+During the summer months, citizens benefit from the remaining water due to recreational use (e.g. swimming). This leads to the recreational users' utility function which can be described as
 
 \begin{equation}
-%  G(s, x) = \frac{a_2}{1+b_2} * (s-x)^{1+b_2}
-  G(s, x) = -\frac{1}{(s-x)^2}
+  G(s, x) = \frac{a_2}{1+b_2} * (s-x)^{1+b_2}
+%  G(s, x) = -\frac{1}{(s-x)^2}
 \end{equation}
 
-with the additional paramter $s$ which indicates the water level.
+with the additional paramter $s$ which indicates the water level at the beginning of the year. For recreational users, the preference parameters shell be $a_2 = 2$ and $b_2 = -3$.\\\\
 
-The rainfall shall be lognormal distributed: 
+At the end of the year, water in the reservoir is replenished by rainfall, which we assume to be randomly following a lognormal distribution: %Evtl Zitat: https://www.waterlog.info/pdf/freqtxt.pdf
 \begin{equation}
 	r \sim lognormal(\mu, \sigma^2)
+\end{equation}
+
+with parameters $\mu = 0$ and $\sigma^2 = 1$.\\\\
+
+Other parameters are the maximum capacity of the reservoir
+\begin{equation}
+	M = 7
+\end{equation}
+
+and the discount factor for future utility
+
+\begin{equation}
+	\beta = 0.9
 \end{equation}

### 2020-01-03 Code/code.m

@@ -29,7 +29,7 @@ maxIter = 10000;
 tol = 1e-6;
 
 % Number of periods for forward simulation
-T = 1000;
+T = 100;
 
 %% Rainfall simulation
 % rainfall: lognormal distribution
@@ -122,7 +122,7 @@ plot(waterLevel(irrigationInd));
 plot([1 T],[steadyStateLvl steadyStateLvl],'--g');
 xlim([1 T]);
 legend('water left','water used for irrigation','steady state level');
-title('Optimal Extraction Policy');
+title('Optimal Irrigation Policy');
 xlabel('period');
 ylabel('water level in reservoir');
 hold off

### 2020-01-12 Code/code.m
calculate mean over periods

@@ -109,19 +109,24 @@ waterInd(1) = 1;
 
 % Index for water that is used for irrigation
 irrigationInd = zeros(1, T);
-
+steadyStateLvls = zeros(2, T);
 for i=1:T
     
     irrigationInd(i) = optIrrigation_ind(waterInd(i));
     
     waterInd(i+1) = min(waterInd(i) - irrigationInd(i) + round(r(i)/(M/dimWL)),dimWL);
-    
+    steadyStateLvls(1,i) = mean(waterLevel(waterInd(1:i)));
+    if (i~=1)
+        % the differences to the previous mean
+        steadyStateLvls(2,i) = steadyStateLvls(1,i)-steadyStateLvls(1,i-1);
+    end
 end
 
 steadyStateThreshold = 20; % Change this parameter to determine the period 
                            % after which the system is expected to be in a 
                            % steady state (has to be smaller then T!)
-                       
+
+%% TODO write function that returns the period of steady state depending on a difference threshold
 steadyStateLvl = mean(waterLevel(waterInd(steadyStateThreshold:end)));
 
 % Plot water level, amount of water used for irrigation and steady state
@@ -130,9 +135,11 @@ figure(2)
 hold on
 plot(waterLevel(waterInd));    
 plot(waterLevel(irrigationInd));
+plot(steadyStateLvls(1,:));
+plot(steadyStateLvls(2,:));
 plot([1 T],[steadyStateLvl steadyStateLvl],'--g');
 xlim([1 T]);
-legend('water left','water used for irrigation','steady state level');
+legend('water left','water used for irrigation','Mean until period','Difference in mean to previous','steady state level');
 title('Optimal Irrigation Policy');
 xlabel('period');
 ylabel('water level in reservoir');

### 2020-01-12 Code/code.m
dynamicly calculate threshold period

@@ -110,6 +110,7 @@ waterInd(1) = 1;
 % Index for water that is used for irrigation
 irrigationInd = zeros(1, T);
 steadyStateLvls = zeros(2, T);
+steadyStateThreshold = T;
 for i=1:T
     
     irrigationInd(i) = optIrrigation_ind(waterInd(i));
@@ -122,11 +123,17 @@ for i=1:T
     end
 end
 
-steadyStateThreshold = 20; % Change this parameter to determine the period 
+for i=2:T
+    if (abs(steadyStateLvls(2,i)) < 0.05)
+        steadyStateThreshold = i;
+        break;
+    end
+end
+%steadyStateThreshold = 20; % Change this parameter to determine the period
                            % after which the system is expected to be in a 
                            % steady state (has to be smaller then T!)
 
-%% TODO write function that returns the period of steady state depending on a difference threshold
+%% TODO better threshold calculation by using more that one period to evaluate the difference
 steadyStateLvl = mean(waterLevel(waterInd(steadyStateThreshold:end)));
 
 % Plot water level, amount of water used for irrigation and steady state

### 2020-01-16 Paper/paper.tex
example figures
Introduction changes + Citations

@@
 
+\usepackage{graphicx} % for figures
+
+\usepackage{longtable}
+
 % TODO delete the lorem ipsum
@@
 			\textbf{Niklas Sayer} \\
-			stud. M.Sc. Business Administration and Engineering: Materials- and Process Engineering\\
+			stud. 2nd semester M.Sc. Business Administration and Engineering: Materials- and Process Engineering\\
 			Matr. Nr. 332404
@@
 	\subsection{Background}
-In this work, we want to examine the variations in use of a (hypothetical) water reservoir.
-Water management of reservoir lakes poses a non trivial task.
-Different stakeholders have different interests in how to use the water and as lakes usually are public goods, claims of all parties shall be considered.
-As climate change proceeds, rainfall as the natural influx of lakes becomes more volatile and prolonged dry periods will make water an even more precious resource than it is today.
-Therefore, the optimal allocation of the available water will be of increasing importance in the future.
+With world population expected to hit 9.7 billion in 2050 and 11.2 billion in 2100, more and more people have to be fed, which will increase the demand of water for agriculture \cite{UN:2019}.
+Already today, around one third of the world's population live under physical water scarcity \cite{Vörösmarty:2000, Alcamo:2003, Oki and Kanae:2006}.
+With climate change proceeding, rainfall as the natural influx of lakes will become more volatile in some regions while prolonged dry periods will make water an even more precious resource in other areas \cite{Guhathakurta:2011}.
+This situation will get even worse due to the projected increase in the demand of water per capita in many regions of the world, driven by lifestyle factors, especially diet \cite{Vörösmarty:2000}.
+Despite all those challenges, research suggests that just by using water more efficiently, we will be able to meet the acute water, environment and poverty challenges facing us over the next 50 years \cite {Molden:2013}.
+Therefore, the optimal allocation and sustainable use of the available water will be of increasing importance in the future.
+
+In this work we want to examine the variations in use of water from a (hypothetical) water reservoir.
+One of the most important uses of surface water is the irrigation of fields for food production.
+However, as farming and food production are private enterprises and lakes are public goods, interests of all different parties have to be considered.
 
@@
 	\subsection{bla}
+	\begin{figure}[ht] % TODO remove example
+		\includegraphics[width=1\textwidth]{figures/example.png}
+		\caption{Example Image}
+		\label{fig:example-pic}
+	\end{figure}
 	\blindtext
+	\begin{figure}[ht] % TODO remove example
+		\begin{longtable}{|p{2,5cm}|p{2,5cm}|p{2,5cm}|p{3cm}|p{3cm}|}
+			\hline
+			&\multicolumn{4}{c|}{\textbf{Klasse}} \\
+			\hline
+			& \textbf{S0} & \textbf{S1} & \textbf{S2} & \textbf{S3} \\
+			\hline
+			\textbf{lorem} & ipsum & ipsum & ipsum & ipsum \\
+			\hline
+		\end{longtable}
+		\label{fig:example-figure}
+		\caption{Example Table}
+	\end{figure}
 	\blindtext
@@
 
+		\bibitem[United Nations, 2019]{UN:2019}
+		United Nations, Department of Economic and Social Affairs, Population Division (2019). 
+		\newblock World Population Prospects 2019: Press Release.
+
+		\bibitem[Vörösmarty et al., 2000]{Vörösmarty:2000}
+		Vörösmarty C J, Green P, Salisbury J and Lammers R B (2000).
+		\newblock Global water resources: vulnerability from climate change and population growth.
+		\newblock {\em Science}, 289:284--288.
+
+		\bibitem[Alcamo et al., 2003]{Alcamo:2003}
+		Alcamo J, Döll P, Henrichs T, Kaspar F, Lehner B, Rösch T and Siebert S (2003).
+		\newblock Global estimates of water withdrawals and availability under current and future ‘business-as-usual’ conditions.
+		\newblock {\em Hydrol. Sci. J.}, 48:339--348.
+
+		\bibitem[Oki and Kanae, 2006]{Oki and Kanae:2006}
+		Oki T and Kanae S (2006).
+		\newblock Global hydrological cycles and world water resources.
+		\newblock {\em Science}, 313:1068--1072.
+
+		\bibitem[Guhathakurta et al., 2011]{Guhathakurta:2011}	
+		Guhathakurta, P., Sreejith, O. P. and Menon, P. A. (2011).
+		\newblock Impact of climate change on extreme rainfall events and flood risk in India. 
+		\newblock {\em Journal of earth system science}, 120(3):359.
+
+		\bibitem[D. Molden, 2013]{Molden:2013}
+		Molden, D. (2013). 
+		\newblock Water for food water for life: A comprehensive assessment of water management in agriculture. 
+		\newblock {\em Routledge}
+
+		%Molden, D. (Ed.). (2007). Water for Food Water for Life. London: Routledge, https://doi.org/10.4324/9781849773799
 	\end{thebibliography}

### 2020-01-16 Paper/paper.tex
Introduction

@@
 
-	\subsection{Background}
-With world population expected to hit 9.7 billion in 2050 and 11.2 billion in 2100, more and more people have to be fed, which will increase the demand of water for agriculture \cite{UN:2019}.
+	\subsection{Motivation}
+With world population expected to hit 9.7 billion in 2050 and 11.2 billion in 2100, more and more people have to be fed, which will increase the demand of water for agricultural uses \cite{UN:2019}.
 Already today, around one third of the world's population live under physical water scarcity \cite{Vörösmarty:2000, Alcamo:2003, Oki and Kanae:2006}.
@@
 
+\subsection{Assumptions and Conditions}
+
 In this work we want to examine the variations in use of water from a (hypothetical) water reservoir.
@@
 However, as farming and food production are private enterprises and lakes are public goods, interests of all different parties have to be considered.
-
-\subsection{Assumptions and Conditions}
-
 The water reservoir to be considered has two stakeholders: farmers, who want to use the water for irrigation purposes and citizens, who want to use the lake for recreational activities.
@@
 	F(x) = \frac{a_1}{1+b_1} * x^{1+b_1}
-%	F(x) = -\frac{1}{x}
 \end{equation}
@@
 
-During the summer months, citizens benefit from the remaining water due to recreational use (e.g. swimming). This leads to the recreational users' utility function which can be described as
+During the summer months, citizens benefit from the remaining water due to recreational use (e.g. swimming). This can be described by the recreational users' utility function which is
 
@@
   G(s, x) = \frac{a_2}{1+b_2} * (s-x)^{1+b_2}
-%  G(s, x) = -\frac{1}{(s-x)^2}
 \end{equation}
 
-with the additional paramter $s$ which indicates the water level at the beginning of the year. For recreational users, the preference parameters shell be $a_2 = 2$ and $b_2 = -3$.\\\\
+with the additional paramter $s$ that indicates the water level at the beginning of the year. For recreational users, the preference parameters shell be $a_2 = 2$ and $b_2 = -3$.\\\\
 

### 2020-01-16 Paper/paper.tex
merge changes

--- 2020-01-03 Paper/paper.tex
@@
 
+\usepackage{graphicx} % for figures
+
+\usepackage{longtable}
+
+% packages used for citations
+\usepackage[
+backend=biber,
+style=alphabetic,
+citestyle=authoryear,
+natbib=true
+]{biblatex}
+\addbibresource{collection.bib}
+
 % TODO delete the lorem ipsum
@@
 
-	\subsection{Background}
-In this work, we want to examine the variations in use of a (hypothetical) water reservoir.
-Water management of reservoir lakes poses a non trivial task.
-Different stakeholders have different interests in how to use the water and as lakes usually are public goods, claims of all parties shall be considered.
-As climate change proceeds, rainfall as the natural influx of lakes becomes more volatile and prolonged dry periods will make water an even more precious resource than it is today.
-Therefore, the optimal allocation of the available water will be of increasing importance in the future.
+	\subsection{Motivation}
+With world population expected to hit 9.7 billion in 2050 and 11.2 billion in 2100, more and more people have to be fed, which will increase the demand of water for agriculture uses \citep{un2019}.
+Already today, around one third of the world's population live under physical water scarcity \citep{vorosmarty2000global, alcamo2003global, oki2006global}.
+With climate change proceeding, rainfall as the natural influx of lakes will become more volatile in some regions while prolonged dry periods will make water an even more precious resource in other areas \citep{guhathakurta2011impact}.
+This situation will get even worse due to the projected increase in the demand of water per capita in many regions of the world, driven by lifestyle factors, especially diet \citep{vorosmarty2000global}.
+Despite all those challenges, research suggests that just by using water more efficiently, we will be able to meet the acute water, environment and poverty challenges facing us over the next 50 years \citep{molden2013water}.
+Therefore, the optimal allocation and sustainable use of the available water will be of increasing importance in the future.
 
@@
 
+In this work we want to examine the variations in use of water from a (hypothetical) water reservoir.
+One of the most important uses of surface water is the irrigation of fields for food production.
+However, as farming and food production are private enterprises and lakes are public goods, interests of all different parties have to be considered.
 The water reservoir to be considered has two stakeholders: farmers, who want to use the water for irrigation purposes and citizens, who want to use the lake for recreational activities.
@@
 
-	\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template
-
-		\bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}
-		Figueredo, A.~J. and Wolf, P. S.~A. (2009).
-		\newblock Assortative pairing and life history strategy - a cross-cultural
-		study.
-		\newblock {\em Human Nature}, 20:317--330.
-
-	\end{thebibliography}
+	\printbibliography
 

### 2020-01-16 Paper/paper.tex
biber -> bibtex
introduction changes

@@
 \usepackage[
-backend=biber,
+backend=bibtex,
 style=alphabetic,
@@
 
-At the end of the year, water in the reservoir is replenished by rainfall, which we assume to be randomly following a lognormal distribution: %Evtl Zitat: https://www.waterlog.info/pdf/freqtxt.pdf
+At the end of the year, water in the reservoir is replenished by rainfall, which we can assume to be randomly following a lognormal distribution \citep{oosterbaan1994frequency}:
 \begin{equation}
@@
 \end{equation}
+
+Using dynamic programming...
 

### 2020-01-16 Code/code.m
change calculation of steady state

@@ -111,12 +111,17 @@ waterInd(1) = 1;
 irrigationInd = zeros(1, T);
 steadyStateLvls = zeros(2, T);
 steadyStateThreshold = T;
+periodsForMean = 150;
 for i=1:T
     
     irrigationInd(i) = optIrrigation_ind(waterInd(i));
     
     waterInd(i+1) = min(waterInd(i) - irrigationInd(i) + round(r(i)/(M/dimWL)),dimWL);
-    steadyStateLvls(1,i) = mean(waterLevel(waterInd(1:i)));
+    if (i>periodsForMean)
+        steadyStateLvls(1,i) = mean(waterLevel(waterInd(i-periodsForMean:i)));
+    else
+        steadyStateLvls(1,i) = mean(waterLevel(waterInd(1:i)));
+    end
     if (i~=1)
         % the differences to the previous mean
         steadyStateLvls(2,i) = steadyStateLvls(1,i)-steadyStateLvls(1,i-1);
@@ -124,8 +129,9 @@ for i=1:T
 end
 
 for i=2:T
-    if (abs(steadyStateLvls(2,i)) < 0.05)
+    if (abs(steadyStateLvls(2,i)) < 0.01)
         steadyStateThreshold = i;
+        fprintf('Steady state found in period %s\n',num2str(i));
         break;
     end
 end

### 2020-01-16 Code/code.m
also display the amount of rain

@@ -148,14 +148,15 @@ figure(2)
 hold on
 plot(waterLevel(waterInd));    
 plot(waterLevel(irrigationInd));
+plot(r);
 plot(steadyStateLvls(1,:));
 plot(steadyStateLvls(2,:));
 plot([1 T],[steadyStateLvl steadyStateLvl],'--g');
 xlim([1 T]);
-legend('water left','water used for irrigation','Mean until period','Difference in mean to previous','steady state level');
+legend('water level in reservoir','water used for irrigation','rain','Mean last 50 periods','Difference in mean to previous','steady state level');
 title('Optimal Irrigation Policy');
 xlabel('period');
-ylabel('water level in reservoir');
+ylabel('amount');
 hold off
 
 % plot histogram

### 2020-01-16 Code/code.m
use realdaten with trigger to shut in on or off

@@ -3,6 +3,9 @@ close all
 clear all
 %% Parameters & utility functions
 
+% set realdaten to true or false
+realdaten = true;
+
 % Preferences
 a1 = 1;
 a2 = 2;
@@ -29,7 +32,7 @@ maxIter = 10000;
 tol = 1e-6;
 
 % Number of periods for forward simulation
-T = 100;
+T = 64;
 
 %% Rainfall simulation
 % Rainfall parameters (might be overwritten by real data)
@@ -37,24 +40,37 @@ mu = 0;
 sigma = 1;
 
 % % Realdaten
-% c = 4; % Column: 1,2,3,4 = POONDI,CHOLAVARAM,REDHILLS,CHEMBARAMBAKKAM
-% data = csvread('chennai_reservoir_rainfall_formatted.csv',1,c,[1 c 16 c]); % Syntax csvread(filename,R1,C1,[R1 C1 R2 C2])
+% Column: 1,2,3,4 = POONDI,CHOLAVARAM,REDHILLS,CHEMBARAMBAKKAM
+data = csvread('chennai_reservoir_rainfall_formatted.csv',1,1,[1 1 16 1]); % Syntax csvread(filename,R1,C1,[R1 C1 R2 C2])
+data = [data; csvread('chennai_reservoir_rainfall_formatted.csv',1,2,[1 2 16 2])];
+data = [data; csvread('chennai_reservoir_rainfall_formatted.csv',1,3,[1 3 16 3])];
+data = [data; csvread('chennai_reservoir_rainfall_formatted.csv',1,4,[1 4 16 4])];
+% scale down to be sensible in our example
+data = data / 900;
+
 % mu = mean(data);
 % sigma = std(data);
 
 % rainfall: lognormal distribution with parameters mu and sigma
 % r = lognrnd(mu, sigma, T, 1); % Alternative
-r = exp(mu + sigma.*randn(T,1)); 
+% r = exp(mu + sigma.*randn(T,1));
+if (realdaten == true)
+    r = data;
+end
 
 %% Gauss-Hermite to calculate the expected value of the rain distribution
-n = 10;
-[x_i,w] = GaussHermite(n);
+if (realdaten == false)
+    n = 10;
+    [x_i,w] = GaussHermite(n);
 
-x_trans = sqrt(2)*sigma.*x_i + mu;
+    x_trans = sqrt(2)*sigma.*x_i + mu;
 
-% Expected value of rain
[…]

### 2020-01-16 Code/code.m
restructure trigger functionality

@@ -4,7 +4,7 @@ clear all
 %% Parameters & utility functions
 
 % set realdaten to true or false
-realdaten = true;
+realdaten = false;
 
 % Preferences
 a1 = 1;
@@ -53,13 +53,16 @@ data = data / 900;
 
 % rainfall: lognormal distribution with parameters mu and sigma
 % r = lognrnd(mu, sigma, T, 1); % Alternative
-% r = exp(mu + sigma.*randn(T,1));
-if (realdaten == true)
+if (realdaten)
     r = data;
+else
+    r = exp(mu + sigma.*randn(T,1));
 end
 
 %% Gauss-Hermite to calculate the expected value of the rain distribution
-if (realdaten == false)
+if (realdaten)
+    E_rain = mean(r);
+else
     n = 10;
     [x_i,w] = GaussHermite(n);
 
@@ -68,8 +71,6 @@ if (realdaten == false)
     % Expected value of rain
     E_rain = (1/sqrt(pi)) * w' * exp(x_trans);
     % E_rain = lognstat(mu, sigma) % Alternative
-else
-    E_rain = mean(r);
 end
 
 % Discretization of expected value of rain and rounding to fit our grid 

### 2020-01-16 Code/code.m
code style

@@ -41,7 +41,8 @@ sigma = 1;
 
 % % Realdaten
 % Column: 1,2,3,4 = POONDI,CHOLAVARAM,REDHILLS,CHEMBARAMBAKKAM
-data = csvread('chennai_reservoir_rainfall_formatted.csv',1,1,[1 1 16 1]); % Syntax csvread(filename,R1,C1,[R1 C1 R2 C2])
+% Syntax csvread(filename,R1,C1,[R1 C1 R2 C2])
+data = csvread('chennai_reservoir_rainfall_formatted.csv',1,1,[1 1 16 1]);
 data = [data; csvread('chennai_reservoir_rainfall_formatted.csv',1,2,[1 2 16 2])];
 data = [data; csvread('chennai_reservoir_rainfall_formatted.csv',1,3,[1 3 16 3])];
 data = [data; csvread('chennai_reservoir_rainfall_formatted.csv',1,4,[1 4 16 4])];
@@ -101,7 +102,8 @@ for j = 1:maxIter
     end
     [V, optIrrigation_ind]= max(aux,[],2);  
     
-    % Termination check: Break if norm is smaller then tolerance for all values that are not -inf        
+    % Termination check: Break if norm is smaller then tolerance for all
+    %values that are not -inf
     if norm(V_old(V ~= -inf) - V(V ~= -inf)) < tol
         break;
     end
@@ -170,7 +172,8 @@ plot(steadyStateLvls(1,:));
 plot(steadyStateLvls(2,:));
 plot([1 T],[steadyStateLvl steadyStateLvl],'--g');
 xlim([1 T]);
-legend('water level in reservoir','water used for irrigation','rain','Mean last 50 periods','Difference in mean to previous','steady state level');
+legend('water level in reservoir','water used for irrigation','rain', ...
+    'Mean last 50 periods','Difference in mean to previous','steady state level');
 title('Optimal Irrigation Policy');
 xlabel('period');
 ylabel('amount');

### 2020-01-16 Paper/paper.tex
ChgIntro

@@
 
-with the additional paramter $s$ that indicates the water level at the beginning of the year. For recreational users, the preference parameters shell be $a_2 = 2$ and $b_2 = -3$.\\\\
+with the additional paramter $s$ that indicates the water level at the beginning of the year.
+For recreational users, the preference parameters shell be $a_2 = 2$ and $b_2 = -3$.\\\\
 
@@
 
-Using dynamic programming...
+Using dynamic programming in combination with a stochastic simulation of the rainfall, we determined the optimal irrigation policy that maximizes the combined utility of farmers and citizens for a number of years $T$.
+The sequence of events for every period is displayed in figure x.
+
+\subsubsection{Modification 1: Variation of the discount factor}
+
+An theoretical explanation of the used concepts will be given in chapter x
 

### 2020-01-16 Code/code.m
Chicago Style for Diagramm Titles

@@ -183,7 +183,7 @@ hold off
 figure(3)
 hold on
 histogram(waterLevel(waterInd(steadyStateThreshold:end)),30, 'Normalization','probability');
-title('Steady-State distribution of water level');
+title('Steady-State Distribution of Water Level');
 ylabel('probability');
 xlabel('water level in reservoir');
 hold off

### 2020-01-16 Code/code.m
additional plot showing the optimal irrigation policy

@@ -187,3 +187,12 @@ title('Steady-State Distribution of Water Level');
 ylabel('probability');
 xlabel('water level in reservoir');
 hold off
+
+% plot optimal irrigation policy
+figure (4)
+hold on
+plot(linspace(0,M,dimWL), waterLevel(optIrrigation_ind))
+title('Optimal Irrigation Policy')
+xlabel('water level in reservoir')
+ylabel('irrigation amount')
+hold off

### 2020-01-16 Paper/paper.tex
change style of subsection to arabic numbers
set TeX spellchecker to US english
draft for results section

@@
 %----------------------------------------------------------------------------------------
+% !TeX spellcheck = en_US
 
@@
 %\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
-\renewcommand\thesubsection{\roman{subsection}} % roman numerals for subsections
 \titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
@@
 	\section{Conclusion}
-	\blindtext
+	This chapter will discuss the outcomes and results of the implemented functionality in MATLAB.
+	The discussion will be started on the basic questions asked by the project description.
+	Furthermore we made several modifications to the task which will be discussed additionally.
+	All results will be visualized by plots made within our simulation in MATLAB.
+	\subsection{Optimal irrigation policy}
+	\subsection{Steady State}
+	\subsection{Steady State Probability}
+	\subsection{Effect of Changed Discount Factors to Optimal Irrigation Policy}
+	\subsection{Changed Volatility of Rainfalls}
 

### 2020-01-17 Paper/paper.tex
Introduction

@@
 	\subsection{Motivation}
-With world population expected to hit 9.7 billion in 2050 and 11.2 billion in 2100, more and more people have to be fed, which will increase the demand of water for agriculture uses \citep{un2019}.
+With world population expected to hit 9.7 billion in 2050 and 11.2 billion in 2100, more and more people will have to be fed, which will increase the demand of water for agriculture uses \citep{un2019}.
 Already today, around one third of the world's population live under physical water scarcity \citep{vorosmarty2000global, alcamo2003global, oki2006global}.
@@
 The water reservoir to be considered has two stakeholders: farmers, who want to use the water for irrigation purposes and citizens, who want to use the lake for recreational activities.
+Note that while looking at just two agents is a vast simplification compared to a real allocation problem, one can easily add more stakeholders to the model, as long as there is enough computational power.
 At the beginning of every year we have to decide how much of the water from the reservoir is used for the irrigation of fields, which benefits the farmers' harvest.\\\\
@@
 
-During the summer months, citizens benefit from the remaining water due to recreational use (e.g. swimming). This can be described by the recreational users' utility function which is
+During the summer months, citizens benefit from the remaining water due to recreational use.
+This can be described by the recreational users' utility function which is
 
@@
 
+The sequence of events for every period is displayed in figure x. \\
+
 Using dynamic programming in combination with a stochastic simulation of the rainfall, we determined the optimal irrigation policy that maximizes the combined utility of farmers and citizens for a number of years $T$.
-The sequence of events for every period is displayed in figure x.
+We then go on and modify the original problem in a way that will be described later on. 
+In the main part we will give a brief summary of the underlying theory and the optimization problem.
+After that we will present the results and our conclusions for the original model and its modifications, as well as possible shortfalls of our analysis.
+At the end we also give a little outlook into further possible extensions.
 
-\subsubsection{Modification 1: Variation of the discount factor}
+\subsubsection{Modification 1: Variation of the Discount Factor}
+For our first modification we study the effects of different betas on the optimal irrigation policy. We take a look at the discount factors x and x.
 
-An theoretical explanation of the used concepts will be given in chapter x
+\subsubsection{Modification 2: Increased Rainfall Volatility}
 

### 2020-01-18 Paper/paper.tex
Introduction mostly finished

@@
 Using dynamic programming in combination with a stochastic simulation of the rainfall, we determined the optimal irrigation policy that maximizes the combined utility of farmers and citizens for a number of years $T$.
-We then go on and modify the original problem in a way that will be described later on. 
 In the main part we will give a brief summary of the underlying theory and the optimization problem.
@@
 \subsubsection{Modification 1: Variation of the Discount Factor}
-For our first modification we study the effects of different betas on the optimal irrigation policy. We take a look at the discount factors x and x.
+We also study the effects of different betas on the optimal irrigation policy. 
+The discount factor takes into account that agents in a simple two-stage model often value current consumption over future consumption.
+It expresses how much one unit of future utility is worth in terms of current utility and therefore makes them easily comparable \citep{kruschwitz2014investitionsrechnung}.
+We take a look at the discount factors x and x.
 
 \subsubsection{Modification 2: Increased Rainfall Volatility}
+To consider the predicted impacts of climate change \citep{guhathakurta2011impact}, we look at the effect of increased rainfall volatility. 
+We do this by using a  $\sigma^2$ of x.
+
+\subsubsection{Modification 3: Using Real Data for Rainfall}
+Instead of simulating the rainfall with random numbers, we also use real data.
+Comparing the results then allows us to draw conclusions on the accuracy of the simulation for that region. 
+We use a dataset aquired from \cite{kaggle:2019}, that contains real rainfall data for the four main water reservoirs of the city Chennai, India. % Zitation anpassen
+
+\subsubsection{Modification 4: Taking Evaporation into Account}
+The dataset that contains the real rainfall data also contains the water levels in the reservoirs.
+Using these values we calculate the evaporation of water and include it in our model. \\\\
+
+Note that all modifications, except Modification 4, are based on the original model without any modifications.
+Modification 4 is an extension of Modification 3.
 

### 2020-01-19 Paper/paper.tex
Introduction Alpha

@@
 Note that while looking at just two agents is a vast simplification compared to a real allocation problem, one can easily add more stakeholders to the model, as long as there is enough computational power.
-At the beginning of every year we have to decide how much of the water from the reservoir is used for the irrigation of fields, which benefits the farmers' harvest.\\\\
+The sequence of events for every period is displayed in figure ~\ref{fig:Sequence}.
+At the beginning of every year (Step 1) we have to decide how much of the water from the reservoir is used for the irrigation of fields, which benefits the farmers' harvest.\\\\
 The farmers utility function follows the form
@@
 
-During the summer months, citizens benefit from the remaining water due to recreational use.
+During the summer months, citizens benefit from the remaining water due to recreational use (Step 2).
 This can be described by the recreational users' utility function which is
@@
 
-At the end of the year, water in the reservoir is replenished by rainfall, which we can assume to be randomly following a lognormal distribution \citep{oosterbaan1994frequency}:
+At the end of the year, water in the reservoir is replenished by rainfall (Step 3), which we can assume to be randomly following a lognormal distribution \citep{oosterbaan1994frequency}:
 \begin{equation}
@@
 
-The sequence of events for every period is displayed in figure x. \\
+\begin{figure}[ht]
+	\includegraphics[width=1\textwidth]{figures/CESchaubild_Cut.png}
+	\caption{Sequence of Events in One Period}
+	\label{fig:Sequence}
+\end{figure}
 
@@
 \subsubsection{Modification 4: Taking Evaporation into Account}
-The dataset that contains the real rainfall data also contains the water levels in the reservoirs.
-Using these values we calculate the evaporation of water and include it in our model. \\\\
+The evaporation of water from a water reservoir is the most important natural "outflow" to be considered when it comes to managing water levels \citep{tanny2008evaporation}.
+It therefore should definitely be considered in a real model that is used to determine the optimal irrigation strategy.
+The dataset that we use for the real rainfall also contains the water levels of the four reservoirs.
+Using these values together, we calculate the evaporation of water and include it in our model. \\\\
 

### 2020-01-19 Paper/paper.tex
replace bad umlaut characters

@@
 		\textbf{Belehrung:}\\
-		Wer vorsätzlich gegen eine die Täuschung über Prüfungsleistungen betreffende Regelung einer Hochschulprüfungsordnung verstößt, handelt ordnungswidrig.
+		Wer vorsätzlich gegen eine die Täuschung über Prüfungsleistungen betreffende Regelung einer Hochschulprüfungsordnung verstößt, handelt ordnungswidrig.
 Die Ordnungswidrigkeit kann mit einer Geldbuße bis zu 50 000 Euro geahndet werden.
-Zuständige Verwaltungsbehörde für die Verfolgung und Ahndung von Ordnungswidrigkeiten ist die Kanzlerin oder der Kanzler der RWTH Aachen.
-Im Falle eines mehrfachen oder sonstigen schwerwiegenden Täuschungsversuches kann der Prüfling zudem exmatrikuliert werden (§ 63 Abs.
+Zuständige Verwaltungsbehörde für die Verfolgung und Ahndung von Ordnungswidrigkeiten ist die Kanzlerin oder der Kanzler der RWTH Aachen.
+Im Falle eines mehrfachen oder sonstigen schwerwiegenden Täuschungsversuches kann der Prüfling zudem exmatrikuliert werden (§ 63 Abs.
 5 HG NRW).

### 2020-01-19 Code/code.m
Added code to read in and use real evaporation data in model
Daten evtl. überarbeiten

@@ -4,7 +4,10 @@ clear all
 %% Parameters & utility functions
 
 % set realdaten to true or false
-realdaten = false;
+realdaten = true;
+
+% set use of real Evaporation data to true or false
+real_Evap = true;
 
 % Preferences
 a1 = 1;
@@ -35,31 +38,47 @@ tol = 1e-6;
 T = 64;
 
 %% Rainfall simulation
-% Rainfall parameters (might be overwritten by real data)
+% Rainfall parameters for simulation
 mu = 0;
 sigma = 1;
 
 % % Realdaten
 % Column: 1,2,3,4 = POONDI,CHOLAVARAM,REDHILLS,CHEMBARAMBAKKAM
 % Syntax csvread(filename,R1,C1,[R1 C1 R2 C2])
-data = csvread('chennai_reservoir_rainfall_formatted.csv',1,1,[1 1 16 1]);
-data = [data; csvread('chennai_reservoir_rainfall_formatted.csv',1,2,[1 2 16 2])];
-data = [data; csvread('chennai_reservoir_rainfall_formatted.csv',1,3,[1 3 16 3])];
-data = [data; csvread('chennai_reservoir_rainfall_formatted.csv',1,4,[1 4 16 4])];
+data = csvread('chennai_reservoir_rainfall_formatted.csv',1,1,[1 1 16 4]);
+data = data(:);
+
 % scale down to be sensible in our example
 data = data / 900;
 
+% calculate mu and sigma from the real dataset
 % mu = mean(data);
 % sigma = std(data);
 
 % rainfall: lognormal distribution with parameters mu and sigma
-% r = lognrnd(mu, sigma, T, 1); % Alternative
 if (realdaten)
     r = data;
 else
     r = exp(mu + sigma.*randn(T,1));
 end
 
+%% Potentially read evaporation data
+dimE_Evap = 0;
+dataEvap = zeros(T,1);
+
+if (real_Evap)
+    dataEvap = csvread('chennai_evaporation_filtered_formatted.csv',1,1,[1 1 16 4]);
+    dataEvap = dataEvap(:);
+    
+    % scale down to be sensible in our example
+    dataEvap = dataEvap / 900;
+    
+    % Expected value of evaporation
+    E_Evap = mean(dataEvap);
+    
+    % Discretization of expected value of rain and rounding to fit our grid 
+    dimE_Evap = round(E_Evap/(M/dimWL));
+end
[…]

### 2020-01-20 Code/code.m
set no of periods depending on realdaten true or false

@@ -35,7 +35,11 @@ maxIter = 10000;
 tol = 1e-6;
 
 % Number of periods for forward simulation
-T = 64;
+if (realdaten)
+    T = 64;
+else
+    T = 300;
+end
 
 %% Rainfall simulation
 % Rainfall parameters for simulation

### 2020-01-20 Code/code.m
real evaporation can only be used in combination with real rainfall for now

@@ -7,7 +7,11 @@ clear all
 realdaten = true;
 
 % set use of real Evaporation data to true or false
-real_Evap = true;
+if (realdaten)
+    real_Evap = true;
+else
+    real_evap = false;
+end
 
 % Preferences
 a1 = 1;

### 2020-01-20 Code/code.m
fix typo

@@ -10,7 +10,7 @@ realdaten = true;
 if (realdaten)
     real_Evap = true;
 else
-    real_evap = false;
+    real_Evap = false;
 end
 
 % Preferences

### 2020-01-20 Code/code.m
monte carlo simulation for steady state. Could be put in another script, but works that way as well

@@ -1,226 +1,286 @@
 clc 
 close all
 clear all
+
+%% turn monteCarloSimulation for distribution of the steady state on/off
+monteCarloSimulation = true;
+if(monteCarloSimulation)
+    monteCarloMaxIter = 500;
+else
+    monteCarloMaxIter = 1;
+end
+monteCarloSteadyState = zeros(monteCarloMaxIter,1);
+
+for monteInd=1:monteCarloMaxIter
 %% Parameters & utility functions
 
-% set realdaten to true or false
-realdaten = true;
+    % set realdaten to true or false
+    realdaten = false;
 
-% set use of real Evaporation data to true or false
-if (realdaten)
-    real_Evap = true;
-else
-    real_Evap = false;
-end
+    % set use of real Evaporation data to true or false
+    if (realdaten)
+        real_Evap = true;
+    else
+        real_Evap = false;
+    end
 
-% Preferences
-a1 = 1;
-a2 = 2;
-b1 = -2;
-b2 = -3;
+    % Preferences
+    a1 = 1;
+    a2 = 2;
+    b1 = -2;
+    b2 = -3;
 
-M = 7; % Capacity Reservior
-beta = 0.9; % Discount factor
+    M = 7; % Capacity Reservior
+    beta = 0.9; % Discount factor
 
-% utility functions
-utilFar = @(x) (a1 / (1 + b1)) * x .^ (1 + b1);
-utilRec = @(s,x) (a2 / (1 + b2)) * (s - x) .^ (1 + b2);
+    % utility functions
+    utilFar = @(x) (a1 / (1 + b1)) * x .^ (1 + b1);
+    utilRec = @(s,x) (a2 / (1 + b2)) * (s - x) .^ (1 + b2);
 
-%% Create grid and iteration parameters
+    %% Create grid and iteration parameters
 
-% We need a grid to evaluate the value function for different water levels
-% and different periods.
-% We define dimWL different water levels
-dimWL = 1000;
-waterLevel = linspace(0,M,dimWL);
+    % We need a grid to evaluate the value function for different water levels
+    % and different periods.
+    % We define dimWL different water levels
+    dimWL = 100;
+    waterLevel = linspace(0,M,dimWL);
 
-% Iteration parameters
-maxIter = 10000;
-tol = 1e-6;
+    % Iteration parameters
+    maxIter = 10000;
+    tol = 1e-6;
 
-% Number of periods for forward simulation
-if (realdaten)
[…]

### 2020-01-21 Code/code.m
nicer steady state visualisation

@@ -53,7 +53,7 @@ for monteInd=1:monteCarloMaxIter
     if (realdaten)
         T = 64;
     else
-        T = 300;
+        T = 1000;
     end
 
     %% Rainfall simulation
@@ -191,8 +191,8 @@ for monteInd=1:monteCarloMaxIter
     % Index for water that is used for irrigation
     irrigationInd = zeros(1, T);
     steadyStateLvls = zeros(2, T);
-    steadyStateThreshold = T;
-    periodsForMean = 15;
+    steadyStatePeriod = T;
+    periodsForMean = 200;
     for i=1:T
 
         irrigationInd(i) = optIrrigation_ind(waterInd(i));
@@ -210,20 +210,36 @@ for monteInd=1:monteCarloMaxIter
     end
 
     for i=2:T
-        if (abs(steadyStateLvls(2,i)) < 0.01)
-            steadyStateThreshold = i;
+        if (abs(steadyStateLvls(2,i)) < 0.0001)
+            steadyStatePeriod = i;
+            steadyStateLvl = steadyStateLvls(1,i);
             fprintf('Steady state found in period %s\n',num2str(i));
             break;
         end
     end
-    %steadyStateThreshold = 20; % Change this parameter to determine the period
+    %steadyStatePeriod = 20; % Change this parameter to determine the period
                                % after which the system is expected to be in a 
                                % steady state (has to be smaller then T!)
 
     %% TODO better threshold calculation by using more that one period to evaluate the difference
-    steadyStateLvl = mean(waterLevel(waterInd(steadyStateThreshold:end)));
+    % steadyStateLvl = mean(waterLevel(waterInd(steadyStatePeriod:end)));
     monteCarloSteadyState(monteInd) = steadyStateLvl;
     fprintf('Iteration %s ended with steady State: %s\n',num2str(monteInd), num2str(steadyStateLvl));
+    
+    figure(8)
+    hold on
+    plot(waterLevel(waterInd));
+    plot(steadyStateLvls(1,:));
+    plot(steadyStateLvls(2,:));
+    plot([1 T],[steadyStateLvl steadyStateLvl],'--g');
+    plot([steadyStatePeriod steadyStatePeriod],[0 7],'--g');
+    xlim([1 T]);
+    legend('water level in reservoir',strcat('Mean last ', periodsForMean, ' periods'), ...
+        'Difference in mean to previous', 'steady state level');
+    title('Steady State of Water Level');
+    xlabel('period');
[…]

### 2020-01-21 Code/code.m
highlight area

@@ -232,7 +232,8 @@ for monteInd=1:monteCarloMaxIter
     plot(steadyStateLvls(1,:));
     plot(steadyStateLvls(2,:));
     plot([1 T],[steadyStateLvl steadyStateLvl],'--g');
-    plot([steadyStatePeriod steadyStatePeriod],[0 7],'--g');
+    area = patch([steadyStatePeriod-periodsForMean steadyStatePeriod steadyStatePeriod steadyStatePeriod-periodsForMean],[0 0 7 7],'r');
+    alpha(area,.2)
     xlim([1 T]);
     legend('water level in reservoir',strcat('Mean last ', periodsForMean, ' periods'), ...
         'Difference in mean to previous', 'steady state level');

### 2020-01-23 Code/code.m
plot styling

@@ -225,22 +225,6 @@ for monteInd=1:monteCarloMaxIter
     % steadyStateLvl = mean(waterLevel(waterInd(steadyStatePeriod:end)));
     monteCarloSteadyState(monteInd) = steadyStateLvl;
     fprintf('Iteration %s ended with steady State: %s\n',num2str(monteInd), num2str(steadyStateLvl));
-    
-    figure(8)
-    hold on
-    plot(waterLevel(waterInd));
-    plot(steadyStateLvls(1,:));
-    plot(steadyStateLvls(2,:));
-    plot([1 T],[steadyStateLvl steadyStateLvl],'--g');
-    area = patch([steadyStatePeriod-periodsForMean steadyStatePeriod steadyStatePeriod steadyStatePeriod-periodsForMean],[0 0 7 7],'r');
-    alpha(area,.2)
-    xlim([1 T]);
-    legend('water level in reservoir',strcat('Mean last ', periodsForMean, ' periods'), ...
-        'Difference in mean to previous', 'steady state level');
-    title('Steady State of Water Level');
-    xlabel('period');
-    ylabel('amount');
-    hold off
 end
 
 if(monteCarloSimulation)
@@ -253,6 +237,25 @@ if(monteCarloSimulation)
     xlabel('water level in reservoir');
     hold off
 else
+    figure(8)
+    hold on
+    plot(waterLevel(waterInd));
+    plot(steadyStateLvls(1,:));
+    plot(steadyStateLvls(2,:));
+    plot([1 T],[steadyStateLvl steadyStateLvl],'--g');
+    area = patch([steadyStatePeriod-periodsForMean steadyStatePeriod steadyStatePeriod steadyStatePeriod-periodsForMean],[0 0 7 7],'r');
+    alpha(area,.2)
+    xlim([1 T]);
+    ylim([-1 11]);
+    legend('water level in reservoir',append('Mean last ', num2str(periodsForMean), ' periods'), ...
+        'Difference in mean to previous', 'steady state level');
+    title('Steady State of Water Level');
+    xlabel('period');
+    ylabel('amount');
+    set(gca,'FontSize',12)
+    set(gcf,'Units','Centimeters','position',[0,0,16,12]);
+    hold off
+    
     % Plot water level, amount of water used for irrigation and steady state
     % value
     figure(2)

### 2020-01-23 Code/theoryPlots.m
Create theoryPlots.m

@@ -1,0 +1,12 @@
+clc;
+clear;
+f = @(t) sin(t*pi);
+x1 = linspace(0,1, 101);
+x2 = linspace(0,1, 11);
+figure
+hold on
+bar(x2,f(x2), 1, "w")
+plot(x2,f(x2), "o", "color", "k")
+plot(x1,f(x1), "color", "#0072BD")
+
+sum(f(linspace(0,1,101))/101)

### 2020-01-23 Paper/paper.tex
temporary changes
plot graphs using matlab and include them in latex

@@
 	\section{Main Part}
-	\subsection{bla}
-	\begin{figure}[ht] % TODO remove example
-		\includegraphics[width=1\textwidth]{figures/example.png}
-		\caption{Example Image}
-		\label{fig:example-pic}
-	\end{figure}
+	\subsection{Theory}
+	(Für die Präsentation)
+
+%	\begin{equation}
+%	V_t =  max\{u_{farmer}(irrigation_t) + u_{recreationalUser}(irrigation_t, waterLevel_t) \\
+%		
+%	+ 	\beta*(u_{farmer}(irrigation_{t-1}) + u_{recreationalUser}(irrigation_{t-1}, waterLevel_{t-1})\}
+%	\end{equation}
+
 	\blindtext
-	\begin{figure}[ht] % TODO remove example
+	\begin{figure}[ht]
 		\begin{longtable}{|p{2,5cm}|p{2,5cm}|p{2,5cm}|p{3cm}|p{3cm}|}
@@
 	\subsection{Optimal irrigation policy}
+	When questioning what the optimal irrigation policy is, the first question to be asked is how the aggregated value function of both recreational users and farmers looks like.
+	The value function as posed in % TODO put links to theory here
+	contains the maximum possible usage that can be gained by irrigating a specific amount of water under a given water level. 
+	\ref{fig:value-function} show the value function calculated for all possible water levels between the allowed boundaries of 0 and 7.
+	\begin{figure}[ht]
+		\includegraphics{figures/value_function.eps}
+		\caption{Value Function}
+		\label{fig:value-function}
+	\end{figure}
+	\begin{figure}[ht]
+		\includegraphics{figures/OptimalIrrigationPolicy.eps}
+		\caption{Optimal Irrigation Policy}
+		\label{fig:optimal-irrigation-policy}
+	\end{figure}
 	\subsection{Steady State}
+	\begin{figure}[ht]
+		\includegraphics{figures/steadyState.eps}
+		\caption{Exemplary Calculation of the Steady State}
+		\label{fig:steadyState}
+	\end{figure}
 	\subsection{Steady State Probability}
+	\begin{figure}[ht]
+		\includegraphics{figures/steadyStateMonteCarlo.eps}
+		\caption{Steady State Probability}
+		\label{fig:steadyStateMonteCarlo}
+	\end{figure}
 	\subsection{Effect of Changed Discount Factors to Optimal Irrigation Policy}

### 2020-01-23 Paper/paper.tex
begin adding some theory

@@
 	\subsection{Theory}
+
+	In this paper, the theoretical knowledge obtained in the seminar shall be applied.
+
+	\subsubsection{Quadrature}
+	In many computational applications, formulas 	are used to calculate values of any kind. For example, let the flow of water through a pipe be
+
+	\begin{equation}
+	f(t) = sin(\pi t), \; t \in [0,1]
+	\end{equation}
+
+Now, it's trivial to compute the flow at any given time using code, simply by implementing the formula as given above.
+However, if we want to compute the amount of water flowing through the pipe in the interval $[0.25, 0.5]$, we can not calculate that directly, as we would have to know the antiderivative of $f(t)$.
+In this example, this might be solved analytically
+
+---
+
 	(Für die Präsentation)

### 2020-01-23 Code/theoryPlots.m
Update theoryPlots.m

@@ -8,5 +8,9 @@ hold on
 bar(x2,f(x2), 1, "w")
 plot(x2,f(x2), "o", "color", "k")
 plot(x1,f(x1), "color", "#0072BD")
-
+ylabel('f(t)');
+xlabel('t');
+set(gca,'FontSize',8)
+set(gcf,'Units','Centimeters','position',[0,0,16,6]);
+saveas(gcf,'quadratureA.eps', 'epsc');
 sum(f(linspace(0,1,101))/101)

### 2020-01-23 Paper/paper.tex
font size 8. Include eps files in repo
add section quadrature

@@
 Now, it's trivial to compute the flow at any given time using code, simply by implementing the formula as given above.
-However, if we want to compute the amount of water flowing through the pipe in the interval $[0.25, 0.5]$, we can not calculate that directly, as we would have to know the antiderivative of $f(t)$.
-In this example, this might be solved analytically
+However, if we want to compute the amount of water flowing through the pipe in the time interval $[0.25, 0.5]$, we can not calculate that directly, as we would have to know the antiderivative of $f(t)$.
+Apart from this example, it's not always trivial to calculate the antiderivative of a function analytically.
+However it is possible compute the area below (almost) any function using numerical integration, also called numerical quadrature.
 
----
+A naive implementation of a quadrature could be to calculate the total area of equidistant rectangles whose height corresponds to the function value of $f$:
 
-	(Für die Präsentation)
+	\begin{figure}[ht] 
+		\includegraphics[width=1\textwidth]{figures/quadratureA.eps}
+		\caption{Quadrature using Midpoint Rule}
+		\label{fig:quadrature-a}
+	\end{figure}
+This implementation of a quadrature is called \emph{Midpoint Rule}. 
+\\
 
-%	\begin{equation}
-%	V_t =  max\{u_{farmer}(irrigation_t) + u_{recreationalUser}(irrigation_t, waterLevel_t) \\
-%		
-%	+ 	\beta*(u_{farmer}(irrigation_{t-1}) + u_{recreationalUser}(irrigation_{t-1}, waterLevel_{t-1})\}
-%	\end{equation}
+A more sophisticated way to determine the area of an unbounded integral is to use the Gauss-Hermite Formula
 
+	\begin{equation}
+	\int_{-\infty}^{\infty} f(x)e^{-x^2} dx \approx \sum_{i=1}^{n}{w_if(x_i)}
+	\end{equation}
+
+where the weights $w_i$ are given as the solution to a linear system of equations and the nodes $x_i$ correspond to the zeros of the nth Hermite polynomial. \cite{seminar:week6}
+
+	\subsection{bla}
+	\begin{figure}[ht] % TODO remove example
+		\includegraphics[width=1\textwidth]{figures/example.png}
+		\caption{Example Image}
+		\label{fig:example-pic}
+	\end{figure}
 	\blindtext
-	\begin{figure}[ht]
+	\begin{figure}[ht] % TODO remove example
 		\begin{longtable}{|p{2,5cm}|p{2,5cm}|p{2,5cm}|p{3cm}|p{3cm}|}
@@
 	\ref{fig:value-function} show the value function calculated for all possible water levels between the allowed boundaries of 0 and 7.
+	The plot shows a rising proposed irrigation amount with rising water level up to a water level of approximately 1.5 where a declining irrigation amount is favored.
 	\begin{figure}[ht]
-		\includegraphics{figures/value_function.eps}
+		\includegraphics[width=1\textwidth]{figures/value_function.eps}
 		\caption{Value Function}
@@
 	\begin{figure}[ht]
-		\includegraphics{figures/OptimalIrrigationPolicy.eps}
+		\includegraphics[width=1\textwidth]{figures/OptimalIrrigationPolicy.eps}
 		\caption{Optimal Irrigation Policy}
@@
 	\begin{figure}[ht]
-		\includegraphics{figures/steadyState.eps}
+		\includegraphics[width=1\textwidth]{figures/steadyState.eps}
 		\caption{Exemplary Calculation of the Steady State}
@@
 	\begin{figure}[ht]
-		\includegraphics{figures/steadyStateMonteCarlo.eps}
+		\includegraphics[width=1\textwidth]{figures/steadyStateMonteCarlo.eps}
 		\caption{Steady State Probability}

### 2020-01-23 Paper/paper.tex
increase line space
further work in results

@@
 \usepackage{booktabs} % Horizontal rules in tables
+
+\usepackage[onehalfspacing]{setspace}
 
@@
 	%------------------------------------------------
-
-	\section{Conclusion}
+	\newpage
+	\section{Results}
 	This chapter will discuss the outcomes and results of the implemented functionality in MATLAB.
@@
 	contains the maximum possible usage that can be gained by irrigating a specific amount of water under a given water level. 
-	\ref{fig:value-function} show the value function calculated for all possible water levels between the allowed boundaries of 0 and 7.
-	The plot shows a rising proposed irrigation amount with rising water level up to a water level of approximately 1.5 where a declining irrigation amount is favored.
+	Figure \ref{fig:value-function} shows the value function calculated for all possible water levels between the allowed boundaries of 0 and 7. 
 	\begin{figure}[ht]
@@
 	\end{figure}
+	\newline
+	Every value along the possible water level leads to a irrigation amount that optimizes the value for all users in the given situation.
+Summarizing these values over the range of possible water levels results in a optimal irrigation policy which is shown in figure \ref{fig:optimal-irrigation-policy}.
+	The plot shows a rising proposed irrigation amount with rising water level. There are 2 particularities that can be found within the graph:
+	\begin{itemize}
+		\item Due to the effect that a almost empty reservoir has a very low use to recreational users, the optimal irrigation strategy suggests a very low irrigation amount whose gradient increases with increasing water level.
+		\item The diminishing margin utility of farmers lead to a decreasing gradient for the proposed irrigation amount with increasing water level.
+	\end{itemize}
 	\begin{figure}[ht]
@@
 	\subsection{Steady State}
+	% TODO maybe move to theory?
+	In systems theory, a system or a process is in a steady state if the variables which define the behavior of the system are unchanging in time \cite{gagniuc2017markov}. 
+	The variable defining the behavior of our system is the water level. 
+	Because of this, we have to calculate the difference in water level for each period. 
+	Due to the fact, that our water level highly depends on the randomly distributed rainfalls that rise up the water level at the end of each period, it is not possible to find a state steady. 
+	To accomplish this, we defined our steady state, to be a state, where the mean water level of the last 200 periods did not change more than $1x10^{-7}$ from a period to the next.
+	A exemplary graph of the mean in every period and the difference to its predecessor is shown in figure \ref{fig:steadyState}.
+	Once a period is found, where the mean does not change more than the given value, the mean of that period is defined as steady state water level.
+	The water level found to be a steady state and the period in which this condition occurs changes from calculation to calculation due to the fact, that the rain fall is still a random distribution.
 	\begin{figure}[ht]
@@
 	\subsection{Steady State Probability}
+	As mentioned above, the steady state changes from simulation to simulation.
+	In order to find out, how the steady states are distributed and what the probability of their occurrences are, we did a Monte Carlo simulation.
+	To do so, we ran our code 1000 times and put the resulting water levels into bins with the size of 0.1 each.
+	We then counted the occurrences, normalized them and calculated the corresponding probability.
+	The result is put in a histogram and shown in figure \ref{fig:steadyStateMonteCarlo}
 	\begin{figure}[ht]

### 2020-01-25 Code/code.m
Einlesen neuer Realdaten + Verschieben MC-SS
Einlesen neuer Realdaten für Regen und Errechnen von Evaporation aus Altitude, Latitude und durchschnittlichen Temperaturen.
Außerdem habe ich den Beginn des Loops für die Monte Carlo Simulation des Steady States auf kurz vor die Forward Simulation verschoben. Jetzt kann das auch mit feinem Raster recht flott ablaufen, da die Value Function nur 1x berechnet wird

@@ -11,178 +11,208 @@ else
 end
 monteCarloSteadyState = zeros(monteCarloMaxIter,1);
 
-for monteInd=1:monteCarloMaxIter
 %% Parameters & utility functions
 
-    % set realdaten to true or false
-    realdaten = false;
+% set realdaten to true or false
+realdaten = false;
 
-    % set use of real Evaporation data to true or false
-    if (realdaten)
-        real_Evap = true;
-    else
-        real_Evap = false;
-    end
+% Define which real data you want to use. Has to correspond exactly to
+% names from dataset
+useRealDataFrom = 'Barisal';
 
-    % Preferences
-    a1 = 1;
-    a2 = 2;
-    b1 = -2;
-    b2 = -3;
+% set use of real Evaporation data to true or false
+if (realdaten)
+    real_Evap = true;
+else
+    real_Evap = false;
+end
 
-    M = 7; % Capacity Reservior
-    beta = 0.9; % Discount factor
+% Preferences
+a1 = 1;
+a2 = 2;
+b1 = -2;
+b2 = -3;
 
-    % utility functions
-    utilFar = @(x) (a1 / (1 + b1)) * x .^ (1 + b1);
-    utilRec = @(s,x) (a2 / (1 + b2)) * (s - x) .^ (1 + b2);
+M = 7; % Capacity Reservior
+beta = 0.9; % Discount factor
 
-    %% Create grid and iteration parameters
+% utility functions
+utilFar = @(x) (a1 / (1 + b1)) * x .^ (1 + b1);
+utilRec = @(s,x) (a2 / (1 + b2)) * (s - x) .^ (1 + b2);
 
-    % We need a grid to evaluate the value function for different water levels
-    % and different periods.
-    % We define dimWL different water levels
-    dimWL = 100;
-    waterLevel = linspace(0,M,dimWL);
+%% Create grid and iteration parameters
 
-    % Iteration parameters
-    maxIter = 10000;
-    tol = 1e-6;
+% We need a grid to evaluate the value function for different water levels
+% and different periods.
+% We define dimWL different water levels
+dimWL = 1000;
+waterLevel = linspace(0,M,dimWL);
 
-    % Number of periods for forward simulation
-    if (realdaten)
-        T = 64;
-    else
-        T = 1000;
-    end
+% Iteration parameters
+maxIter = 10000;
+tol = 1e-6;
 
-    %% Rainfall simulation
-    % Rainfall parameters for simulation
-    mu = 0;
-    sigma = 1;
[…]

### 2020-01-25 Paper/paper.tex
Modified introduction

@@
 Comparing the results then allows us to draw conclusions on the accuracy of the simulation for that region. 
-We use a dataset aquired from \cite{kaggle:2019}, that contains real rainfall data for the four main water reservoirs of the city Chennai, India. % Zitation anpassen
+We use a dataset aquired from \cite{kaggle:2019}, that contains real rainfall data for the last 65 years in different regions of Bangladesh. % Zitation anpassen
 
@@
 It therefore should definitely be considered in a real model that is used to determine the optimal irrigation strategy.
-The dataset that we use for the real rainfall also contains the water levels of the four reservoirs.
-Using these values together, we calculate the evaporation of water and include it in our model. \\\\
+The dataset that we use for the real rainfall also contains many other weather parameters.
+With the mean temperature $T$, the elevation $h$, the latitude $A$ and the dew point $T_d$ we calculate an estimate for evaporation of water in mm per day for that specific location, using the formula \citep{linacre1977simple}
+
+\begin{equation}
+	{E_0} = \frac{700 * (T + 0.006h) / (100 - A) + 15 * (T - T_d)}{(80 - T)}
+\end{equation}
+
+and then multiply the result with 365.25 to get the estimated evaporation in mm per year. 
+Note, however, that due to a lack of data on the real mean temperature in a year, we use the average of the minimum and maximum temperature of every month for all months in a year as $T$. \\
+The average dew point $T_d$ can be estimated by
+
+\begin{equation}
+	{T_d} = T - (\frac{100 - RH}{5})
+\end{equation}
+
+where RH is the average relative humidity of the year in percent \citep{lawrence2005relationship}.\\\\
 

### 2020-02-03 Code/code.m
Improved code structure, standardized code style, put all parameters in the top section, added/improved/standardized comments, ...

@@ -1,89 +1,124 @@
-clc 
-close all
-clear all
+%% Reset the Script Environment
+clc;
+close all;
+clear all;
 
-%% turn monteCarloSimulation for distribution of the steady state on/off
-monteCarloSimulation = true;
-if(monteCarloSimulation)
-    monteCarloMaxIter = 500;
-else
-    monteCarloMaxIter = 1;
-end
-monteCarloSteadyState = zeros(monteCarloMaxIter,1);
+%% Simulation Parameters 
 
-%% Parameters & utility functions
+% Switch to use a monte Carlo simulation to determine the steady state
+useMonteCarloSimulation = true;
 
-% set realdaten to true or false
-realdaten = false;
+% The count of iterations to be used in the monte carlo simulation, if
+% enabled
+monteCarloMaxIter = 500;
 
-% Define which real data you want to use. Has to correspond exactly to
-% names from dataset
-useRealDataFrom = 'Barisal';
+% Switch to use real rainfall and evaporation data 
+useRealData = false;
 
-% set use of real Evaporation data to true or false
-if (realdaten)
-    real_Evap = true;
-else
-    real_Evap = false;
-end
+% Set the file name of the weather dataset
+realDataSource = 'bangladesh_weather_formatted.csv';
 
-% Preferences
+% Set the name of the location in the weather dataset to be used to
+% simulate the evaporation
+locationName = 'Barisal';
+
+% Parameters of the value function
 a1 = 1;
 a2 = 2;
 b1 = -2;
 b2 = -3;
 
-M = 7; % Capacity Reservior
-beta = 0.9; % Discount factor
+% Total capacity of the reservoir
+M = 7;
 
-% utility functions
-utilFar = @(x) (a1 / (1 + b1)) * x .^ (1 + b1);
-utilRec = @(s,x) (a2 / (1 + b2)) * (s - x) .^ (1 + b2);
+% Discount factor
+beta = 0.9;
 
-%% Create grid and iteration parameters
-
-% We need a grid to evaluate the value function for different water levels
-% and different periods.
-% We define dimWL different water levels
-dimWL = 1000;
-waterLevel = linspace(0,M,dimWL);
-
-% Iteration parameters
-maxIter = 10000;
-tol = 1e-6;
-
-% Number of periods for forward simulation (will be overwritten if
-% using realdaten
+% Number of periods for the forward simulation, will be overwritten when
+% using real data
 T = 1000;
 
-%% Rainfall simulation
-% Rainfall parameters for simulation
[…]

### 2020-02-10 Code/code.m
add comment

@@ -173,6 +173,7 @@ for j = 1:valueFunctionMaxIter
             aux(iWL, iIrrigation) = utilFar(waterLevel(iIrrigation)) +...
                 utilRec(waterLevel(iWL), waterLevel(iIrrigation)) +...
                 beta*V_old(min(max(iWL-iIrrigation+1+dimE_Rain-dimE_Evap,1),dimWL));
+            % Generate aux_farm and aux_rec for plotting in a later step 
             aux_farm(iWL, iIrrigation) = utilFar(waterLevel(iIrrigation));
             aux_rec(iWL, iIrrigation) = utilRec(waterLevel(iWL), waterLevel(iIrrigation));
         end

### 2020-02-10 Paper/paper.tex
small changes
start writing chapters dynamic programming and implementation

@@
 			\textbf{Martin Wicke} \\
-			stud. M.Sc. Business Administration and Engineering: Electrical Power Engineering\\
+			stud. 4th semester M.Sc. Business Administration and Engineering: Electrical Power Engineering\\
 			Matr. Nr. 336301 \\
@@
 		\includegraphics[width=1\textwidth]{figures/quadratureA.eps}
-		\caption{Quadrature using Midpoint Rule}
+		\caption{Quadrature using the Midpoint Rule}
 		\label{fig:quadrature-a}
@@
 
-where the weights $w_i$ are given as the solution to a linear system of equations and the nodes $x_i$ correspond to the zeros of the nth Hermite polynomial. \cite{seminar:week6}
+where the weights $w_i$ are given as the solution to a linear system of equations and the nodes $x_i$ correspond to the zeros of the nth Hermite polynomial. (\cite{seminar:week6})
+
+\subsection{Dynamic Programming}
+
+A common technique to solve optimization or combinatorial problems is called Dynamic Programming.
+It is based on the idea of dividing problems into smaller subproblems, solving these recursively and caching intermediate results (\textit{memoization}).
+In his paper "The Theory Of Dynamic Programming" from 1957, Richard Bellman aimed to provide a solution to multi-stage-decision processes.
+He defines a \texit{policy} as "a sequence of decisions" (\cite{bellman:1957}, p.2), and an \textit{optimal policy} as "a policy, which is most advantageous according to some preassigned criterion" (\cite{bellman:1957}, p.2).
+This is reflected in Bellman's Principle of Optimality: "An optimal policy has the property that whatever initial state and initial decisions are, the remaining decisions must constitute an optimal policy with regard o the state resulting from the first decisions" (\cite{bellman:1957}, p.2).
+
+\subsection{Implementation} % TODO: Better title?
+
+In a first step, we want to compute the value function $V$ for each state $s$.
+In order to do so, we need to calculate an auxiliary matrix $aux$ that contains the combined utilities of the farmers and the recreational users.
+Obviously, the $aux$ matrix can only a finite, discrete amount of values.
+In this work, a resolution of $n = 1000$ different water levels and irrigation amounts should suffice.
+Thus we discretize both water level $s$ and irrigation amount $x$ as vectors
+$$
+s = \left( \begin{array}{c}0\\...\\7\end{array}\right), \hspace{3mm}x = \left( \begin{array}{c}0\\...\\7\end{array}\right)
+$$
+Note that the maximum reservoir capacity is $7$ and $|s|=|x| = n = 1000$.
+
+For a combined utility of the farmers and the recreational users  $u(s_i,x_j) = F(x_j) + G(s_i,x_j)$, depending on the water level $s_i$ and the amount of irrigation $x_i$ we get an auxiliary matrix
+$$
+aux = \left[ \begin{array}{ccccc}                                u(s_1, x_1)& \varnothing & \varnothing & ...
+& \varnothing \\                                               u(s_2, x_1)& u(s_2, x_2)& \varnothing & ...
+& \varnothing \\                                               u(s_3, x_1)& ...
+& u(s_3, x_3)& ...
+& \varnothing \\                                              ...
+& ...
+& ...
+& ...
+& \varnothing \\                                              u(s_n, x_1)& ...
+& ...
+& ...
+& u(s_n, x_n)\\                                               \end{array}\right]\\\\\\+\beta*\left[ \begin{array}{ccccc}                                V'(s_1') & \varnothing & \varnothing & ...
+& \varnothing \\                                               V'(s_1') & V'(s_2') & \varnothing & ...
+& \varnothing \\                                               V'(s_1') & ...
+& V'(s_3') & ...
+& \varnothing \\                                              ...
+& ...
+& ...
+& ...
+& \varnothing \\                                              V'(s_1') & ...
+& ...
+& ...
+& V'(s_n') \\                                               \end{array}\right]
+$$
+where upper right triangle of this matrix will be empty, as the amount of water, that can be taken from the reservoir must be smaller or equal to the amount of water in the reservoir.
+
+Here, $V'$ represents the value for $V$ in the previous iteration and will be set to $V'(s'_1) = (0,...0)^T$. Iteratively, we use the bellman equation to calculate a new optimal V:
+
+$$V(s) = \underset{x}{max}(u(s,x) + \beta*V'(s')))$$
+
+which equals to finding the column $aux$ that maximizes the expected value over all states (TODO: kann man das so sagen?).
+We repeat this step over and over again until the our value function $V$ converges.
 

### 2020-02-11 Code/code.m
comments

@@ -178,6 +178,8 @@ for j = 1:valueFunctionMaxIter
             aux_rec(iWL, iIrrigation) = utilRec(waterLevel(iWL), waterLevel(iIrrigation));
         end
     end
+    % Compute V and the optimum irrigation index (for each water level
+    % step) using the bellman equation
     [V, optIrrigation_ind]= max(aux,[],2);  
 
     % Termination check: break if the norm is smaller than the tolerance 
@@ -252,6 +254,9 @@ for monteInd=1:monteCarloMaxIter
     
     % Steady state levels
     steadyStateLvls = zeros(2, T);
+    % Initialize the steady state period with the amount of
+    % periods of the forward iteration in case no steady state will be
+    % found
     steadyStatePeriod = T;
     
     % Perform the forward iteration

### 2020-02-11 Paper/paper.tex
working on section bellman equation and implementation

@@
 \usepackage[english]{babel} % Language hyphenation and typographical rules
+
+\usepackage{breqn} % for line breaks in equations
 
@@
 
+\subsection{The Bellman Equation}
+
+The so-called "Bellman Equation" answers the question, assuming a given state, what long-term utility can be expected by taking the best possible action now and at each subsequent step - or simply put: "What is the value $V$ of a state $s$?".
+
+Mathematically, we can express this as  
+\begin{equation}
+V(s) = \max_{x}(u(s,x) + \beta*V'(s')))
+\end{equation}
+
+where $u(s,x)$ is the utility to gain from the control variable (action) $x$ in state $s$ [TODO: QUOTE Seminar Week 3].
+
+So the value of the control variable to be set is the one that maximizes both the utility of the current state plus (recursively) the value $V$ of the next state $s'$ which is discounted by a factor of $\beta$.
+
 \subsection{Implementation} % TODO: Better title?
@@
 Thus we discretize both water level $s$ and irrigation amount $x$ as vectors
-$$
-s = \left( \begin{array}{c}0\\...\\7\end{array}\right), \hspace{3mm}x = \left( \begin{array}{c}0\\...\\7\end{array}\right)
-$$
+
+	\begin{equation}
+s = \left( \begin{array}{c}0\\...\\7\end{array}\right),\hspace{2mm}x = \left( \begin{array}{c}0\\...\\7\end{array}\right)
+\end{equation}
+
 Note that the maximum reservoir capacity is $7$ and $|s|=|x| = n = 1000$.
@@
 For a combined utility of the farmers and the recreational users  $u(s_i,x_j) = F(x_j) + G(s_i,x_j)$, depending on the water level $s_i$ and the amount of irrigation $x_i$ we get an auxiliary matrix
-$$
-aux = \left[ \begin{array}{ccccc}                                u(s_1, x_1)& \varnothing & \varnothing & ...
-& \varnothing \\                                               u(s_2, x_1)& u(s_2, x_2)& \varnothing & ...
-& \varnothing \\                                               u(s_3, x_1)& ...
+
+\begin{dmath}
+aux &= \left[
+ \begin{array}{ccccc} 
+u(s_1, x_1)& \emptyset & \emptyset & ...
+& \emptyset \\ u(s_2, x_1)& u(s_2, x_2)& \emptyset & ...
+& \emptyset \\                                               u(s_3, x_1)& ...
 & u(s_3, x_3)& ...
-& \varnothing \\                                              ...
+& \emptyset \\                                              ...
 & ...
@@
 & ...
-& \varnothing \\                                              u(s_n, x_1)& ...
+& \emptyset \\                                              u(s_n, x_1)& ...
 & ...
 & ...
-& u(s_n, x_n)\\                                               \end{array}\right]\\\\\\+\beta*\left[ \begin{array}{ccccc}                                V'(s_1') & \varnothing & \varnothing & ...
-& \varnothing \\                                               V'(s_1') & V'(s_2') & \varnothing & ...
-& \varnothing \\                                               V'(s_1') & ...
+& u(s_n, x_n)\\                                               \end{array}\right]+\beta*\left[ \begin{array}{ccccc}
+V'(s_1') & \emptyset & \emptyset & ...
+& \emptyset \\                                               V'(s_1') & V'(s_2') & \emptyset & ...
+& \emptyset \\                                               V'(s_1') & ...
 & V'(s_3') & ...
-& \varnothing \\                                              ...
+& \emptyset \\                                              ...
 & ...
@@
 & ...
-& \varnothing \\                                              V'(s_1') & ...
+& \emptyset \\                                              V'(s_1') & ...
 & ...
@@
 & V'(s_n') \\                                               \end{array}\right]
-$$
-where upper right triangle of this matrix will be empty, as the amount of water, that can be taken from the reservoir must be smaller or equal to the amount of water in the reservoir.
+\end{dmath}
+
+where the upper right triangle of this matrix will be empty, as the amount of water, that can be taken from the reservoir must be smaller or equal to the amount of water in the reservoir.
 
@@
 
-$$V(s) = \underset{x}{max}(u(s,x) + \beta*V'(s')))$$
+\begin{equation}
+V(s) = \max_{x}(u(s,x) + \beta*V'(s')))
+\end{equation}
 
 which equals to finding the column $aux$ that maximizes the expected value over all states (TODO: kann man das so sagen?).
-We repeat this step over and over again until the our value function $V$ converges.
+We repeat this step over and over again until the our value function $V$ converges (technically: the norm of the difference between $V'$ and $V$ falls below a certain tolerance value).
+
+Finally, we want to compute the value for the steady state.
 

### 2020-02-11 Code/code.m
move calulation of value function to a seperate function (making it easier to parameterize it)

@@ -150,45 +150,7 @@ else
     dimE_Evap = 0;
 end
 
-%% Computation of the Value Function
-
-% Create a value function for the combined value of farmers and 
-% recreational users for each water level, respecting the discounted value
-% for the next period and the expected rainfall and evaporation
-V = zeros(dimWL,1);
-
-tic
-for j = 1:valueFunctionMaxIter
-    V_old = V;
-
-    % Create an auxiliary matrix for each water level and depending on that
-    % each possible amount of water to be used for irrigation. Then use
-    % the auxiliary matrix to find the maximum over all amounts of water 
-    % used for irrigation for each water level in the reservoir
-    aux = zeros(dimWL, dimWL) + NaN;
-    aux_farm = zeros(dimWL, dimWL) + NaN;
-    aux_rec = zeros(dimWL, dimWL) + NaN;
-    for iWL = 1:dimWL
-        for iIrrigation = 1:iWL
-            aux(iWL, iIrrigation) = utilFar(waterLevel(iIrrigation)) +...
-                utilRec(waterLevel(iWL), waterLevel(iIrrigation)) +...
-                beta*V_old(min(max(iWL-iIrrigation+1+dimE_Rain-dimE_Evap,1),dimWL));
-            % Generate aux_farm and aux_rec for plotting in a later step 
-            aux_farm(iWL, iIrrigation) = utilFar(waterLevel(iIrrigation));
-            aux_rec(iWL, iIrrigation) = utilRec(waterLevel(iWL), waterLevel(iIrrigation));
-        end
-    end
-    % Compute V and the optimum irrigation index (for each water level
-    % step) using the bellman equation
-    [V, optIrrigation_ind]= max(aux,[],2);  
-
-    % Termination check: break if the norm is smaller than the tolerance 
-    % (excluding -inf values)
-    if norm(V_old(V ~= -inf) - V(V ~= -inf)) < valueFunctionTolerance
-        break;
-    end
-toc
-end
+[V, optIrrigation_ind, aux_farm, aux_rec] = ValueFunction(dimWL, valueFunctionMaxIter, waterLevel, dimE_Rain, dimE_Evap, utilFar, utilRec, beta, valueFunctionTolerance);
 
 
 % Do not print all the diagrams if running monte carlo simulation
@@ -368,4 +330,47 @@ else
     xlabel('Period');
     ylabel('Amount');
     hold off
+end
+
[…]

### 2020-02-11 Code/code.m
plot impact of different betas on optimal irrigation policy

@@ -330,6 +330,20 @@ else
     xlabel('Period');
     ylabel('Amount');
     hold off
+
+    colormap_jet = colormap(parula);
+    % Plot optimal irrigation policy
+    figure(9)
+    hold on
+    for beta=1:-0.05:0.1
+        % Computation of the Value Function
+        [V, optIrrigation_ind, aux_farm, aux_rec] = ValueFunction(dimWL, valueFunctionMaxIter, waterLevel, dimE_Rain, dimE_Evap, utilFar, utilRec, beta, valueFunctionTolerance);
+        plot(linspace(0,M,dimWL), waterLevel(optIrrigation_ind), 'color', colormap_jet(round(256*beta),:))
+    end
+    title('Optimal Irrigation Policy')
+    xlabel('Water Level of the Reservoir')
+    ylabel('Irrigation Amount')
+    hold off
 end
 
 function [V, optIrrigation_ind, aux_farm, aux_rec] = ValueFunction(dimWL, valueFunctionMaxIter, waterLevel, dimE_Rain, dimE_Evap, utilFar, utilRec, beta, valueFunctionTolerance)

### 2020-02-11 Code/code.m
add comments, format too long lines

@@ -6,14 +6,14 @@ clear all;
 %% Simulation Parameters 
 
 % Switch to use a monte Carlo simulation to determine the steady state
-useMonteCarloSimulation = true;
+useMonteCarloSimulation = false;
 
 % The count of iterations to be used in the monte carlo simulation, if
 % enabled
 monteCarloMaxIter = 500;
 
 % Switch to use real rainfall and evaporation data 
-useRealData = false;
+useRealData = true;
 
 % Set the file name of the weather dataset
 realDataSource = 'bangladesh_weather_formatted.csv';
@@ -35,7 +35,7 @@ M = 7;
 beta = 0.9;
 
 % Number of periods for the forward simulation, will be overwritten when
-% using real data
+% using real data with the actually available amount of period data
 T = 1000;
 
 % Parameters for rainfall distribution, if simulated
@@ -62,7 +62,7 @@ steadyStateTolerance = 0.0001;
 
 if(useRealData)
     % Load weather dataset
-    data = readtable(weatherDataSource);
+    data = readtable(realDataSource);
 end
 
 % Initialize the monte carlo simulation
@@ -211,10 +211,10 @@ for monteInd=1:monteCarloMaxIter
     % Set the index to 1 (empty) in period 1
     waterInd(1) = 1;
 
-    % Index for water that is used for irrigation
+    % Amount of water (as index) over time to be used for irrigation
     irrigationInd = zeros(1, T);
     
-    % Steady state levels
+    % Steady state levels of water in the reservoir over time
     steadyStateLvls = zeros(2, T);
     % Initialize the steady state period with the amount of
     % periods of the forward iteration in case no steady state will be
@@ -223,31 +223,44 @@ for monteInd=1:monteCarloMaxIter
     
     % Perform the forward iteration
     for i=1:T
+        % Irrigate with the optimal amount of water concerning the current
+        % water level
         irrigationInd(i) = optIrrigation_ind(waterInd(i));
-
-        waterInd(i+1) = min(max(waterInd(i) - irrigationInd(i) + round(r(i)/(M/dimWL)) - round(e(i)/(M/dimWL)),1),dimWL);
         
-        if (i>steadyStateMeanPeriods)
-            steadyStateLvls(1,i) = mean(waterLevel(waterInd(i-steadyStateMeanPeriods:i)));
+        % Calculate the water level in the following period
+        waterInd(i+1) = min(max(waterInd(i) ...
[…]

### 2020-02-11 Paper/paper.tex
some more details on steady states
fix typo in textit

@@
 Apart from this example, it's not always trivial to calculate the antiderivative of a function analytically. 
-However it is possible compute the area below (almost) any function using numerical integration, also called numerical quadrature.
+[TODO: zwei mal however]However it is possible compute the area below (almost) any function using numerical integration, also called numerical quadrature.
 
@@
 In order to do so, we need to calculate an auxiliary matrix $aux$ that contains the combined utilities of the farmers and the recreational users. 
-Obviously, the $aux$ matrix can only a finite, discrete amount of values.
-In this work, a resolution of $n = 1000$ different water levels and irrigation amounts should suffice.
+Obviously, the $aux$ matrix can only a finite, discrete amount of values. In this work, a resolution of $n = 1000$ different water levels and irrigation amounts should suffice. 
 Thus we discretize both water level $s$ and irrigation amount $x$ as vectors
@@
 which equals to finding the column $aux$ that maximizes the expected value over all states (TODO: kann man das so sagen?).
-We repeat this step over and over again until the our value function $V$ converges (technically: the norm of the difference between $V'$ and $V$ falls below a certain tolerance value).
+We repeat this step over and over again until the our value function $V$ converges (technically: the norm of the difference between $V'$ and $V$ falls below a certain tolerance value). \\
 
-Finally, we want to compute the value for the steady state.
+Finally, we want to compute the value of the steady state by using a monte carlo simulation.
 
@@
 	We then counted the occurrences, normalized them and calculated the corresponding probability.
-	The result is put in a histogram and shown in figure \ref{fig:steadyStateMonteCarlo}
+	The result is put in a histogram and shown in figure \ref{fig:steadyStateMonteCarlo}.
+	The highest possibility for a steady state is around 4.2. 
+	In edge cases one can find steady states as low as 3.5 or as high as 5.2. 
+	The curve resembles a normal distribution. 
 	\begin{figure}[ht]

### 2020-02-11 Paper/paper.tex
fix typo in textit

@@
 In his paper "The Theory Of Dynamic Programming" from 1957, Richard Bellman aimed to provide a solution to multi-stage-decision processes. 
-He defines a \texit{policy} as "a sequence of decisions" (\cite{bellman:1957}, p.2), and an \textit{optimal policy} as "a policy, which is most advantageous according to some preassigned criterion" (\cite{bellman:1957}, p.2).
+He defines a \textit{policy} as "a sequence of decisions" (\cite{bellman:1957}, p.2), and an \textit{optimal policy} as "a policy, which is most advantageous according to some preassigned criterion" (\cite{bellman:1957}, p.2).
 This is reflected in Bellman's Principle of Optimality: "An optimal policy has the property that whatever initial state and initial decisions are, the remaining decisions must constitute an optimal policy with regard o the state resulting from the first decisions" (\cite{bellman:1957}, p.2).

### 2020-02-11 Paper/paper.tex
typos

@@
 			stud. 2nd semester M.Sc. Business Administration and Engineering: Materials- and Process Engineering\\
-			Matr. Nr. 332404
+			Matr. Nr. 332404 \\
 			\href{mailto:niklas.sayer@rwth-aachen.de}{niklas.sayer@rwth-aachen.de}
@@
 	\newline
-	Every value along the possible water level leads to a irrigation amount that optimizes the value for all users in the given situation.
+	Every value along the possible water levels leads to an irrigation amount that optimizes the value for all parties in the given situation.
 Summarizing these values over the range of possible water levels results in a optimal irrigation policy which is shown in figure \ref{fig:optimal-irrigation-policy}.
@@
 	We then counted the occurrences, normalized them and calculated the corresponding probability.
-	The result is put in a histogram and shown in figure \ref{fig:steadyStateMonteCarlo}.
-	The highest possibility for a steady state is around 4.2. 
-	In edge cases one can find steady states as low as 3.5 or as high as 5.2. 
-	The curve resembles a normal distribution. 
+	The result is put in a histogram and shown in figure \ref{fig:steadyStateMonteCarlo}
 	\begin{figure}[ht]

### 2020-02-11 Code/code.m
commenting

@@ -221,7 +221,7 @@ for monteInd=1:monteCarloMaxIter
     % found
     steadyStatePeriod = T;
     
-    % Perform the forward iteration
+    % Perform a forward iteration
     for i=1:T
         % Irrigate with the optimal amount of water concerning the current
         % water level
@@ -248,8 +248,10 @@ for monteInd=1:monteCarloMaxIter
         end
     end
     
-    
+    % Try to find a steady state
     for i=2:T
+        % A steady state has been found if the changes in the steady state
+        % levels are below the steady state tolerance
         if (abs(steadyStateLvls(2,i)) < steadyStateTolerance)
             steadyStatePeriod = i;
             steadyStateLvl = steadyStateLvls(1,i);
@@ -257,7 +259,8 @@ for monteInd=1:monteCarloMaxIter
             break;
         end
     end
-
+    
+    % Save the steady state of the current monte carlo iteration
     monteCarloSteadyState(monteInd) = steadyStateLvl;
     fprintf('Iteration %s ended with Steady State %s\n', ...
              num2str(monteInd), num2str(steadyStateLvl));

### 2020-02-11 Paper/paper.tex
small fix
minor changes

@@
 
-In this work we want to examine the variations in use of water from a (hypothetical) water reservoir.
+In this work, we want to examine the variations in use of water from a (hypothetical) water reservoir.
 One of the most important uses of surface water is the irrigation of fields for food production.
@@
 \begin{dmath}
-aux &= \left[
- \begin{array}{ccccc} 
+aux = \left[\begin{array}{ccccc} 
 u(s_1, x_1)& \emptyset & \emptyset & ...
@@
 	% TODO maybe move to theory?
-	In systems theory, a system or a process is in a steady state if the variables which define the behavior of the system are unchanging in time \cite{gagniuc2017markov}. 
+	In systems theory, a system or a process is in a steady state, if the variables which define the behavior of the system are unchanging in time \cite{gagniuc2017markov}. 
 	The variable defining the behavior of our system is the water level. 
 	Because of this, we have to calculate the difference in water level for each period. 
-	Due to the fact, that our water level highly depends on the randomly distributed rainfalls that rise up the water level at the end of each period, it is not possible to find a state steady. 
+	Due to the fact, that our water level highly depends on the randomly distributed rainfalls that fill the reservoir at end of each period, it is not possible to find a state steady.
+% TODO: auch noch evaporation nennen?
 	To accomplish this, we defined our steady state, to be a state, where the mean water level of the last 200 periods did not change more than $1x10^{-7}$ from a period to the next.

### 2020-02-11 Code/code.m
renaming variables, code formating to match with those super short suggested line breaks

@@ -128,29 +128,37 @@ if (useRealData)
 
     % Calculate the evaporation with the formula given in the literature
     % (average evaporation in mm per year)
-    e = ((700 .* (avgMeanTemp + 0.006 .* altitude) ./...
+    avgEvap = ((700 .* (avgMeanTemp + 0.006 .* altitude) ./...
     (100 - latitude) + 15 .* ((100 - avgHumidity) / 5))...
     ./ (80 - avgMeanTemp)) * 365.25; 
     
     % Exclude missing values
-    e = e(~isnan(e));
+    avgEvap = avgEvap(~isnan(avgEvap));
 
     % Scale the real data down to be sensible in this example
-    e = e / 900;
+    avgEvap = avgEvap / 900;
 
     % Expected value of the evaporation
-    E_Evap = mean(e);
+    E_Evap = mean(avgEvap);
 
     % Discretize and round the expected value of the evaporation to fit to
     % the grid
     dimE_Evap = round(E_Evap/(M/dimWL));
 else 
     % Use no evaporation
-    e = zeros(T,1);
+    avgEvap = zeros(T,1);
     dimE_Evap = 0;
 end
 
-[V, optIrrigation_ind, aux_farm, aux_rec] = ValueFunction(dimWL, valueFunctionMaxIter, waterLevel, dimE_Rain, dimE_Evap, utilFar, utilRec, beta, valueFunctionTolerance);
+[V, optIrrigation_ind, aux_farm, aux_rec] = ValueFunction(dimWL, ...
+                                                valueFunctionMaxIter, ...
+                                                waterLevel, ...
+                                                dimE_Rain, ...
+                                                dimE_Evap, ...
+                                                utilFar, ...
+                                                utilRec, ...
+                                                beta, ...
+                                                valueFunctionTolerance);
 
 
 % Do not print all the diagrams if running monte carlo simulation
@@ -231,7 +239,7 @@ for monteInd=1:monteCarloMaxIter
         waterInd(i+1) = min(max(waterInd(i) ...
                             - irrigationInd(i) ...
                             + round(r(i)/(M/dimWL)) ...
-                            - round(e(i)/(M/dimWL)),1),dimWL);
+                            - round(avgEvap(i)/(M/dimWL)),1),dimWL);
         
         % Calculate the steady state level of the water reservoir
         if (i > steadyStateMeanPeriods)
@@ -307,7 +315,7 @@ else
[…]

### 2020-02-11 Code/code.m
fix patch, fix steady state when not using monte carlo simulation

@@ -257,17 +257,20 @@ for monteInd=1:monteCarloMaxIter
     end
     
     % Try to find a steady state
-    for i=2:T
-        % A steady state has been found if the changes in the steady state
-        % levels are below the steady state tolerance
-        if (abs(steadyStateLvls(2,i)) < steadyStateTolerance)
-            steadyStatePeriod = i;
-            steadyStateLvl = steadyStateLvls(1,i);
-            fprintf('Steady State found in period %s\n', num2str(i));
-            break;
+    if(~useMonteCarloSimulation)
+        steadyStateLvl = steadyStateLvls(1,i);
+    else
+        for i=2:T
+            % A steady state has been found if the changes in the steady state
+            % levels are below the steady state tolerance
+            if (abs(steadyStateLvls(2,i)) < steadyStateTolerance)
+                steadyStatePeriod = i;
+                steadyStateLvl = steadyStateLvls(1,i);
+                fprintf('Steady State found in period %s\n', num2str(i));
+                break;
+            end
         end
     end
-    
     % Save the steady state of the current monte carlo iteration
     monteCarloSteadyState(monteInd) = steadyStateLvl;
     fprintf('Iteration %s ended with Steady State %s\n', ...
@@ -292,9 +295,7 @@ else
     plot(steadyStateLvls(1,:));
     plot(steadyStateLvls(2,:));
     plot([1 T],[steadyStateLvl steadyStateLvl],'--g');
-    area = patch([steadyStatePeriod ...
-        -steadyStateMeanPeriods steadyStatePeriod steadyStatePeriod ... 
-        steadyStatePeriod-steadyStateMeanPeriods],[0 0 7 7],'r');
+    area = patch([steadyStatePeriod-steadyStateMeanPeriods steadyStatePeriod steadyStatePeriod steadyStatePeriod-steadyStateMeanPeriods],[0 0 7 7],'r');
     alpha(area,.2)
     xlim([1 T]);
     ylim([-1 11]);

### 2020-02-11 Paper/paper.tex
rename subsection
remove blindtext
section regarding the effect of discount factors

@@
 
-\subsection{Implementation} % TODO: Better title?
+\subsection{Computation of the value function}
 
@@
 
-Finally, we want to compute the value of the steady state by using a monte carlo simulation.
-
-	\subsection{bla}
-	\begin{figure}[ht] % TODO remove example
-		\includegraphics[width=1\textwidth]{figures/example.png}
-		\caption{Example Image}
-		\label{fig:example-pic}
-	\end{figure}
-	\blindtext
-	\begin{figure}[ht] % TODO remove example
-		\begin{longtable}{|p{2,5cm}|p{2,5cm}|p{2,5cm}|p{3cm}|p{3cm}|}
-			\hline
-			&\multicolumn{4}{c|}{\textbf{Klasse}} \\
-			\hline
-			& \textbf{S0} & \textbf{S1} & \textbf{S2} & \textbf{S3} \\
-			\hline
-			\textbf{lorem} & ipsum & ipsum & ipsum & ipsum \\
-			\hline
-		\end{longtable}
-		\label{fig:example-figure}
-		\caption{Example Table}
-	\end{figure}
-	\blindtext
-	\blindtext
-	\blindtext
+% Finally, we want to compute the value of the steady state by using a monte carlo simulation. 
+% TODO: hier steady state und monte carlo simulation erklären oder besser in results?
 
@@
 	We then counted the occurrences, normalized them and calculated the corresponding probability.
-	The result is put in a histogram and shown in figure \ref{fig:steadyStateMonteCarlo}
+	The result is put in a histogram and shown in figure \ref{fig:steadyStateMonteCarlo}.
+	The highest possibility for a steady state is around 4.2. 
+	In edge cases one can find steady states as low as 3.5 or as high as 5.2. 
+	The curve resembles a normal distribution.
+
 	\begin{figure}[ht]
@@
 	\subsection{Effect of Changed Discount Factors to Optimal Irrigation Policy}
+	When altering the discount factor $\beta$ in a range from 0 to 1, the optimal irrigation policy changes.
+	Looking at small water levels within the reservoir the optimal irrigation amount does not change much depending on the discount factor.
+	A small $\beta$ ensures that high water levels at the end of one period are only available to a small extent in the next period. 
+	This leads to a increasing gap between irrigation amounts over $\beta$ with increasing water level since conserving water does not increase the usage as much.
+	\begin{figure}[ht]
+		\includegraphics[width=1\textwidth]{figures/optimum-policy-over-beta.eps}
+		\caption{Optimal irrigation policy as a function of the discount factor $\beta$}
+		\label{fig:optimal-irrigation-beta}
+	\end{figure}
 	\subsection{Changed Volatility of Rainfalls}

### 2020-02-11 Code/code.m
disable realdata by default and calculate effect of betas in between 0.1 and 0.95

@@ -13,7 +13,7 @@ useMonteCarloSimulation = false;
 monteCarloMaxIter = 500;
 
 % Switch to use real rainfall and evaporation data 
-useRealData = true;
+useRealData = false;
 
 % Set the file name of the weather dataset
 realDataSource = 'bangladesh_weather_formatted.csv';
@@ -364,7 +364,7 @@ else
     % Plot optimal irrigation policy
     figure(9)
     hold on
-    for beta=1:-0.05:0.1
+    for beta=0.95:-0.05:0.1
         % Computation of the Value Function
         [V, optIrrigation_ind, aux_farm, aux_rec] = ValueFunction(dimWL, valueFunctionMaxIter, waterLevel, dimE_Rain, dimE_Evap, utilFar, utilRec, beta, valueFunctionTolerance);
         plot(linspace(0,M,dimWL), waterLevel(optIrrigation_ind), 'color', colormap_jet(round(256*beta),:))

### 2020-02-11 Code/code.m
only run the beta analysis once to save calculation speed when not needed

@@ -364,7 +364,7 @@ else
     % Plot optimal irrigation policy
     figure(9)
     hold on
-    for beta=0.95:-0.05:0.1
+    for beta=0.95:-0.05:0.95
         % Computation of the Value Function
         [V, optIrrigation_ind, aux_farm, aux_rec] = ValueFunction(dimWL, valueFunctionMaxIter, waterLevel, dimE_Rain, dimE_Evap, utilFar, utilRec, beta, valueFunctionTolerance);
         plot(linspace(0,M,dimWL), waterLevel(optIrrigation_ind), 'color', colormap_jet(round(256*beta),:))

### 2020-02-11 Paper/paper.tex
update graph
add new pages after logical sections to keep the images together. We might be able to remove this hack again after we finished the texts
rainfall with higher standard deviation

@@
 \usepackage{graphicx} % for figures
+\usepackage{subcaption}
 
@@
 	\end{figure}
+	\newpage
 	\subsection{Steady State}
@@
 	This leads to a increasing gap between irrigation amounts over $\beta$ with increasing water level since conserving water does not increase the usage as much.
+	One can say that a high $\beta$ prefers a long-term policy while a smaller $\beta$ leads to a rather short-term policy.
 	\begin{figure}[ht]
@@
 	\end{figure}
+	\newpage
 	\subsection{Changed Volatility of Rainfalls}
-
+	A more volatile rainfall means a rainfall having a higher standard deviation.
+	Since we want to compare two rainfall distributions of the same mean, the median of the distribution has to be changed.
+	Using a $\sigma^2=2$ includes a mean of $\mu=-1.5$ in order for the mean to stay the same. 
+	(See \ref{eq:rainfall} for reference) 
+	\begin{equation}
+		mean = exp(\mu+\frac{\sigma^2}{2}) \label{eq:rainfall}
+	\end{equation} 
+	Both distributions are visualized in \ref{fig:rain-distributions}.
+	\begin{figure}
+		\begin{subfigure}{0.5\textwidth}
+			\centering
+			\includegraphics[width=1\textwidth]{figures/rainfall-default.eps}
+			\caption{$\mu=0, \sigma=1$}
+			\label{fig:rainfall-default}
+		\end{subfigure}%
+		\begin{subfigure}{.5\textwidth}
+			\centering
+			\includegraphics[width=1\textwidth]{figures/rainfall-more-volatile.eps}
+			\caption{$\mu=-1.5, \sigma=2$}
+			\label{fig:rainfall-changed}
+		\end{subfigure}
+		\caption{Different Rain Distributions}
+		\label{fig:rain-distributions}
+	\end{figure}
+	\newpage
 	\printbibliography

### 2020-02-12 Paper/paper.tex
more results on rainfall distributions

@@
 	\newpage
+	\begin{figure}[ht]
+		\includegraphics[width=1\textwidth]{figures/aggregated-use-rainfall-default.eps}
+		\caption{Aggregated Use for lognormal distributed rainfall $(\mu=0, \sigma=1)$}
+		\label{fig:aggregated-use-rainfall-default}
+	\end{figure}
+	\begin{figure}[ht]
+	\includegraphics[width=1\textwidth]{figures/aggregated-use-rainfall-more-volatile.eps}
+	\caption{Aggregated Use for lognormal distributed rainfall $(\mu=-1.5, \sigma=2)$}
+	\label{fig:aggregated-use-rainfall-more-volatile}
+	\end{figure}
+	\ref{fig:aggregated-use-rainfall-default} shows the aggregated use of a lognormal distributed rainfall including the water level, the amount of water used for irrigation in the rain exemplary for 100 periods.
+\ref{fig:aggregated-use-rainfall-more-volatile} does the same for a changed lognormal distribution parameterized with $\mu=-1.5$ and $\sigma=2$.
+	When comparing both visualizations, it is noticeable that more volatile rainfall leads to much higher peaks of rain in combination with reduced number of periods with a decent amount of rain. 
+	As a consequence the water level in the reservoir fluctuates much more leading to periods with almost no water in the reservoir. 
+	In such periods the farmers use almost no water for irrigation.
+	The aggregated use of farmers and recreational users drops significantly.
+	Since the reservoir is capped at a amount of 7, the high peaks of rainfall can not be used to a large extent to fill up the reservoir.
+	Generally speaking, one can say that a more volatile rain distribution leads to a lower average water level in the reservoir with lower amounts of water to be used for irrigation.
+	This effect is reflected in the aggregated use as well. 
+	The Use of more volatile rainfall is much lower compared to rainfall parameterized with $\mu=0$ and $\sigma=1$.
+	\clearpage	
+	\newpage
 	\printbibliography

### 2020-02-12 Code/code.m
figures for rainfall analysis

@@ -209,6 +209,16 @@ for monteInd=1:monteCarloMaxIter
         r = exp(mu + sigma.*randn(T,1));
     end
     
+    figure(11)
+    set(gcf,'Units','Centimeters','position',[0,0,8,8]);
+    edges = [0:0.05:7];
+    hold on
+    histogram(r, edges, 'Normalization', 'Probability')
+    title('Probability of Rainfall');
+    ylabel('Probability');
+    xlabel('Amount of Rain');
+    hold off
+    
     % Overwrite T to match available data points if using real data
     if (useRealData)
         T = size(r,1);
@@ -372,6 +382,28 @@ else
     title('Optimal Irrigation Policy')
     xlabel('Water Level of the Reservoir')
     ylabel('Irrigation Amount')
+    hold off
+    
+    % Plot the aggregated use for farmers and recreational users
+    figure(10)
+    hold on
+    set(gca,'FontSize',12)
+    set(gcf,'Units','Centimeters','position',[0,0,16,8]);
+    %title('Aggreated Use for Farmers and Recreational Users');
+    yyaxis left;
+    plot(utilRec(waterLevel(waterInd(:, 1:1000)), waterLevel(irrigationInd)) + utilFar(waterLevel(irrigationInd)),'color','#A2142F');
+    ylabel('Use');
+    ylim([-100 0]);
+    yyaxis right;
+    plot(waterLevel(waterInd),'-', 'color', '#0072BD');
+    plot(waterLevel(irrigationInd),'-','color','#D95319');
+    plot(r,'-','color','#EDB120');
+    ylabel('Amount');
+    ylim([0 10]);
+    legend('Aggregated Use','Water Level of the Reservoir','Water used for Irrigation',...
+    'Rain');
+    xlim([1 100]);
+    xlabel('Period');
     hold off
 end
 

### 2020-02-12 Paper/paper.tex
fix indentation
more about using real data

@@
 \usepackage{subcaption}
+\usepackage{wrapfig}
 
@@
 We use a dataset aquired from \cite{kaggle:2019}, that contains real rainfall data for the last 65 years in different regions of Bangladesh. % Zitation anpassen
+% TODO describe how we fitted them to match the rainfall distribution
 
@@
 This effect is reflected in the aggregated use as well. 
-	The Use of more volatile rainfall is much lower compared to rainfall parameterized with $\mu=0$ and $\sigma=1$.
+The use of more volatile rainfall is much lower compared to rainfall parameterized with $\mu=0$ and $\sigma=1$.
+% TODO changed irrigation policy?
+\clearpage
+\subsection{Using Real Data for Rainfall}
+\ref{fig:rainfall-real-data} shows the distribution of rain from a real data set. 
+Comparing this figure with \ref{fig:rainfall-default} shows a similar shape of the distribution. 
+Since there are only 65 periods within the real data set is not as well distributed as in the lognormal distributed simulation.
+\begin{wrapfigure}{l}{0.5\textwidth} 
+	\includegraphics[width=0.5\textwidth]{figures/rainfall-real-data.eps}
+	\caption{Rainfall Distribution from Real Data}
+	\label{fig:rainfall-real-data}
+\end{wrapfigure}
+In the real data set there are no periods with (almost) no rain while in our simulation there are.
+As we can learn from our dataset this assumption is not realistic.
+Since the amount of rain is scaled down to match our model, there is no comparison between means of the model possible.
+Nevertheless one can say that the overall shape of a lognormal distribution is a feasible estimation for rainfall but for estimating whole years as periods the distribution might to be parameterized differently to fit actual rainfalls.
+As shown in the changed distribution of rainfalls also has an effect on the optimal irrigation policy since it depends on the expected amount of rain.
+\begin{wrapfigure}{l}{0.5\textwidth} 
+	\includegraphics[width=0.5\textwidth]{figures/optimalIrrigationPolicy-real-data.eps}
+	\caption{Optimal Irrigation Policy with Real Data}
+	\label{fig:optimalIrrigationPolicy-real-data}
+\end{wrapfigure}
+
 \clearpage	

### 2020-02-12 Paper/paper.tex
use cref for nice automatic refences

@@
 
+\usepackage[noabbrev]{cleveref}
+
 \usepackage{longtable}
@@
 \begin{document}
+	\let\origref\ref % The original \ref command is now accessible by \origref
+	\let\ref\cref    % The command \ref uses \cref from the cleveref package now
 
@@
 Note that while looking at just two agents is a vast simplification compared to a real allocation problem, one can easily add more stakeholders to the model, as long as there is enough computational power.
-The sequence of events for every period is displayed in figure ~\ref{fig:Sequence}.
+The sequence of events for every period is displayed in ~\ref{fig:Sequence}.
 At the beginning of every year (Step 1) we have to decide how much of the water from the reservoir is used for the irrigation of fields, which benefits the farmers' harvest.\\\\
@@
 contains the maximum possible usage that can be gained by irrigating a specific amount of water under a given water level. 
-Figure \ref{fig:value-function} shows the value function calculated for all possible water levels between the allowed boundaries of 0 and 7. 
+Figure \origref{fig:value-function} shows the value function calculated for all possible water levels between the allowed boundaries of 0 and 7. 
 \begin{figure}[ht]
@@
 Every value along the possible water levels leads to an irrigation amount that optimizes the value for all parties in the given situation.
-Summarizing these values over the range of possible water levels results in a optimal irrigation policy which is shown in figure \ref{fig:optimal-irrigation-policy}.
+Summarizing these values over the range of possible water levels results in a optimal irrigation policy which is shown in \ref{fig:optimal-irrigation-policy}.
 The plot shows a rising proposed irrigation amount with rising water level. There are 2 particularities that can be found within the graph:
@@
 To accomplish this, we defined our steady state, to be a state, where the mean water level of the last 200 periods did not change more than $1x10^{-7}$ from a period to the next.
-A exemplary graph of the mean in every period and the difference to its predecessor is shown in figure \ref{fig:steadyState}.
+A exemplary graph of the mean in every period and the difference to its predecessor is shown in \ref{fig:steadyState}.
 Once a period is found, where the mean does not change more than the given value, the mean of that period is defined as steady state water level.
@@
 We then counted the occurrences, normalized them and calculated the corresponding probability.
-The result is put in a histogram and shown in figure \ref{fig:steadyStateMonteCarlo}.
+The result is put in a histogram and shown in \ref{fig:steadyStateMonteCarlo}.
 The highest possibility for a steady state is around 4.2. 
@@
 \ref{fig:rainfall-real-data} shows the distribution of rain from a real data set. 
-Comparing this figure with \ref{fig:rainfall-default} shows a similar shape of the distribution. 
+Comparing this with \ref{fig:rainfall-default} shows a similar shape of the distribution. 
 Since there are only 65 periods within the real data set is not as well distributed as in the lognormal distributed simulation.
@@
 Nevertheless one can say that the overall shape of a lognormal distribution is a feasible estimation for rainfall but for estimating whole years as periods the distribution might to be parameterized differently to fit actual rainfalls.
-As shown in the changed distribution of rainfalls also has an effect on the optimal irrigation policy since it depends on the expected amount of rain.
+As shown in \ref{fig:optimalIrrigationPolicy-real-data} the changed distribution of rainfalls also has an effect on the optimal irrigation policy since it depends ocn the expected amount of rain.
 \begin{wrapfigure}{l}{0.5\textwidth} 

### 2020-02-12 Paper/paper.tex
further text and removed wrapfigure for now. Looks nicer side by side

@@
 Since there are only 65 periods within the real data set is not as well distributed as in the lognormal distributed simulation.
-\begin{wrapfigure}{l}{0.5\textwidth} 
-	\includegraphics[width=0.5\textwidth]{figures/rainfall-real-data.eps}
+\begin{figure}[h]
+\begin{subfigure}{0.5\textwidth}
+	\centering
+	\includegraphics[width=1\textwidth]{figures/rainfall-real-data.eps}
 	\caption{Rainfall Distribution from Real Data}
 	\label{fig:rainfall-real-data}
-\end{wrapfigure}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+	\centering
+	\includegraphics[width=1\textwidth]{figures/optimalIrrigationPolicy-real-data.eps}
+	\caption{Optimal Irrigation Policy with Real Data}
+	\label{fig:optimalIrrigationPolicy-real-data}
+\end{subfigure}
+\end{figure}
 In the real data set there are no periods with (almost) no rain while in our simulation there are.
@@
 Nevertheless one can say that the overall shape of a lognormal distribution is a feasible estimation for rainfall but for estimating whole years as periods the distribution might to be parameterized differently to fit actual rainfalls.
-As shown in \ref{fig:optimalIrrigationPolicy-real-data} the changed distribution of rainfalls also has an effect on the optimal irrigation policy since it depends ocn the expected amount of rain.
-\begin{wrapfigure}{l}{0.5\textwidth} 
-	\includegraphics[width=0.5\textwidth]{figures/optimalIrrigationPolicy-real-data.eps}
-	\caption{Optimal Irrigation Policy with Real Data}
-	\label{fig:optimalIrrigationPolicy-real-data}
-\end{wrapfigure}
+As shown in \ref{fig:optimalIrrigationPolicy-real-data} the changed distribution of rainfalls also has an effect on the optimal irrigation policy since it depends on the expected amount of rain.
+Due to the higher expected value of rainfalls the optimal irrigation policy is steeper over the water level compared to the simulated rainfall.
 

### 2020-02-18 Code/code.m
reformat line width to max. 120

@@ -150,15 +150,8 @@ else
     dimE_Evap = 0;
 end
 
-[V, optIrrigation_ind, aux_farm, aux_rec] = ValueFunction(dimWL, ...
-                                                valueFunctionMaxIter, ...
-                                                waterLevel, ...
-                                                dimE_Rain, ...
-                                                dimE_Evap, ...
-                                                utilFar, ...
-                                                utilRec, ...
-                                                beta, ...
-                                                valueFunctionTolerance);
+[V, optIrrigation_ind, aux_farm, aux_rec] = ValueFunction(dimWL, valueFunctionMaxIter, waterLevel, dimE_Rain, ...
+                                                dimE_Evap, utilFar, utilRec, beta, valueFunctionTolerance);
 
 
 % Do not print all the diagrams if running monte carlo simulation
@@ -246,15 +239,12 @@ for monteInd=1:monteCarloMaxIter
         irrigationInd(i) = optIrrigation_ind(waterInd(i));
         
         % Calculate the water level in the following period
-        waterInd(i+1) = min(max(waterInd(i) ...
-                            - irrigationInd(i) ...
-                            + round(r(i)/(M/dimWL)) ...
+        waterInd(i+1) = min(max(waterInd(i) - irrigationInd(i) + round(r(i)/(M/dimWL)) ...
                             - round(avgEvap(i)/(M/dimWL)),1),dimWL);
         
         % Calculate the steady state level of the water reservoir
         if (i > steadyStateMeanPeriods)
-            steadyStateLvls(1,i) = mean(waterLevel(waterInd(i...
-                                             - steadyStateMeanPeriods:i)));
+            steadyStateLvls(1,i) = mean(waterLevel(waterInd(i - steadyStateMeanPeriods:i)));
         else
             steadyStateLvls(1,i) = mean(waterLevel(waterInd(1:i)));
         end
@@ -362,8 +352,7 @@ else
     plot(steadyStateLvls(2,:));
     plot([1 T],[steadyStateLvl steadyStateLvl],'--g');
     xlim([1 T]);
-    legend('Water Level of the Reservoir',strcat('Mean last ', ...
-        num2str(steadyStateMeanPeriods), ' Periods'), ...
+    legend('Water Level of the Reservoir',strcat('Mean last ', num2str(steadyStateMeanPeriods), ' Periods'), ...
[…]

### 2020-02-18 Code/code.m
reorder plots

@@ -154,42 +154,6 @@ end
                                                 dimE_Evap, utilFar, utilRec, beta, valueFunctionTolerance);
 
 
-% Do not print all the diagrams if running monte carlo simulation
-% Note: figures 6 and 7 require manual panning and zooming in order to
-% see the details
-if(~useMonteCarloSimulation)
-    % Plot the auxiliary matrix of the farmers
-    figure(6)
-    hold on
-    surf(aux_farm);
-    colormap(jet);
-    title('Auxiliary Matrix of the Farmers');
-    xlabel('Amount of Water Used For Irrigation');
-    ylabel('Amount of Water in the Reservoir');
-    zlabel('Value');
-    hold off
-    
-    % Plot the auxiliary matrix of the recreational user
-    figure(7)
-    hold on
-    surf(aux_rec);
-    colormap(jet);
-    title('Auxiliary Matrix of the Recreational Users');
-    xlabel('Amount of Water Used For Irrigation');
-    ylabel('Amount of Water in the Reservoir');
-    zlabel('Value');
-    hold off
-
-    % Plot the value function
-    figure(1)
-    hold on
-    plot(waterLevel,V(:));
-    title('Value Function');
-    xlabel('Water Level');
-    ylabel('Maximum Value V');
-    axis([0 M -20 0]);
-    hold off
-end
 
 %% (Monte Carlo) Simulation for the Steady State
 
@@ -201,16 +165,6 @@ for monteInd=1:monteCarloMaxIter
         % Use lognormal distributed rainfall
         r = exp(mu + sigma.*randn(T,1));
     end
-    
-    figure(11)
-    set(gcf,'Units','Centimeters','position',[0,0,8,8]);
-    edges = [0:0.05:7];
-    hold on
-    histogram(r, edges, 'Normalization', 'Probability')
-    title('Probability of Rainfall');
-    ylabel('Probability');
-    xlabel('Amount of Rain');
-    hold off
     
     % Overwrite T to match available data points if using real data
     if (useRealData)
@@ -277,6 +231,9 @@ for monteInd=1:monteCarloMaxIter
              num2str(monteInd), num2str(steadyStateLvl));
 end
 
+
+%% plots
+
 if(useMonteCarloSimulation)
     edges = [0:0.1:7];
     % Plot monte carlo steady states
@@ -288,25 +245,14 @@ if(useMonteCarloSimulation)
     xlabel('Water Level in of the Reservoir');
     hold off
 else
[…]

### 2020-02-18 Code/code.m
reformat line width to max. 120

@@ -402,16 +402,8 @@ end
 
 
 %% Computation of the Value Function
-function [V, optIrrigation_ind, aux_farm, aux_rec] ...
-                                = ValueFunction(dimWL, ...
-                                                valueFunctionMaxIter, ...
-                                                waterLevel, ...
-                                                dimE_Rain, ...
-                                                dimE_Evap, ...
-                                                utilFar, ...
-                                                utilRec, ...
-                                                beta, ...
-                                                valueFunctionTolerance)
+function [V, optIrrigation_ind, aux_farm, aux_rec] = ValueFunction(dimWL, valueFunctionMaxIter, waterLevel, dimE_Rain, ...
+                                                dimE_Evap, utilFar, utilRec, beta, valueFunctionTolerance)
 
     % Create a value function for the combined value of farmers and
     % recreational users for each water level, respecting the discounted value

### 2020-02-18 Paper/paper.tex
minor changes

@@
 Note that while looking at just two agents is a vast simplification compared to a real allocation problem, one can easily add more stakeholders to the model, as long as there is enough computational power.
-The sequence of events for every period is displayed in ~\ref{fig:Sequence}.
+The sequence of events for every period is displayed in \ref{fig:Sequence}.
 At the beginning of every year (Step 1) we have to decide how much of the water from the reservoir is used for the irrigation of fields, which benefits the farmers' harvest.\\\\
@@
 with the additional paramter $s$ that indicates the water level at the beginning of the year.
-For recreational users, the preference parameters shell be $a_2 = 2$ and $b_2 = -3$.\\\\
+For recreational users, the preference parameters shall be $a_2 = 2$ and $b_2 = -3$.\\\\
 
@@
 
-where the weights $w_i$ are given as the solution to a linear system of equations and the nodes $x_i$ correspond to the zeros of the nth Hermite polynomial. (\cite{seminar:week6})
+where the weights $w_i$ are given as the solution to a linear system of equations and the nodes $x_i$ correspond to the zeros of the n-th Hermite polynomial. (\cite{seminar:week6})
 
@@
 In order to do so, we need to calculate an auxiliary matrix $aux$ that contains the combined utilities of the farmers and the recreational users. 
-Obviously, the $aux$ matrix can only a finite, discrete amount of values. In this work, a resolution of $n = 1000$ different water levels and irrigation amounts should suffice. 
+Obviously, the $aux$ matrix can only contain a finite, discrete amount of values. In this work, a resolution of $n = 1000$ different water levels and irrigation amounts should suffice. 
 Thus we discretize both water level $s$ and irrigation amount $x$ as vectors
@@
 
+% TODO das scheint mir nicht so ganz korrekt zu sein. Meint ihr die Dimension?
 Note that the maximum reservoir capacity is $7$ and $|s|=|x| = n = 1000$.
@@
 \end{equation}
+\label{eq:value-function}
 
 which equals to finding the column $aux$ that maximizes the expected value over all states (TODO: kann man das so sagen?).
-We repeat this step over and over again until the our value function $V$ converges (technically: the norm of the difference between $V'$ and $V$ falls below a certain tolerance value). \\
+We repeat this step over and over again until the our value function $V$ converges.
+In our case the function converges, if the norm of the difference between $V'$ and $V$ falls below a certain tolerance value) \\
 
@@
 When questioning what the optimal irrigation policy is, the first question to be asked is how the aggregated value function of both recreational users and farmers looks like.
-The value function as posed in % TODO put links to theory here
-contains the maximum possible usage that can be gained by irrigating a specific amount of water under a given water level. 
+The value function as posed in \ref{eq:value-function} contains the maximum possible usage that can be gained by irrigating a specific amount of water under a given water level. 
 Figure \origref{fig:value-function} shows the value function calculated for all possible water levels between the allowed boundaries of 0 and 7. 
@@
 \newline
-Every value along the possible water levels leads to an irrigation amount that optimizes the value for all parties in the given situation.
+Every value along the possible water levels is mapped to an irrigation amount that optimizes the value for all parties in the given situation.
 Summarizing these values over the range of possible water levels results in a optimal irrigation policy which is shown in \ref{fig:optimal-irrigation-policy}.
@@
 Because of this, we have to calculate the difference in water level for each period. 
-Due to the fact, that our water level highly depends on the randomly distributed rainfalls that fill the reservoir at end of each period, it is not possible to find a state steady.
+Due to the fact, that our water level highly depends on the randomly distributed rainfalls that fill the reservoir at end of each period, it is not possible to find a steady state in which the water level does not change at all.
 % TODO: auch noch evaporation nennen?
-To accomplish this, we defined our steady state, to be a state, where the mean water level of the last 200 periods did not change more than $1x10^{-7}$ from a period to the next.
+To analyze it nevertheless, we defined our steady state, to be a state, where the mean water level of the last 200 periods did not change more than $1x10^{-7}$ from a period to the next.
 A exemplary graph of the mean in every period and the difference to its predecessor is shown in \ref{fig:steadyState}.

### 2020-02-20 Paper/paper.tex
add subsection evaporation

@@
 \newpage
+
+\subsection{Evaporation}
+In the following, we want to assess the effect of evaporation on the optimal irrigation policy and the steady state. 
+As described in \ref{intro-mod4}, we derive the amount of evaporation over time from other weather related and geographical data. 
+The (calculated) evaporation has a low volatility over the years and has value of approximately 1 unit, as \ref{fig:without-vs-with-evaporation} shows. 
+Due to the fact that the reservoir has a fixed capacity of 7 units, rainfall that would fill the reservoir above this value is not available for further use. 
+When taking evaporation into account, the reservoir never reaches its capacity limits and thus all rainfall will be saved. This leads to an increased volatility of the water level. 
+\\
+
+\begin{figure}[h]
+\begin{subfigure}{0.5\textwidth}
+	\centering
+	\includegraphics[width=1\textwidth]{figures/results-opt-irrigation-over-time-evap-inactive.eps}
+	\caption{Without Evaporation}
+	\label{fig:rainfall-real-data}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+	\centering
+	\includegraphics[width=1\textwidth]{figures/results-opt-irrigation-over-time-evap-active.eps}
+	\caption{With Evaporation}
+	\label{fig:value-function}
+\end{subfigure}
+\caption{Comparison of water level, irrigation amount, rainfall and evaporation over time using real data}
+\label{fig:without-vs-with-evaporation}
+\end{figure}
+
+As expected, the average water level is lower with evaporation than without, which in return leads to a lower irrigation the farmers are allowed to receive. \\
+We should not at this point, that under the assumptions made, this simulation does not allow for evaporation effects that outweigh the rainfalls, as this would dry out the reservoir.
+The reason for this limit is found in the utility functions of the farmers and the recreational users.
+Because of $x\leq s$, we get for $s\to 0 \Rightarrow x\to 0$, and as a consequence
+
+\begin{equation}
+\lim_{s, x\to0}F(x) = \lim_{s,x\to0}-\frac{1}{x} = -\inf
+\end{equation}
+
+as well as
+
+\begin{equation}
+\lim_{s, x\to0}G(s,x) = \lim_{s,x\to0}-\frac{1}{(s-x)^2} = -\inf
+\end{equation}
+
+%TODO den folgenden Absatz rauswerfen?
+How then can these utility functions be interpreted in terms of reality? 
+An utility value of negative infinity may be reasonable, if an actor has no alternative actions available. 
+In this scenario, we could imagine this to be an isolated village, where the only possibility for the farmers to irrigate their fields is to use water from this particular reservoir, and the only way to have food for survive is to make use of their own fields (nothing else to do, no trading for example), assuming that the farmers want to survive.
+
 \printbibliography

### 2020-02-22 Code/code.m
change factors for realdata rain and evaporation

@@ -22,6 +22,11 @@ realDataSource = 'bangladesh_weather_formatted.csv';
 % simulate the evaporation
 locationName = 'Barisal';
 
+% Reservoir surface to volume ratio
+% This exemplary value is calculated for the Rurstausee with data from
+% https://de.wikipedia.org/wiki/Rurtalsperre
+surface_to_volume_ratio = 7.83 / 202.6;
+
 % Parameters of the value function
 a1 = 1;
 a2 = 2;
@@ -84,16 +89,23 @@ waterLevel = linspace(0,M,dimWL);
 if (useRealData)
     % Find column indices of the necessary data
     indData = find(strcmp(table2cell(data(1,:)), locationName));
-
     % Extract the rainfall data and calculate the average per year
-    avgRain = str2double(table2array(data(5:end,indData(1,4)))) * 12;
-
-    % Exclude missing values
-    avgRain = avgRain(~isnan(avgRain));
-
-    % Scale the real data down to be sensible in this example
-    avgRain = avgRain / 900;
+    avgRain_mm = str2double(table2array(data(5:end,indData(1,4)))) * 12;
     
+    % Exclude missing values
+    avgRain_mm = avgRain_mm(~isnan(avgRain_mm));
+
+    % Norm the amount of rain in mm to the volume of the reservoir
+    avgRain = M*surface_to_volume_ratio*avgRain_mm/1000;
+
+    % Note that because of the weather conditions of the location chosen in 
+    % bangladesh, the amount of rainfall is not high enough to provide a
+    % significant amount of value through irrigatio for the farmers. For 
+    % this demo however, we want the reservoir to be refilled in order to 
+    % give the farmers at least some utility. Thus we will
+    % artificially increase the amount of rainfall here. 
+    avgRain = avgRain * 4;
+
     % Expected value of the rainfall
     E_rain = mean(avgRain);
 else
@@ -120,23 +132,38 @@ dimE_Rain = round(E_rain/(M/dimWL));
 if (useRealData)
     % Extract the data necessary for the evaporation simulation
     altitude = str2double(table2array(data(2,indData(1,1))));
-    latitude = str2double(table2array(data(3,indData(1,1))));
+    latitude_n = str2double(table2array(data(3,indData(1,1))));
     avgMaxTemp = str2double(table2array(data(5:end,indData(1,1))));
[…]

### 2020-02-22 Paper/paper.tex
todo note
fix label

@@
 
-\subsubsection{Modification 4: Taking Evaporation into Account}
+\subsubsection{Modification 4: Taking Evaporation into Account} \label{sec:intro-mod4}
 The evaporation of water from a water reservoir is the most important natural "outflow" to be considered when it comes to managing water levels \citep{tanny2008evaporation}.
@@
 \subsection{Using Real Data for Rainfall}
-\ref{fig:rainfall-real-data} shows the distribution of rain from a real data set. 
+\ref{fig:rainfall-real-data} shows the distribution of rain from a real data set. %TODO "figure" -> "Figure" groß schreiben!
 Comparing this with \ref{fig:rainfall-default} shows a similar shape of the distribution. 
@@
 In the following, we want to assess the effect of evaporation on the optimal irrigation policy and the steady state. 
-As described in \ref{intro-mod4}, we derive the amount of evaporation over time from other weather related and geographical data. 
+As described in \ref{sec:intro-mod4}, we derive the amount of evaporation over time from other weather related and geographical data. 
 The (calculated) evaporation has a low volatility over the years and has value of approximately 1 unit, as \ref{fig:without-vs-with-evaporation} shows. 

### 2020-02-22 Code/code.m
small fix

@@ -400,7 +400,7 @@ else
     set(gcf,'Units','Centimeters','position',[0,0,16,8]);
     %title('Aggreated Use for Farmers and Recreational Users');
     yyaxis left;
-    plot(utilRec(waterLevel(waterInd(:, 1:1000)), waterLevel(irrigationInd)) + utilFar(waterLevel(irrigationInd)),'color','#A2142F');
+    plot(utilRec(waterLevel(waterInd(:, 1:T)), waterLevel(irrigationInd)) + utilFar(waterLevel(irrigationInd)),'color','#A2142F');
     ylabel('Use');
     ylim([-100 0]);
     yyaxis right;

### 2020-02-22 Paper/paper.tex
write figur in capital
add todo
fix todos

@@
 \addbibresource{collection.bib}
-
-% TODO delete the lorem ipsum
-\usepackage{blindtext} % Package to generate dummy text throughout this template
 
@@
 Apart from this example, it's not always trivial to calculate the antiderivative of a function analytically. 
-[TODO: zwei mal however]However it is possible compute the area below (almost) any function using numerical integration, also called numerical quadrature.
+Nonetheless it is possible compute the area below (almost) any function using numerical integration, also called numerical quadrature.
 
@@
 
-where $u(s,x)$ is the utility to gain from the control variable (action) $x$ in state $s$ [TODO: QUOTE Seminar Week 3].
+where $u(s,x)$ is the utility to gain from the control variable (action) $x$ in state $s$ %TODO: QUOTE Seminar Week 3.
 
@@
 % Finally, we want to compute the value of the steady state by using a monte carlo simulation.
-% TODO: hier steady state und monte carlo simulation erklären oder besser in results?
 
@@
 \subsection{Steady State}
-% TODO maybe move to theory?
 In systems theory, a system or a process is in a steady state, if the variables which define the behavior of the system are unchanging in time \cite{gagniuc2017markov}. 
@@
 % TODO: auch noch evaporation nennen?
-To analyze it nevertheless, we defined our steady state, to be a state, where the mean water level of the last 200 periods did not change more than $1x10^{-7}$ from a period to the next.
+To analyze it nevertheless, we defined our steady state, to be a state, where the mean water level of the last 200 periods did not change more than $1*10^{-7}$ from a period to the next.
 A exemplary graph of the mean in every period and the difference to its predecessor is shown in \ref{fig:steadyState}.
@@
 \subsection{Using Real Data for Rainfall}
-\ref{fig:rainfall-real-data} shows the distribution of rain from a real data set. %TODO "figure" -> "Figure" groß schreiben!
+%TODO update figures
+Figure \origref{fig:rainfall-real-data} shows the distribution of rain from a real data set.
 Comparing this with \ref{fig:rainfall-default} shows a similar shape of the distribution. 
@@
 
+\section{Conclusion}
+%TODO content
 \printbibliography
@@
 %----------------------------------------------------------------------------------------
-
+\pagestyle{empty} 
 \begingroup
@@
 \begin{tabular}{@{}p{8cm}p{5.8cm}}
+	\underline{\hspace{6cm}} & \underline{\hspace{5.8cm}} \\
+	\vspace{0.02cm}Name, Vorname & \vspace{0.02cm}Matrikelnummer \\
+	&\\
+	\underline{\hspace{6cm}} & \underline{\hspace{5.8cm}} \\
+	\vspace{0.02cm}Name, Vorname & \vspace{0.02cm}Matrikelnummer \\
+	&\\
 	\underline{\hspace{6cm}} & \underline{\hspace{5.8cm}} \\
@@
 \begin{tabular}{@{}p{8cm}p{5.8cm}}
-	\underline{\smash{Aachen, den \today}} & \underline{\hspace{5.8cm}}\\
+	\underline{\smash{Aachen, den \hspace{4cm}}} & \underline{\hspace{5.8cm}}\\
+	Ort, Datum & Unterschrift \\
+	&\\
+	\underline{\smash{Aachen, den \hspace{4cm}}} & \underline{\hspace{5.8cm}}\\
+	Ort, Datum & Unterschrift \\
+	&\\
+	\underline{\smash{Aachen, den \hspace{4cm}}} & \underline{\hspace{5.8cm}}\\
 	Ort, Datum & Unterschrift \\
@@
 \vspace{0.6cm}
-
 \begin{tabular}{@{}p{8cm}p{5.8cm}}
-	\underline{\smash{Aachen, den \today}} & \underline{\hspace{5.8cm}}\\
+	\underline{\smash{Aachen, den \hspace{4cm}}} & \underline{\hspace{5.8cm}}\\
+	Ort, Datum & Unterschrift \\
+	&\\
+	\underline{\smash{Aachen, den \hspace{4cm}}} & \underline{\hspace{5.8cm}}\\
+	Ort, Datum & Unterschrift \\
+	&\\
+	\underline{\smash{Aachen, den \hspace{4cm}}} & \underline{\hspace{5.8cm}}\\
 	Ort, Datum & Unterschrift \\

### 2020-02-22 Paper/paper.tex
unify citation style

@@
 backend=bibtex,
-style=alphabetic,
+style=ieee,
 citestyle=authoryear,
@@
 Comparing the results then allows us to draw conclusions on the accuracy of the simulation for that region. 
-We use a dataset aquired from \cite{kaggle:2019}, that contains real rainfall data for the last 65 years in different regions of Bangladesh. % Zitation anpassen
+We use a dataset aquired from \citep{kaggle:2019}, that contains real rainfall data for the last 65 years in different regions of Bangladesh. % Zitation anpassen
 % TODO describe how we fitted them to match the rainfall distribution
@@
 
-where the weights $w_i$ are given as the solution to a linear system of equations and the nodes $x_i$ correspond to the zeros of the n-th Hermite polynomial. (\cite{seminar:week6})
+where the weights $w_i$ are given as the solution to a linear system of equations and the nodes $x_i$ correspond to the zeros of the n-th Hermite polynomial \citep{seminar:week6}.
 
@@
 
-where $u(s,x)$ is the utility to gain from the control variable (action) $x$ in state $s$ %TODO: QUOTE Seminar Week 3.
+where $u(s,x)$ is the utility to gain from the control variable (action) $x$ in state $s$ \citep{seminar:week3}.
 
@@
 \subsection{Steady State}
-In systems theory, a system or a process is in a steady state, if the variables which define the behavior of the system are unchanging in time \cite{gagniuc2017markov}. 
+In systems theory, a system or a process is in a steady state, if the variables which define the behavior of the system are unchanging in time \citep{gagniuc2017markov}. 
 The variable defining the behavior of our system is the water level. 

### 2020-02-22 Code/code.m
correct formatting of diagramms

@@ -311,6 +311,8 @@ else
     % Plot optimal irrigation policy
     figure(4)
     hold on
+    set(gca,'FontSize',12);
+    set(gcf,'Units','Centimeters','position',[0,0,8,8]);
     plot(linspace(0,M,dimWL), waterLevel(optIrrigation_ind))
     title('Optimal Irrigation Policy')
     xlabel('Water Level of the Reservoir')

### 2020-02-22 Paper/paper.tex
resolve TODOs
proper formatting for new sections

@@
 
-% TODO do we need a abstract?
 \usepackage{abstract} % Allows abstract customization
@@
 \titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles
+\newcommand{\sectionbreak}{\clearpage}
 
@@
 
-which equals to finding the column $aux$ that maximizes the expected value over all states (TODO: kann man das so sagen?).
+which equals to finding the column $aux$ that maximizes the expected value over all states.
 We repeat this step over and over again until the our value function $V$ converges.
@@
 Due to the fact, that our water level highly depends on the randomly distributed rainfalls that fill the reservoir at end of each period, it is not possible to find a steady state in which the water level does not change at all.
-% TODO: auch noch evaporation nennen?
 To analyze it nevertheless, we defined our steady state, to be a state, where the mean water level of the last 200 periods did not change more than $1*10^{-7}$ from a period to the next.
@@
 The use of more volatile rainfall is much lower compared to rainfall parameterized with $\mu=0$ and $\sigma=1$.
-% TODO changed irrigation policy?
+Since the mean stays constant within the analyzed comparison, the value function does not change.
 \clearpage
 \subsection{Using Real Data for Rainfall}
-%TODO update figures
 Figure \origref{fig:rainfall-real-data} shows the distribution of rain from a real data set.
@@
 %TODO content
+\clearpage
 \printbibliography

### 2020-02-22 Paper/paper.tex
Kleinere Anmerkungen

@@
 We use a dataset aquired from \citep{kaggle:2019}, that contains real rainfall data for the last 65 years in different regions of Bangladesh. % Zitation anpassen
+To make this data fit our somewhat arbitrarily selected range of 0 to 7 for water levels in the reservoir, we...
 % TODO describe how we fitted them to match the rainfall distribution
@@
 A common technique to solve optimization or combinatorial problems is called Dynamic Programming. 
-It is based on the idea of dividing problems into smaller subproblems, solving these recursively and caching intermediate results (\textit{memoization}). 
+It is based on the idea of dividing problems into smaller subproblems, solving these recursively and caching intermediate results (\textit{memoization}). % Nicht eher "memoRization"?
 In his paper "The Theory Of Dynamic Programming" from 1957, Richard Bellman aimed to provide a solution to multi-stage-decision processes. 
 He defines a \textit{policy} as "a sequence of decisions" (\cite{bellman:1957}, p.2), and an \textit{optimal policy} as "a policy, which is most advantageous according to some preassigned criterion" (\cite{bellman:1957}, p.2).
-This is reflected in Bellman's Principle of Optimality: "An optimal policy has the property that whatever initial state and initial decisions are, the remaining decisions must constitute an optimal policy with regard o the state resulting from the first decisions" (\cite{bellman:1957}, p.2).
+This is reflected in Bellman's Principle of Optimality: "An optimal policy has the property that whatever initial state and initial decisions are, the remaining decisions must constitute an optimal policy with regard to the state resulting from the first decisions" (\cite{bellman:1957}, p.2).
 
@@
 \end{equation}
-
+% TODO Gibt es hier nicht eine mathematisch schönere Schreibweise um einen Vektor mit 1000 Einträgen darzustellen? (R hoch 1000/ R hoch n)
 % TODO das scheint mir nicht so ganz korrekt zu sein. Meint ihr die Dimension?
@@
 We repeat this step over and over again until the our value function $V$ converges.
-In our case the function converges, if the norm of the difference between $V'$ and $V$ falls below a certain tolerance value) \\
+In our case the function converges, if the norm of the difference between $V'$ and $V$ falls below a certain tolerance value \\
 
@@
 Once a period is found, where the mean does not change more than the given value, the mean of that period is defined as steady state water level.
-The water level found to be a steady state and the period in which this condition occurs changes from calculation to calculation due to the fact, that the rain fall is still a random distribution.
+The water level found to be a steady state and the period in which this condition occurs changes from calculation to calculation due to the fact, that the rainfall is still a random distribution.
 \begin{figure}[ht]
@@
 In order to find out, how the steady states are distributed and what the probability of their occurrences are, we did a Monte Carlo simulation.
-To do so, we ran our code 1000 times and put the resulting water levels into bins with the size of 0.1 each.
+To do so, we ran our code 1000 times and put the resulting water levels into bins with the size of 0.1 each. % "the forward simulation" instead of "our code"?
 We then counted the occurrences, normalized them and calculated the corresponding probability.
@@
 A small $\beta$ ensures that high water levels at the end of one period are only available to a small extent in the next period.
+% Finde die Formulierung nicht so gut: Sie sind ja weiterhin voll verfügbar, aber eben in Zukunft weniger wert.
+Bzw das könnte man diskutieren.
 This leads to a increasing gap between irrigation amounts over $\beta$ with increasing water level since conserving water does not increase the usage as much.

### 2020-02-22 Paper/paper.tex
added evaporation factors

@@
 As expected, the average water level is lower with evaporation than without, which in return leads to a lower irrigation the farmers are allowed to receive. \\
-We should not at this point, that under the assumptions made, this simulation does not allow for evaporation effects that outweigh the rainfalls, as this would dry out the reservoir.
-The reason for this limit is found in the utility functions of the farmers and the recreational users.
-Because of $x\leq s$, we get for $s\to 0 \Rightarrow x\to 0$, and as a consequence
+In the introduction, we stated that due to climate change, prolonged dry periods will occur. We want to reflect this fact in the simulation by scaling the evaporation rate with different factors.
+
+%TODO make the figure better
+\begin{figure}[h]
+\begin{subfigure}{0.5\textwidth}
+	\centering
+	\includegraphics[width=1\textwidth]{figures/results-evap-factors-water-levels.eps}
+	\caption{Water Level}
+	\label{fig:rainfall-real-data}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+	\centering
+	\includegraphics[width=1\textwidth]{figures/results-evap-factors-utilities.eps}
+	\caption{Total Utility}
+	\label{fig:value-function}
+\end{subfigure}
+\caption{Water level and log-scaled total utility over time, scaling evaporation with factors $0, 1, 2, 2.3$ (from top to bottom)}
+\label{fig:evaporation-factors}
+\end{figure}
+
+\ref{fig:evaporation-factors} shows the total utility and the water level over time, using scaling factors of $0, 1, 2, 2.3$.\footnote{The expected value of the evaporation should not exceed the expected value of the rainfall, as this would make it impossible to calculate the value function.
+The expected value of the evaporation with a scaling factor of $2.3$ is just below the expected value of the rainfall.
+} %TODO Fußnote gut so?
+We can see that higher amounts of evaporation not only lead to water levels, but also to significantly lower utilities for the recreational users and the farmers. 
+Especially when the reservoir completely dries out, the total utility approaches $-\infty$. This happens in periods $7,62,63,64$ and $65$ when applying a factor of $2.3$ to the evaporation.
+
+%TODO improve this figure
+\begin{figure}[ht]
+	\includegraphics[width=1\textwidth]{figures/results-evap-factor-2.3.eps}
+	\caption{...}
+	\label{fig:evap-factor-2.3}
+\end{figure}
+
+The reason for this is found in the utility functions of the farmers and the recreational users. Because of $x\leq s$, we get for $s\to 0 \Rightarrow x\to 0$, and as a consequence
 

### 2020-02-22 Code/code.m
simulate different evaporation factors

@@ -25,7 +25,11 @@ locationName = 'Barisal';
 % Reservoir surface to volume ratio
 % This exemplary value is calculated for the Rurstausee with data from
 % https://de.wikipedia.org/wiki/Rurtalsperre
-surface_to_volume_ratio = 7.83 / 202.6;
+surfaceToVolumeRatio = 7.83 / 202.6;
+
+% Factors to artificially scale the amount of evaporation if real data
+% is being used
+evapFactors = [0, 1, 2, 2.3];
 
 % Parameters of the value function
 a1 = 1;
@@ -96,7 +100,7 @@ if (useRealData)
     avgRain_mm = avgRain_mm(~isnan(avgRain_mm));
 
     % Norm the amount of rain in mm to the volume of the reservoir
-    avgRain = M*surface_to_volume_ratio*avgRain_mm/1000;
+    avgRain = M*surfaceToVolumeRatio*avgRain_mm/1000;
 
     % Note that because of the weather conditions of the location chosen in 
     % bangladesh, the amount of rainfall is not high enough to provide a
@@ -126,140 +130,167 @@ end
 % the grid
 dimE_Rain = round(E_rain/(M/dimWL));
 
-
+utilityByEvapFactors = [];
+optIrrigationByEvapFactors = [];
+reservoirLevels = [];
+for evapFactor = evapFactors
+    
 %% Evaporation Simulation
-
-if (useRealData)
-    % Extract the data necessary for the evaporation simulation
-    altitude = str2double(table2array(data(2,indData(1,1))));
-    latitude_n = str2double(table2array(data(3,indData(1,1))));
-    avgMaxTemp = str2double(table2array(data(5:end,indData(1,1))));
-    avgMinTemp = str2double(table2array(data(5:end,indData(1,2))));
-    avgMeanTemp = (avgMaxTemp + avgMinTemp) / 2;
-    avgHumidity = str2double(table2array(data(5:end,indData(1,3))));
-
-    % Calculate the average dew temperature using (lawrence2005relationship)
-    % Inputs: avgHumidity in degrees celsius
-    %         avgHumidity isurface_to_volume_ration percent
-    % Output: avgDewTemp in degrees celsius
-    avgDewTemp = (100 - avgHumidity) / 5;
-
-    % Calculate the altitude in dergees south
-    latitude_s = -latitude_n;
-
-    % Calculate the average evaporation using (linacre1977simple)
-    % Inputs: avgMeanTemp in degrees celsius
[…]

### 2020-02-22 Paper/paper.tex
proper line breaks on the front page

@@
 			\textbf{Felix Fink} \\
-			stud. 2nd semester M.Sc. Business Administration and Engineering: Mechanical Engineering \\
+			stud. 2nd semester M.Sc. Business Administration and Engineering:\\Mechanical Engineering \\
 			Matr. Nr. 345971 \\
@@
 			\textbf{Niklas Sayer} \\
-			stud. 2nd semester M.Sc. Business Administration and Engineering: Materials- and Process Engineering\\
+			stud. 2nd semester M.Sc. Business Administration and Engineering:\\Materials- and Process Engineering\\
 			Matr. Nr. 332404 \\
@@
 			\textbf{Martin Wicke} \\
-			stud. 4th semester M.Sc. Business Administration and Engineering: Electrical Power Engineering\\
+			stud. 4th semester M.Sc. Business Administration and Engineering:\\Electrical Power Engineering\\
 			Matr. Nr. 336301 \\

### 2020-02-22 Code/code.m
small changes

@@ -272,25 +272,20 @@ end
 %% Plots
 close all;
 figure(100)
-colors = colormap(autumn);
 l = size(evapFactors, 2);
 for i=1:l
-    semilogy(utilityByEvapFactors(i, :), 'color', colors(round(0.5*255*i/l), :))
-    %stairs(utilityByEvapFactors(i, :), 'color', colors(round(0.5*255*i/l), :))
-    %set(gca, 'YScale', 'log')
-    hold on;
-end
-hold off;
-return
-figure(101)
-colors = colormap(winter);
-l = size(evapFactors, 2);
-for i=1:l
-    plot(reservoirLevels(i, :), 'color', colors(round(0.5*255*i/l), :))
+    semilogy(utilityByEvapFactors(i, :), 'color', [i/l, 0, 1-i/l])
     hold on;
 end
 hold off;
 
+figure(101)
+l = size(evapFactors, 2);
+for i=1:l
+    plot(reservoirLevels(i, :), 'color', [i/l, 0, 1-i/l])
+    hold on;
+end
+hold off;
 
 if(useMonteCarloSimulation)
     edges = [0:0.1:7];

### 2020-02-23 Paper/paper.tex
trying to get the figures on the right positions
Conclusion First Version

@@
 It expresses how much one unit of future utility is worth in terms of current utility and therefore makes them easily comparable \citep{kruschwitz2014investitionsrechnung}.
-We take a look at the discount factors x and x.
+We take a look at the discount factors from 0 to 1.
 
@@
 To consider the predicted impacts of climate change \citep{guhathakurta2011impact}, we look at the effect of increased rainfall volatility. 
-We do this by using a  $\sigma^2$ of x.
+We do this by using a  $\sigma^2$ of x. % TODO: Das ist noch nicht ganz richtig
 
@@
 
+% TODO Fehlen hier nicht noch ein paar Sätze dazu wie wir die optimal irrigation policy extrahieren?
+
 % Finally, we want to compute the value of the steady state by using a monte carlo simulation.
@@
 In order to find out, how the steady states are distributed and what the probability of their occurrences are, we did a Monte Carlo simulation.
-To do so, we ran our code 1000 times and put the resulting water levels into bins with the size of 0.1 each. % "the forward simulation" instead of "our code"?
+To do so, we ran our code 1000 times and put the resulting water levels into bins with the size of 0.1 each. % "the forward simulation" instead of "our code"? Außerdem laut code 500x statt 1000x
 We then counted the occurrences, normalized them and calculated the corresponding probability.
@@
 Since we want to compare two rainfall distributions of the same mean, the median of the distribution has to be changed.
-Using a $\sigma^2=2$ includes a mean of $\mu=-1.5$ in order for the mean to stay the same. 
+Using a $\sigma^2=2$ includes a mean of $\mu=-1.5$ in order for the mean to stay the same. % Würde das erste "mean of mu" rausnehmen oder spezifizieren -> mean of the underlying normal distribution
 (See \ref{eq:rainfall} for reference) 
@@
 Nevertheless one can say that the overall shape of a lognormal distribution is a feasible estimation for rainfall but for estimating whole years as periods the distribution might to be parameterized differently to fit actual rainfalls.
+% feasible -> sensible?
 As shown in \ref{fig:optimalIrrigationPolicy-real-data} the changed distribution of rainfalls also has an effect on the optimal irrigation policy since it depends on the expected amount of rain.
@@
 %TODO make the figure better
-\begin{figure}[h]
+\begin{figure}[ht]
 \begin{subfigure}{0.5\textwidth}
@@
 \end{figure}
+%\clearpage
 
@@
 %TODO improve this figure
-\begin{figure}[ht]
+\begin{figure}[h]
 	\includegraphics[width=1\textwidth]{figures/results-evap-factor-2.3.eps}
-	\caption{...}
+	\caption{Evaporation multiplied by a factor of $2.3$ outweighs the rainfall in certain periods} %TODO 
 	\label{fig:evap-factor-2.3}
@@
 %TODO content
+In this work we developed an optimal irrigation strategy for an hypothetical water reservoir, maximizing the sum of utilities for farmers and recreational users.
+Using dynamic programming we calculated an aggregated value function based on the given utility functions as described in chapter 2.4.
+We applied this value function to a basic model of the water reservoir, by extracting the optimal irrigation path from the value function and simulating water influx in the form of rainfall with a lognormally distributed random variable.
+The optimal irrigation policy stipulates quite low irrigation amounts for low water levels, mostly due to the very negative utility citizens receive from an almost empty reservoir (Figure 4).
+With $\beta = 0.9$, the optimal irrigation amount tapers off again at the end, due to % TODO: Besprechen!
+MMn wegen Möglichkeit des Mehrfachnutzens bei Non-irrigation (daher nicht erkennbar bei beta = 0)
+In figure 9 you can see one possible scenario in our basic setting with $\mu=0$ and $\sigma=1$.
+However, due to the random variable, every forward simulation can look quite different, even though no parameters are changed.
+Therefore, the steady state of the water level itself, defined as the moving average of 200 periods (see chapter 3.2), is uncertain.
+Running our forward simulation 500 times, we could see that it seems to follow a normal distribution with a mean of around 4.2 and a standard deviation of 0.37.
+% TODO: StdDev auch im Ergebnisteil erwähenen?
+\\\\
+We also looked at how changes in the discount factor $\beta$ influenced the optimal irrigation policy (Figure 7).
+As expected, a high $\beta$ leads to less irrigation at high water levels.
+That's because citizens might be able to benefit from the remaining water again in future periods, while when using water for irrigation farmers can only benefit once in the present period.
+Increasing $\beta$ means putting more weight on future utility up to a factor of $\beta = 1$, where stakeholders are indifferent between current and future utility.
+A low $\beta$ on the other hand leads to a higher amount of water used for irrigation, especially at higher water levels. 
+This can be easily imagined with the extreme case $\beta = 0$, which means that no one benefits from leaving water in the reservoir for the next period.
+Of course there is still water that's being left in the reservoir, but that is only to benefit citizens in the current period and not because of the benefits of the remaining water for the next period.
+In this case the model is reduced to a single-stage optimization problem with no uncertainty, where the only task is the optimal allocation of the given current amount of water on the basis of the stakeholders' utility functions.
+\\
+We then went on and looked at the effect of increased rainfall volatility (Figures 9 and 10).
+Because of the constant mean, the total amount of rain and therefore of water influx in the reservoir stays the same.
+The result is a lower average water level in the reservoir.
+The aggregated use is a good indicator of the efficiency our allocation strategy is achieving.
+We could see that with rainfalls getting more volatile the sum of aggregated use over all simulated periods drops significantly, due to efficiency losses caused by increased likelihood of reservoir overflow and drought.
+% Evlt noch schnell ausrechnen?
+This vividly illustrates a very real consequence of climate change that governments and counties might face in the future.
+\\
+Instead of simulated rainfall, we also used real weather data from Barisal, Bangladesh. 
+The real rainfall distribution (Figure 11 a) looked quite a bit different then our lognormal distribution with $\mu=0$ and $\sigma=1$.
+This indicates that other parameters will have to be found to better describe the annual rainfall in that specific area.
+The higher expected value of rain in the real distribution leads to the optimal irrigation policy showing a higher amount of water used for irrigation at higher water levels (Figure 11 b).
+Further adding evaporation to the model shows...
+%Evap
+% Efficiency loss with higher evap scale factor -> same as more volatile rainfalls
+
+% Ausblick
+% Whole different problem, more of empirical nature: What even is the "right" discount factor in this context
+
 \clearpage

### 2020-02-24 Paper/paper.tex
changes as discussed

@@
 So the value of the control variable to be set is the one that maximizes both the utility of the current state plus (recursively) the value $V$ of the next state $s'$ which is discounted by a factor of $\beta$.
+%TODO hier noch auf die Definition von beta eingehen? Was ist beta und warum ist das relevant?
 
@@
 \end{equation}
-% TODO Gibt es hier nicht eine mathematisch schönere Schreibweise um einen Vektor mit 1000 Einträgen darzustellen? (R hoch 1000/ R hoch n)
-% TODO das scheint mir nicht so ganz korrekt zu sein. Meint ihr die Dimension?
-Note that the maximum reservoir capacity is $7$ and $|s|=|x| = n = 1000$.
+Note that the maximum reservoir capacity is $7$ and $s, x \in R^{1000}$.
 
@@
 	\item Due to the effect that a almost empty reservoir has a very low use to recreational users, the optimal irrigation strategy suggests a very low irrigation amount whose gradient increases with increasing water level.
-	\item The diminishing margin utility of farmers lead to a decreasing gradient for the proposed irrigation amount with increasing water level.
+	\item As long as the discount factor $\beta$ is larger than 0, the water within the reservoir has a value in the next period for both recreational users and farmers.
+Due to this effect, saving waters makes sense and leads to a decreasing gradient for the proposed irrigation amount with increasing water level.
 \end{itemize}
@@
 In order to find out, how the steady states are distributed and what the probability of their occurrences are, we did a Monte Carlo simulation.
-To do so, we ran our code 1000 times and put the resulting water levels into bins with the size of 0.1 each. % "the forward simulation" instead of "our code"? Außerdem laut code 500x statt 1000x
+To do so, we ran our code 1000 times (in specific, the forward simulation) and put the resulting water levels into bins with the size of 0.1 each. 
 We then counted the occurrences, normalized them and calculated the corresponding probability.
@@
 A small $\beta$ ensures that high water levels at the end of one period are only available to a small extent in the next period. 
-% Finde die Formulierung nicht so gut: Sie sind ja weiterhin voll verfügbar, aber eben in Zukunft weniger wert.
-Bzw das könnte man diskutieren.
+%TODO Finde die Formulierung nicht so gut: Sie sind ja weiterhin voll verfügbar, aber eben in Zukunft weniger wert. Bzw das könnte man diskutieren.
 This leads to a increasing gap between irrigation amounts over $\beta$ with increasing water level since conserving water does not increase the usage as much.
@@
 Since we want to compare two rainfall distributions of the same mean, the median of the distribution has to be changed.
-Using a $\sigma^2=2$ includes a mean of $\mu=-1.5$ in order for the mean to stay the same. % Würde das erste "mean of mu" rausnehmen oder spezifizieren -> mean of the underlying normal distribution
+Using a $\sigma^2=2$ includes setting $\mu=-1.5$ in order for the mean to stay the same.
 (See \ref{eq:rainfall} for reference) 
@@
 Since the amount of rain is scaled down to match our model, there is no comparison between means of the model possible.
-Nevertheless one can say that the overall shape of a lognormal distribution is a feasible estimation for rainfall but for estimating whole years as periods the distribution might to be parameterized differently to fit actual rainfalls.
-% feasible -> sensible?
+Nevertheless one can say that the overall shape of a lognormal distribution is a sensible estimation for rainfall but for estimating whole years as periods the distribution might to be parameterized differently to fit actual rainfalls.
 As shown in \ref{fig:optimalIrrigationPolicy-real-data} the changed distribution of rainfalls also has an effect on the optimal irrigation policy since it depends on the expected amount of rain.
@@
 \begin{figure}[h]
-	\includegraphics[width=1\textwidth]{figures/results-evap-factor-2.3.eps}
+	\includegraphics[width=1\textwidth]{figures/results-evap-factor-2-3.eps}
 	\caption{Evaporation multiplied by a factor of $2.3$ outweighs the rainfall in certain periods} %TODO 

### 2020-02-24 Paper/paper.tex
changes as discussed

@@
 
-\subsection{Dynamic Programming}
+\subsubsection{Dynamic Programming}
 
@@
 
-\subsection{The Bellman Equation}
+\subsubsection{The Bellman Equation}
 
@@
 %TODO hier noch auf die Definition von beta eingehen? Was ist beta und warum ist das relevant?
-
-\subsection{Computation of the value function}
+\subsection{Implementation}
+\subsubsection{Computation of the value function}
 
@@
 In our case the function converges, if the norm of the difference between $V'$ and $V$ falls below a certain tolerance value \\
+% TODO Fehlen hier nicht noch ein paar Sätze dazu wie wir die optimal irrigation policy extrahieren?
 
-% TODO Fehlen hier nicht noch ein paar Sätze dazu wie wir die optimal irrigation policy extrahieren?
+\subsubsection{TODO: Forward Simulation}
 
@@
 The highest possibility for a steady state is around 4.2. 
-In edge cases one can find steady states as low as 3.5 or as high as 5.2. 
-The curve resembles a normal distribution.
+The curve resembles a normal distribution with a standard deviation of $\sigma=0.37$.
 
@@
 \\\\
-We also looked at how changes in the discount factor $\beta$ influenced the optimal irrigation policy (Figure 7).
+We also looked at how changes in the discount factor $\beta$ influenced the optimal irrigation policy (\ref{fig:optimal-irrigation-beta}).
 As expected, a high $\beta$ leads to less irrigation at high water levels.
@@
 \\
-We then went on and looked at the effect of increased rainfall volatility (Figures 9 and 10).
+We then went on and looked at the effect of increased rainfall volatility (\ref{fig:aggregated-use-rainfall-default} and \ref{fig:aggregated-use-rainfall-more-volatile}).
 Because of the constant mean, the total amount of rain and therefore of water influx in the reservoir stays the same.

### 2020-02-24 Paper/paper.tex
fix figures

@@
 \end{subfigure}
+\caption{Using Real Data for Rainfall}
+\label{fig:real-data}
 \end{figure}
@@
 	\caption{Without Evaporation}
-	\label{fig:rainfall-real-data}
+	\label{fig:water-level-wo-evap}
 \end{subfigure}%
@@
 	\caption{With Evaporation}
-	\label{fig:value-function}
+	\label{fig:total-utility-wo-evap}
 \end{subfigure}
@@
 	\caption{Water Level}
-	\label{fig:rainfall-real-data}
+	\label{fig:water-level-evap}
 \end{subfigure}%
@@
 	\caption{Total Utility}
-	\label{fig:value-function}
+	\label{fig:total-utility-evap}
 \end{subfigure}
@@
 Instead of simulated rainfall, we also used real weather data from Barisal, Bangladesh. 
-The real rainfall distribution (Figure 11 a) looked quite a bit different then our lognormal distribution with $\mu=0$ and $\sigma=1$.
+The real rainfall distribution (\ref{fig:rainfall-real-data}) looked quite a bit different then our lognormal distribution with $\mu=0$ and $\sigma=1$.
 This indicates that other parameters will have to be found to better describe the annual rainfall in that specific area.
-The higher expected value of rain in the real distribution leads to the optimal irrigation policy showing a higher amount of water used for irrigation at higher water levels (Figure 11 b).
+The higher expected value of rain in the real distribution leads to the optimal irrigation policy showing a higher amount of water used for irrigation at higher water levels (\ref{fig:optimalIrrigationPolicy-real-data}).
 Further adding evaporation to the model shows...

### 2020-02-24 Paper/paper.tex
forward simulation

@@
 
-\subsubsection{TODO: Forward Simulation}
-
+\subsubsection{Forward Simulation}
+In order to simulate the problem, we developed a so called forward simulation which is run once for each period to be simulated.
+It consists of several steps that will be described in more detail here.
+At the begin of each period the code determines how much water according to the irrigation policy to be used for irrigation.
+Since the simulation starts with an empty reservoir the amount irrigated in the first period is zero.
+Then the water level for the next iteration is calculated.
+This calculation is based on the irrigation amount, the current water level, the amount of rain and the amount that went away by evaporation (see \ref{sec:intro-mod4}).
+Afterwards the simulation calculates the difference in water level of the period compared to its predecessor, which will be used for finding a steady state.
 % Finally, we want to compute the value of the steady state by using a monte carlo simulation.

### 2020-02-24 Paper/paper.tex
unify headings

@@
 
-\subsubsection{Modification 2: Increased Rainfall Volatility}
+\subsubsection{Modification 2: Increased Volatility of Rainfalls}
 To consider the predicted impacts of climate change \citep{guhathakurta2011impact}, we look at the effect of increased rainfall volatility. 
@@
 
-\subsection{Evaporation}
+\subsection{Taking Evaporation Into Account}
 In the following, we want to assess the effect of evaporation on the optimal irrigation policy and the steady state. 

### 2020-02-24 Paper/paper.tex
Conclusion finished

@@
 A low $\beta$ on the other hand leads to a higher amount of water used for irrigation, especially at higher water levels. 
-This can be easily imagined with the extreme case $\beta = 0$, which means that no one benefits from leaving water in the reservoir for the next period.
+This can be easily imagined with the extreme case of $\beta = 0$, which means that no one benefits from leaving water in the reservoir for the next period.
 Of course there is still water that's being left in the reservoir, but that is only to benefit citizens in the current period and not because of the benefits of the remaining water for the next period.
@@
 The aggregated use is a good indicator of the efficiency our allocation strategy is achieving.
-We could see that with rainfalls getting more volatile the sum of aggregated use over all simulated periods drops significantly, due to efficiency losses caused by increased likelihood of reservoir overflow and drought.
-% Evlt noch schnell ausrechnen?
+We could see that with rainfalls getting more volatile the sum of aggregated use over all simulated periods drops significantly, due to utility losses caused by increased likelihood of reservoir overflow and drought.
 This vividly illustrates a very real consequence of climate change that governments and counties might face in the future.
@@
 The higher expected value of rain in the real distribution leads to the optimal irrigation policy showing a higher amount of water used for irrigation at higher water levels (\ref{fig:optimalIrrigationPolicy-real-data}).
-Further adding evaporation to the model shows...
-%Evap
-% Efficiency loss with higher evap scale factor -> same as more volatile rainfalls
-
-% Ausblick
-% Whole different problem, more of empirical nature: What even is the "right" discount factor in this context
+Further adding evaporation to the model lowers the average amount of water in the reservoir and therefore also the average amount of water used for irrigation and the sum of utility the optimal irrigation strategy gives us (\ref{fig:total-utility-evap}).
+In the scenario with the highest scale factor for evaporation the reservoir runs completely dry in multiple periods, which leads to a utility of negative infinity in all of those periods.
+This could be sensible in a scenario of an isolated village, where the survival of villagers completely depends on the local farmers' food production.
+In this case our model would predict a famine, which under all circumstances has to be prevented.
+An option would be to supply the village from the outside and assign a value to using this option to respect it in the model.
+The value should somehow correlate to the effort this would take.
+\\
+Summing up, we created a basic model of a water allocation problem that can be solved to optimality by using dynamic programming and can thereby support decisionmakers for the allocation of limited water resources.
+The model can work with random distributions and with real datasets and could of course also work with random distributions fitted to real datasets.
+To get more satisfactory recommendations from the model additional parameters and/or policies could be introduced.
+For example it would make sense to consider the temperature during the summer months when it comes to determine the utility citizens get from recreational water use.
+An additional policy that might be worth considering in the context could be food supply from external sources, as suggested above.
+\\
+As we learned from our research, the discount factor plays a big role when it comes to finding the optimal irrigation strategy.
+A whole different problem, more of an empirical nature however, would therefore be the determination of the "correct" discount factor, whatever that may be.
+This is a different area of research and a lot of publications are dedicated to this topic but further advancements could greatly strengthen the theoretical foundation of our model.
 

### 2020-02-24 Paper/paper.tex
changes in chapters 1 and 3 (3.1-3.5)

@@
 This situation will get even worse due to the projected increase in the demand of water per capita in many regions of the world, driven by lifestyle factors, especially diet \citep{vorosmarty2000global}.
-Despite all those challenges, research suggests that just by using water more efficiently, we will be able to meet the acute water, environment and poverty challenges facing us over the next 50 years \citep{molden2013water}.
+However, research suggests that just by using water more efficiently, we will be able to meet the challenges in water, environment and poverty facing us over the next 50 years \citep{molden2013water}.
 Therefore, the optimal allocation and sustainable use of the available water will be of increasing importance in the future.
@@
 One of the most important uses of surface water is the irrigation of fields for food production.
-However, as farming and food production are private enterprises and lakes are public goods, interests of all different parties have to be considered.
+However, as farming and food production are private enterprises and lakes are public goods, interests of different parties have to be considered.
 The water reservoir to be considered has two stakeholders: farmers, who want to use the water for irrigation purposes and citizens, who want to use the lake for recreational activities. 
-Note that while looking at just two agents is a vast simplification compared to a real allocation problem, one can easily add more stakeholders to the model, as long as there is enough computational power.
-The sequence of events for every period is displayed in \ref{fig:Sequence}.
+Though looking at just two agents is a vast simplification compared to real allocation problems, one can easily add more stakeholders to the model, as long as there is enough computational power.
+The sequence of events for every period is displayed in \ref{fig:Sequence}. %TODO Die Figure ist im PDF zu weit weg gerutscht! 
 At the beginning of every year (Step 1) we have to decide how much of the water from the reservoir is used for the irrigation of fields, which benefits the farmers' harvest.\\\\
@@
 
-where $x$ represents the amount of water to be used for irrigation in a year. The preference parameters are $a_1 = 1$ and $b_1 = -2$.\\\\
+where $x$ represents the amount of water to be used for irrigation during a period. 
+The preference parameters are given as $a_1 = 1$ and $b_1 = -2$.\\\\
 
-During the summer months, citizens benefit from the remaining water due to recreational use (Step 2).
-This can be described by the recreational users' utility function which is
+During the summer months, citizens benefit from the remaining water in their recreational activities (Step 2).
+This can be described by the recreational users' utility function
 
@@
 
-with the additional paramter $s$ that indicates the water level at the beginning of the year.
+with the additional parameter $s$ that indicates the water level at the beginning of the year.
 For recreational users, the preference parameters shall be $a_2 = 2$ and $b_2 = -3$.\\\\
 
-At the end of the year, water in the reservoir is replenished by rainfall (Step 3), which we can assume to be randomly following a lognormal distribution \citep{oosterbaan1994frequency}:
+At the end of the year, water in the reservoir is replenished by rainfall (Step 3), which we can assume to be randomly distributed following a lognormal distribution \citep{oosterbaan1994frequency}: %TODO cite steht über den Rand hinaus
 \begin{equation}
@@
 
-Using dynamic programming in combination with a stochastic simulation of the rainfall, we determined the optimal irrigation policy that maximizes the combined utility of farmers and citizens for a number of years $T$.
+In this work, we will determine an optimal irrigation policy that maximizes the combined utility of farmers and citizens over $T$ years by using dynamic programming in combination with a stochastic simulation of the rainfall.
 In the main part we will give a brief summary of the underlying theory and the optimization problem. 
 After that we will present the results and our conclusions for the original model and its modifications, as well as possible shortfalls of our analysis.
-At the end we also give a little outlook into further possible extensions.
+%TODO statt "possible shortfalls" besser "possible enhancements" schreiben?
+(klingt positiver?)
+Finally, we will give an outlook for further possible extensions.
 
 \subsubsection{Modification 1: Variation of the Discount Factor}
-We also study the effects of different betas on the optimal irrigation policy. 
+In this modification, we're interested in studying the effects of different betas on the optimal irrigation policy. 
 The discount factor takes into account that agents in a simple two-stage model often value current consumption over future consumption.
-It expresses how much one unit of future utility is worth in terms of current utility and therefore makes them easily comparable \citep{kruschwitz2014investitionsrechnung}.
-We take a look at the discount factors from 0 to 1.
+It expresses how much one unit of future utility is worth in terms of current utility and therefore makes them mathematically comparable \citep{kruschwitz2014investitionsrechnung}.
+%TODO wirklich deutsche Citation verwenden?
+We take a look at the discount factors from X to Y. %TODO tatsächliche Grenzen einfügen
 
-\subsubsection{Modification 2: Increased Volatility of Rainfalls}
-To consider the predicted impacts of climate change \citep{guhathakurta2011impact}, we look at the effect of increased rainfall volatility. 
-We do this by using a  $\sigma^2$ of x. % TODO: Das ist noch nicht ganz richtig
+\subsubsection{Modification 2: Increasing the Volatility of Rainfalls}
+In order to take into account expected future impacts of climate change \citep{guhathakurta2011impact}, we consider the effects of increased volatility of precipitation.
+%TODO cite steht über den Rand hinaus
+We do so by setting $\sigma$ to 2. %TODO Das ist noch nicht ganz richtig
 
-\subsubsection{Modification 3: Using Real Data for Rainfall}
-Instead of simulating the rainfall with random numbers, we also use real data.
-Comparing the results then allows us to draw conclusions on the accuracy of the simulation for that region. 
-We use a dataset aquired from \citep{kaggle:2019}, that contains real rainfall data for the last 65 years in different regions of Bangladesh. % Zitation anpassen
-To make this data fit our somewhat arbitrarily selected range of 0 to 7 for water levels in the reservoir, we...
-% TODO describe how we fitted them to match the rainfall distribution
+\subsubsection{Modification 3: Using Real Data for Rainfalls}
+Instead of simulating the rainfall with randomly generated numbers, we want to use real data. 
+We make use of a dataset aquired from \citep{kaggle:2019}, which contains rainfall and other weather related data from the last 65 years in different regions of Bangladesh (here we will focus on the city Barisal).
+%TODO Zitation anpassen
+In order to make the dataset compatible to the basic model, we assume an exemplary surface-to-volume-ratio of the reservoir.
+For the sake of this simulation, we multiply the given rainfall by a factor of $4$, in order to appropriately replenish the reservoir.
 
 \subsubsection{Modification 4: Taking Evaporation into Account} \label{sec:intro-mod4}
-The evaporation of water from a water reservoir is the most important natural "outflow" to be considered when it comes to managing water levels \citep{tanny2008evaporation}.
-It therefore should definitely be considered in a real model that is used to determine the optimal irrigation strategy.
+The evaporation of water from a reservoir is the most important natural "outflow" to be considered when it comes to managing water levels \citep{tanny2008evaporation}. 
+To determine a more realistic optimal irrigation strategy, we should therefore consider evaporation effects in our model. 
 The dataset that we use for the real rainfall also contains many other weather parameters.
-With the mean temperature $T$, the elevation $h$, the latitude $A$ and the dew point $T_d$ we calculate an estimate for evaporation of water in mm per day for that specific location, using the formula \citep{linacre1977simple}
+The dataset from \citep{kaggle:2019} does not contain evaporation information directly.
+However, with the mean temperature $T$ in °C, the elevation $h$ in m, the latitude $A$ in °S and the dew point $T_d$ in °C we can calculate an estimate for the evaporation of water in mm per day for a specific location, using the formula \citep{linacre1977simple}
 
@@
 
-where RH is the average relative humidity of the year in percent \citep{lawrence2005relationship}.\\\\
+where $RH$ is the average relative humidity of the year in percent \citep{lawrence2005relationship}.\\\\
 
+%TODO sind diese beiden Sätze so schön? oder nötig?
 Note that all modifications, except Modification 4, are based on the original model without any modifications. 
@@
 \subsection{Optimal irrigation policy}
-When questioning what the optimal irrigation policy is, the first question to be asked is how the aggregated value function of both recreational users and farmers looks like.
-The value function as posed in \ref{eq:value-function} contains the maximum possible usage that can be gained by irrigating a specific amount of water under a given water level. 
-Figure \origref{fig:value-function} shows the value function calculated for all possible water levels between the allowed boundaries of 0 and 7. 
+When examining what an optimal irrigation policy could be, the first question to be asked is how the aggregated value function of both recreational users and farmers looks like. 
+The value function as posed in \ref{eq:value-function} contains the maximum possible (expected) utility that can be gained by irrigating a certain amount of water under a given water level in the reservoir.
+Figure \origref{fig:value-function} shows the value function calculated for all possible water levels between the allowed boundaries of $0$ and $7$. 
 \begin{figure}[ht]
@@
 \newline
-Every value along the possible water levels is mapped to an irrigation amount that optimizes the value for all parties in the given situation.
-Summarizing these values over the range of possible water levels results in a optimal irrigation policy which is shown in \ref{fig:optimal-irrigation-policy}.
-The plot shows a rising proposed irrigation amount with rising water level. There are 2 particularities that can be found within the graph:
+Every value along the possible water levels is mapped to an irrigation amount that optimizes the expected value for all parties in the given situation. 
+The resulting optimal irrigation policy is shown in \ref{fig:optimal-irrigation-policy}. 
+As the plot shows, a higher water level leads to a higher proposed irrigation amount. There are two particularities that can be found within the graph:
 \begin{itemize}
-	\item Due to the effect that a almost empty reservoir has a very low use to recreational users, the optimal irrigation strategy suggests a very low irrigation amount whose gradient increases with increasing water level.
-	\item As long as the discount factor $\beta$ is larger than 0, the water within the reservoir has a value in the next period for both recreational users and farmers.
+	\item Due to the effect that an almost empty reservoir has a very low use to recreational users, the optimal irrigation strategy suggests a very low irrigation amount whose gradient increases with increasing water level.
+	\item As long as the discount factor $\beta$ is greater than $0$, the water within the reservoir has a value in the next period for both recreational users and farmers.
 Due to this effect, saving waters makes sense and leads to a decreasing gradient for the proposed irrigation amount with increasing water level.
@@
 The variable defining the behavior of our system is the water level. 
-Because of this, we have to calculate the difference in water level for each period. 
-Due to the fact, that our water level highly depends on the randomly distributed rainfalls that fill the reservoir at end of each period, it is not possible to find a steady state in which the water level does not change at all.
-To analyze it nevertheless, we defined our steady state, to be a state, where the mean water level of the last 200 periods did not change more than $1*10^{-7}$ from a period to the next.
+Thus, we have to calculate the change in water level for each period. 
+Due to the fact that the water level highly depends on the randomly distributed rainfalls (that fill the reservoir at end of each period), it is not possible to find a steady state in which the water level does not change \textit{at all}.
+In this context, we define a steady state to be a state, where the mean water level of the last 200 periods does not change more than $1*10^{-7}$ from one period to the next.
+%TODO ist dieser Wert noch richtig?
+Im Code steht im Moment steadyStateTolerance = 0.0001;
 A exemplary graph of the mean in every period and the difference to its predecessor is shown in \ref{fig:steadyState}.
-Once a period is found, where the mean does not change more than the given value, the mean of that period is defined as steady state water level.
-The water level found to be a steady state and the period in which this condition occurs changes from calculation to calculation due to the fact, that the rainfall is still a random distribution.
+Once a period is found where the mean does not change more than the given tolerance, the mean of that period is defined as the steady state of the water level. %TODO hier "moving average" schreiben?
+The water level found to be a steady state and the period in which this condition occurs changes from one simulation to the next, since the rainfall is generated randomly following the given distribution.
+
 \begin{figure}[ht]
@@
 \end{figure}
+
 \subsection{Steady State Probability}
-As mentioned above, the steady state changes from simulation to simulation.
-In order to find out, how the steady states are distributed and what the probability of their occurrences are, we did a Monte Carlo simulation.
-To do so, we ran our code 1000 times (in specific, the forward simulation) and put the resulting water levels into bins with the size of 0.1 each. 
+As just mentioned, the steady state changes from simulation to simulation. 
+In order to find out how the steady states are distributed and what the probability of their occurrences are, we implement a Monte Carlo simulation. 
+To do so, we evaluate the forward simulation 1000 times and put the resulting water levels into bins with the size of 0.1 each. 
 We then counted the occurrences, normalized them and calculated the corresponding probability.
 The result is put in a histogram and shown in \ref{fig:steadyStateMonteCarlo}.
+ %TODO ich (martin) würde vorschlagen, einfach zu schreiben ...
+"1000 times and calculate a relative frequency distribution" (ist hier erklärt: https://www.sciencedirect.com/topics/mathematics/relative-frequency-distribution)
 The highest possibility for a steady state is around 4.2. 
-The curve resembles a normal distribution with a standard deviation of $\sigma=0.37$.
+The curve resembles a normal distribution with a standard deviation of $\sigma=0.37$ and peaks at a water level around $4.2$.
 
@@
 \end{figure}
-\subsection{Effect of Changed Discount Factors to Optimal Irrigation Policy}
-When altering the discount factor $\beta$ in a range from 0 to 1, the optimal irrigation policy changes.
-Looking at small water levels within the reservoir the optimal irrigation amount does not change much depending on the discount factor.
+
+\subsection{Effect of the Discount Factor on the Optimal Irrigation Policy}
+When altering the discount factor $\beta$ in a range from 0 to 1, the optimal irrigation policy changes. %TODO auch hier die tatsächlichen Werte anpassen?!
+For small water levels of the reservoir, the optimal irrigation amount is rather insensitive to the discount factor. 
 A small $\beta$ ensures that high water levels at the end of one period are only available to a small extent in the next period. 
 %TODO Finde die Formulierung nicht so gut: Sie sind ja weiterhin voll verfügbar, aber eben in Zukunft weniger wert. Bzw das könnte man diskutieren. 
-This leads to a increasing gap between irrigation amounts over $\beta$ with increasing water level since conserving water does not increase the usage as much.
+This leads to an increasing gap between irrigation amounts over $\beta$ with increasing water level since conserving water does not increase the usage as much. 
 One can say that a high $\beta$ prefers a long-term policy while a smaller $\beta$ leads to a rather short-term policy.
+%TODO ich (martin) würde diesen absatz so schreiben "A smaller $\beta$ allows more water to be used for irrigation today, optimizing the short-term utility.
+On the other hand, a higher $\beta$ increases the today's valuation of future utility, and therefore reduces the amount of water to be consumed (irrigated) today.
+"
 \begin{figure}[ht]
@@
 \newpage
-\subsection{Changed Volatility of Rainfalls}
-A more volatile rainfall means a rainfall having a higher standard deviation.
+\subsection{Increased Volatility of the Rainfalls} %TODO überall use durch utility ersetzen!
+More volatile rainfalls means that rainfalls are having a higher standard deviation.
 Since we want to compare two rainfall distributions of the same mean, the median of the distribution has to be changed. 
-Using a $\sigma^2=2$ includes setting $\mu=-1.5$ in order for the mean to stay the same.
-(See \ref{eq:rainfall} for reference) 
+Originally, we assumed $\mu=0$ and $\sigma=1$.
+Now, we choose $\sigma=2$ and thus have to set $\mu=-1.5$, in order to retain the same mean (see \ref{eq:rainfall}).
+%TODO ich (martin) hab diesen Absatz so umgeschrieben, einverstanden?
+Vorher stand hier insbesondere auch (fälschlicherweise?!) $\sigma^2=2$:
+%Since we want to compare two rainfall distributions of the same mean, the median of the distribution has to be changed.
+%Using a $\sigma^2=2$ includes setting $\mu=-1.5$ in order for the mean to stay the same.
+%(See \ref{eq:rainfall} for reference)
+
 \begin{equation}
@@
 	\end{subfigure}
-	\caption{Different Rain Distributions}
+	\caption{Different Rainfall Distributions}
 	\label{fig:rain-distributions}
@@
 \end{figure}
-\ref{fig:aggregated-use-rainfall-default} shows the aggregated use of a lognormal distributed rainfall including the water level, the amount of water used for irrigation in the rain exemplary for 100 periods.
-\ref{fig:aggregated-use-rainfall-more-volatile} does the same for a changed lognormal distribution parameterized with $\mu=-1.5$ and $\sigma=2$.
-When comparing both visualizations, it is noticeable that more volatile rainfall leads to much higher peaks of rain in combination with reduced number of periods with a decent amount of rain. 
-As a consequence the water level in the reservoir fluctuates much more leading to periods with almost no water in the reservoir. 
-In such periods the farmers use almost no water for irrigation.
-The aggregated use of farmers and recreational users drops significantly.
-Since the reservoir is capped at a amount of 7, the high peaks of rainfall can not be used to a large extent to fill up the reservoir.
+
+The following figures demonstrate the effect of the changed volatility.
+They show the lognormal rainfalls with the original volatility (\ref{fig:aggregated-use-rainfall-default}), and with the increased volatility (\ref{fig:aggregated-use-rainfall-more-volatile}) over 100 periods.
+When comparing both visualizations, we see that more extreme rainfalls significantly affect the water level of the reservoir: there longer periods with a rather low amount of water in the reservoir, but some periods where the reservoir reaches it's capacity limit.
+%When comparing both visualizations,it is noticeable that more volatile rainfall leads to much higher peaks of rain in combination with reduced number of periods with a decent amount of rain. 
+%As a consequence the water level in the reservoir fluctuates much more leading to periods with almost no water in the reservoir. 
+In the periods with only low precipitation, the farmers use almost no water for irrigation. 
+The aggregated utility of farmers and recreational users drops significantly. 
+Since the reservoir is capped at an amount of 7, the high peaks of the rainfalls can only partially be used in following (potentially dryer) periods. 
 Generally speaking, one can say that a more volatile rain distribution leads to a lower average water level in the reservoir with lower amounts of water to be used for irrigation.
-This effect is reflected in the aggregated use as well. 
-The use of more volatile rainfall is much lower compared to rainfall parameterized with $\mu=0$ and $\sigma=1$.
-Since the mean stays constant within the analyzed comparison, the value function does not change.
+This effect is reflected in the aggregated utility as well. 
+The utility of more volatile rainfall is much lower compared to rainfall parameterized with $\mu=0$ and $\sigma=1$. 
+Since the mean stays the same in the two cases considered, the value function does not change.
+
 \clearpage

### 2020-02-24 Code/code.m
typo

@@ -104,7 +104,7 @@ if (useRealData)
 
     % Note that because of the weather conditions of the location chosen in 
     % bangladesh, the amount of rainfall is not high enough to provide a
-    % significant amount of value through irrigatio for the farmers. For 
+    % significant amount of value through irrigation for the farmers. For 
     % this demo however, we want the reservoir to be refilled in order to 
     % give the farmers at least some utility. Thus we will
     % artificially increase the amount of rainfall here. 

### 2020-02-27 Paper/paper.tex
add TODO
fix a bunch of TODOs

@@
 Though looking at just two agents is a vast simplification compared to real allocation problems, one can easily add more stakeholders to the model, as long as there is enough computational power.
-The sequence of events for every period is displayed in \ref{fig:Sequence}. %TODO Die Figure ist im PDF zu weit weg gerutscht! 
+The sequence of events for every period is displayed in \ref{fig:Sequence}.
 At the beginning of every year (Step 1) we have to decide how much of the water from the reservoir is used for the irrigation of fields, which benefits the farmers' harvest.\\\\
 The farmers utility function follows the form
+
+\begin{figure}[ht]
+	\includegraphics[width=1\textwidth]{figures/CESchaubild_Cut.png}
+	\caption{Sequence of Events in One Period}
+	\label{fig:Sequence}
+\end{figure}
 
@@
 
-At the end of the year, water in the reservoir is replenished by rainfall (Step 3), which we can assume to be randomly distributed following a lognormal distribution \citep{oosterbaan1994frequency}: %TODO cite steht über den Rand hinaus
+At the end of the year, water in the reservoir is replenished by rainfall (Step 3), which we can assume to be randomly distributed following a lognormal distribution \citep{oosterbaan1994frequency}: 
 \begin{equation}
@@
 
-\begin{figure}[ht]
-	\includegraphics[width=1\textwidth]{figures/CESchaubild_Cut.png}
-	\caption{Sequence of Events in One Period}
-	\label{fig:Sequence}
-\end{figure}
-
 In this work, we will determine an optimal irrigation policy that maximizes the combined utility of farmers and citizens over $T$ years by using dynamic programming in combination with a stochastic simulation of the rainfall.
 In the main part we will give a brief summary of the underlying theory and the optimization problem. 
-After that we will present the results and our conclusions for the original model and its modifications, as well as possible shortfalls of our analysis.
-%TODO statt "possible shortfalls" besser "possible enhancements" schreiben?
-(klingt positiver?)
+After that we will present the results and our conclusions for the original model and its modifications, as well as possible enhancements of our analysis.
 Finally, we will give an outlook for further possible extensions.
@@
 It expresses how much one unit of future utility is worth in terms of current utility and therefore makes them mathematically comparable \citep{kruschwitz2014investitionsrechnung}. 
-%TODO wirklich deutsche Citation verwenden?
-We take a look at the discount factors from X to Y. %TODO tatsächliche Grenzen einfügen
+We take a look at the discount factors from 0.05 to 0.95.
 
 \subsubsection{Modification 2: Increasing the Volatility of Rainfalls}
-In order to take into account expected future impacts of climate change \citep{guhathakurta2011impact}, we consider the effects of increased volatility of precipitation.
-%TODO cite steht über den Rand hinaus
-We do so by setting $\sigma$ to 2. %TODO Das ist noch nicht ganz richtig
+In order to take into account expected future impacts of climate change \citep{guhathakurta2011impact}, we consider the effects of increased volatility
+We do so by setting $\sigma=2$ and $\mu=-1.5$.
 
@@
 \subsubsection{Modification 4: Taking Evaporation into Account} \label{sec:intro-mod4}
-The evaporation of water from a reservoir is the most important natural "outflow" to be considered when it comes to managing water levels \citep{tanny2008evaporation}. 
+The evaporation of water from a reservoir is the most important natural "drain" to be considered when it comes to managing water levels \citep{tanny2008evaporation}. 
 To determine a more realistic optimal irrigation strategy, we should therefore consider evaporation effects in our model. 
@@
 where $RH$ is the average relative humidity of the year in percent \citep{lawrence2005relationship}.\\\\
-
-%TODO sind diese beiden Sätze so schön? oder nötig?
-Note that all modifications, except Modification 4, are based on the original model without any modifications. 
-Modification 4 is an extension of Modification 3.
 
@@
 In his paper "The Theory Of Dynamic Programming" from 1957, Richard Bellman aimed to provide a solution to multi-stage-decision processes. 
+% TODO unbedingt noch Bellman als Quelle hinzufügen!!!
 He defines a \textit{policy} as "a sequence of decisions" (\cite{bellman:1957}, p.2), and an \textit{optimal policy} as "a policy, which is most advantageous according to some preassigned criterion" (\cite{bellman:1957}, p.2).
@@
 \subsection{Implementation}
+%TODO theoretisch ist es sehr unschön ein Kapitel ohne Text aufzumachen. Ist aber nur Makulatur
 \subsubsection{Computation of the value function}
@@
 We repeat this step over and over again until the our value function $V$ converges.
-In our case the function converges, if the norm of the difference between $V'$ and $V$ falls below a certain tolerance value \\
-% TODO Fehlen hier nicht noch ein paar Sätze dazu wie wir die optimal irrigation policy extrahieren?
+In our case the we define the function to be converging, if the norm of the difference between $V'$ and $V$ falls below a certain tolerance value.
+Summarizing the optimal irrigation amounts for all possible water levels into one function results in the optimal irrigation policy.
 
@@
 Due to the fact that the water level highly depends on the randomly distributed rainfalls (that fill the reservoir at end of each period), it is not possible to find a steady state in which the water level does not change \textit{at all}.
-In this context, we define a steady state to be a state, where the mean water level of the last 200 periods does not change more than $1*10^{-7}$ from one period to the next.
-%TODO ist dieser Wert noch richtig?
-Im Code steht im Moment steadyStateTolerance = 0.0001;
+In this context, we define a steady state to be a state, where the mean water level of the last 200 periods does not change more than $1*10^{-4}$ from one period to the next. 
 A exemplary graph of the mean in every period and the difference to its predecessor is shown in \ref{fig:steadyState}.
-Once a period is found where the mean does not change more than the given tolerance, the mean of that period is defined as the steady state of the water level. %TODO hier "moving average" schreiben?
+Once a period is found where the mean does not change more than the given tolerance, the mean of that period is defined as the steady state of the water level. 
 The water level found to be a steady state and the period in which this condition occurs changes from one simulation to the next, since the rainfall is generated randomly following the given distribution.
@@
 The result is put in a histogram and shown in \ref{fig:steadyStateMonteCarlo}.
- %TODO ich (martin) würde vorschlagen, einfach zu schreiben ...
-"1000 times and calculate a relative frequency distribution" (ist hier erklärt: https://www.sciencedirect.com/topics/mathematics/relative-frequency-distribution)
 The highest possibility for a steady state is around 4.2. 
@@
 \subsection{Effect of the Discount Factor on the Optimal Irrigation Policy}
-When altering the discount factor $\beta$ in a range from 0 to 1, the optimal irrigation policy changes. %TODO auch hier die tatsächlichen Werte anpassen?!
+When altering the discount factor $\beta$ in a range from 0.05 to 0.95, the optimal irrigation policy changes. 
 For small water levels of the reservoir, the optimal irrigation amount is rather insensitive to the discount factor. 
-A small $\beta$ ensures that high water levels at the end of one period are only available to a small extent in the next period. 
-%TODO Finde die Formulierung nicht so gut: Sie sind ja weiterhin voll verfügbar, aber eben in Zukunft weniger wert. Bzw das könnte man diskutieren. 
+A small $\beta$ ensures that high water levels at the end of one period only keep their value only to a small extent in the next period.  
 This leads to an increasing gap between irrigation amounts over $\beta$ with increasing water level since conserving water does not increase the usage as much. 
-One can say that a high $\beta$ prefers a long-term policy while a smaller $\beta$ leads to a rather short-term policy.
-%TODO ich (martin) würde diesen absatz so schreiben "A smaller $\beta$ allows more water to be used for irrigation today, optimizing the short-term utility.
+A smaller $\beta$ allows more water to be used for irrigation today, optimizing the short-term utility.
 On the other hand, a higher $\beta$ increases the today's valuation of future utility, and therefore reduces the amount of water to be consumed (irrigated) today.
-"
 \begin{figure}[ht]
@@
 \newpage
-\subsection{Increased Volatility of the Rainfalls} %TODO überall use durch utility ersetzen!
+\subsection{Increased Volatility of the Rainfalls} 
 More volatile rainfalls means that rainfalls are having a higher standard deviation.
 Since we want to compare two rainfall distributions of the same mean, the median of the distribution has to be changed. 
-Originally, we assumed $\mu=0$ and $\sigma=1$.
-Now, we choose $\sigma=2$ and thus have to set $\mu=-1.5$, in order to retain the same mean (see \ref{eq:rainfall}).
-%TODO ich (martin) hab diesen Absatz so umgeschrieben, einverstanden?
-Vorher stand hier insbesondere auch (fälschlicherweise?!) $\sigma^2=2$:
-%Since we want to compare two rainfall distributions of the same mean, the median of the distribution has to be changed.
-%Using a $\sigma^2=2$ includes setting $\mu=-1.5$ in order for the mean to stay the same.
-%(See \ref{eq:rainfall} for reference)
+Originally, we assumed $\mu=0$ and $\sigma=1$. Now, we choose $\sigma=2$ and thus have to set $\mu=-1.5$, in order to retain the same mean (see \ref{eq:rainfall}).
 
@@
 The expected value of the evaporation with a scaling factor of $2.3$ is just below the expected value of the rainfall.
-} %TODO Fußnote gut so?
+}
 We can see that higher amounts of evaporation not only lead to water levels, but also to significantly lower utilities for the recreational users and the farmers. 
@@
 \section{Conclusion}
-%TODO content
 In this work we developed an optimal irrigation strategy for an hypothetical water reservoir, maximizing the sum of utilities for farmers and recreational users.
@@
 Running our forward simulation 500 times, we could see that it seems to follow a normal distribution with a mean of around 4.2 and a standard deviation of 0.37.
-% TODO: StdDev auch im Ergebnisteil erwähenen?
 \\\\
@@
 A whole different problem, more of an empirical nature however, would therefore be the determination of the "correct" discount factor, whatever that may be.
+% TODO "whatever that may be" klingt echt merkwürdig
 This is a different area of research and a lot of publications are dedicated to this topic but further advancements could greatly strengthen the theoretical foundation of our model.

### 2020-02-27 Paper/paper.tex
Minor overhaul, still some TODOs

@@
 However, as farming and food production are private enterprises and lakes are public goods, interests of different parties have to be considered.
-The water reservoir to be considered has two stakeholders: farmers, who want to use the water for irrigation purposes and citizens, who want to use the lake for recreational activities. 
+The water reservoir we're modeling has two stakeholders: farmers, who want to use the water for irrigation purposes and citizens, who want to use the lake for recreational activities. 
 Though looking at just two agents is a vast simplification compared to real allocation problems, one can easily add more stakeholders to the model, as long as there is enough computational power.
@@
 \subsubsection{Modification 2: Increasing the Volatility of Rainfalls}
-In order to take into account expected future impacts of climate change \citep{guhathakurta2011impact}, we consider the effects of increased volatility
-We do so by setting $\sigma=2$ and $\mu=-1.5$.
+In order to take into account expected future impacts of climate change \citep{guhathakurta2011impact}, we consider the effects of increased volatility.
+We do so by changing the parameters of the lognormal distribution of the rainfall to $\sigma^2=4$ and $\mu=-1.5$.
 
@@
 A common technique to solve optimization or combinatorial problems is called Dynamic Programming. 
-It is based on the idea of dividing problems into smaller subproblems, solving these recursively and caching intermediate results (\textit{memoization}). % Nicht eher "memoRization"?
+It is based on the idea of dividing problems into smaller subproblems, solving these recursively and caching intermediate results (\textit{memoization}). % TODO Nicht eher "memoRization"?
 In his paper "The Theory Of Dynamic Programming" from 1957, Richard Bellman aimed to provide a solution to multi-stage-decision processes. 
@@
 So the value of the control variable to be set is the one that maximizes both the utility of the current state plus (recursively) the value $V$ of the next state $s'$ which is discounted by a factor of $\beta$.
-%TODO hier noch auf die Definition von beta eingehen? Was ist beta und warum ist das relevant?
 \subsection{Implementation}
@@
 Since we want to compare two rainfall distributions of the same mean, the median of the distribution has to be changed.
-Originally, we assumed $\mu=0$ and $\sigma=1$. Now, we choose $\sigma=2$ and thus have to set $\mu=-1.5$, in order to retain the same mean (see \ref{eq:rainfall}).
+% TODO: Das macht für mich nicht so ganz Sinn; entweder: "parameter mu has to also be changed..." oder "mean of the underlying normally distributed variable has to be changed"
+Originally, we assumed $\mu=0$ and $\sigma^2=1$. Now, we choose $\sigma^2=4$ and thus have to set $\mu=-1.5$, in order to retain the same mean (see \ref{eq:rainfall}).
 
@@
 		\includegraphics[width=1\textwidth]{figures/rainfall-default.eps}
-		\caption{$\mu=0, \sigma=1$}
+		\caption{$\mu=0, \sigma^2=1$}
 		\label{fig:rainfall-default}
@@
 		\includegraphics[width=1\textwidth]{figures/rainfall-more-volatile.eps}
-		\caption{$\mu=-1.5, \sigma=2$}
+		\caption{$\mu=-1.5, \sigma^2=4$}
 		\label{fig:rainfall-changed}
@@
 	\includegraphics[width=1\textwidth]{figures/aggregated-use-rainfall-default.eps}
-	\caption{Aggregated Use for lognormal distributed rainfall $(\mu=0, \sigma=1)$}
+	\caption{Aggregated Use for lognormal distributed rainfall $(\mu=0, \sigma^2=1)$}
 	\label{fig:aggregated-use-rainfall-default}
@@
 \includegraphics[width=1\textwidth]{figures/aggregated-use-rainfall-more-volatile.eps}
-\caption{Aggregated Use for lognormal distributed rainfall $(\mu=-1.5, \sigma=2)$}
+\caption{Aggregated Use for lognormal distributed rainfall $(\mu=-1.5, \sigma^2=4)$}
 \label{fig:aggregated-use-rainfall-more-volatile}
@@
 They show the lognormal rainfalls with the original volatility (\ref{fig:aggregated-use-rainfall-default}), and with the increased volatility (\ref{fig:aggregated-use-rainfall-more-volatile}) over 100 periods.
-When comparing both visualizations, we see that more extreme rainfalls significantly affect the water level of the reservoir: there longer periods with a rather low amount of water in the reservoir, but some periods where the reservoir reaches it's capacity limit.
+When comparing both visualizations, we see that more extreme rainfalls significantly affect the water level of the reservoir: there are longer periods with a rather low amount of water in the reservoir, but some periods where the reservoir reaches it's capacity limit.
 %When comparing both visualizations,it is noticeable that more volatile rainfall leads to much higher peaks of rain in combination with reduced number of periods with a decent amount of rain. 
@@
 This effect is reflected in the aggregated utility as well. 
-The utility of more volatile rainfall is much lower compared to rainfall parameterized with $\mu=0$ and $\sigma=1$. 
+The utility of more volatile rainfall is much lower compared to rainfall parameterized with $\mu=0$ and $\sigma^2=1$. 
 Since the mean stays the same in the two cases considered, the value function does not change.
@@
 The optimal irrigation policy stipulates quite low irrigation amounts for low water levels, mostly due to the very negative utility citizens receive from an almost empty reservoir (Figure 4).
-With $\beta = 0.9$, the optimal irrigation amount tapers off again at the end, due to % TODO: Besprechen!
-MMn wegen Möglichkeit des Mehrfachnutzens bei Non-irrigation (daher nicht erkennbar bei beta = 0)
-In figure 9 you can see one possible scenario in our basic setting with $\mu=0$ and $\sigma=1$.
+With $\beta = 0.9$, the optimal irrigation amount tapers off again at the end, due to the potential of benefitting from the remaining water again in the future (see next paragraph).
+In figure 9 you can see one possible scenario in our basic setting with $\mu=0$ and $\sigma^2=1$.
 However, due to the random variable, every forward simulation can look quite different, even though no parameters are changed.
@@
 That's because citizens might be able to benefit from the remaining water again in future periods, while when using water for irrigation farmers can only benefit once in the present period.
-Increasing $\beta$ means putting more weight on future utility up to a factor of $\beta = 1$, where stakeholders are indifferent between current and future utility.
+Increasing $\beta$ means putting more weight on future utility in relation to current utility.
+For example with a factor of $\beta = 1$, stakeholders are indifferent between current and future utility.
 A low $\beta$ on the other hand leads to a higher amount of water used for irrigation, especially at higher water levels. 
@@
 Instead of simulated rainfall, we also used real weather data from Barisal, Bangladesh. 
-The real rainfall distribution (\ref{fig:rainfall-real-data}) looked quite a bit different then our lognormal distribution with $\mu=0$ and $\sigma=1$.
+The real rainfall distribution (\ref{fig:rainfall-real-data}) looked quite a bit different then our lognormal distribution with $\mu=0$ and $\sigma^2=1$.
 This indicates that other parameters will have to be found to better describe the annual rainfall in that specific area.
@@
 As we learned from our research, the discount factor plays a big role when it comes to finding the optimal irrigation strategy.
-A whole different problem, more of an empirical nature however, would therefore be the determination of the "correct" discount factor, whatever that may be.
-% TODO "whatever that may be" klingt echt merkwürdig
-This is a different area of research and a lot of publications are dedicated to this topic but further advancements could greatly strengthen the theoretical foundation of our model.
+A whole different problem would therefore be the determination of the "correct" discount factor.
+This is a different area of research and a lot of publications are dedicated to this topic. % TODO Quellen hinzufügen!
+Still, further advancements could greatly strengthen the theoretical foundation of our model.
 

### 2020-02-28 Code/code.m
use->utilities, line breaks

@@ -421,30 +421,30 @@ else
     ylabel('Irrigation Amount')
     hold off
 
-    % Plot the aggregated use for farmers and recreational users
+    % Plot the aggregated utility for farmers and recreational users
     figure(10)
     hold on
     set(gca,'FontSize',12)
     set(gcf,'Units','Centimeters','position',[0,0,16,8]);
-    %title('Aggreated Use for Farmers and Recreational Users');
+    %title('Aggreated Utility for Farmers and Recreational Users');
     yyaxis left;
-    plot(utilRec(waterLevel(waterInd(:, 1:T)), waterLevel(irrigationInd)) + utilFar(waterLevel(irrigationInd)),'color','#A2142F');
-    ylabel('Use');
+    plot(utilRec(waterLevel(waterInd(:, 1:T)), waterLevel(irrigationInd)) + utilFar(waterLevel(irrigationInd)), ...
+         'color','#A2142F');
+    ylabel('Utility');
     ylim([-100 0]);
     yyaxis right;
-    plot(waterLevel(waterInd),'-', 'color', '#0072BD');
-    plot(waterLevel(irrigationInd),'-','color','#D95319');
-    plot(r,'-','color','#EDB120');
+    plot(waterLevel(waterInd), '-', 'color', '#0072BD');
+    plot(waterLevel(irrigationInd),'-','color', '#D95319');
+    plot(r, '-', 'color', '#EDB120');
     ylabel('Amount');
     ylim([0 10]);
-    legend('Aggregated Use','Water Level of the Reservoir','Water used for Irrigation',...
-    'Rain');
+    legend('Aggregated Utility', 'Water Level of the Reservoir', 'Water used for Irrigation', 'Rain');
     xlim([1 100]);
     xlabel('Period');
     hold off
-
+    
     figure(11)
-    set(gcf,'Units','Centimeters','position',[0,0,8,8]);
+    set(gcf, 'Units', 'Centimeters', 'position', [0,0,8,8]);
     edges = [0:0.05:7];
     hold on
     histogram(r, edges, 'Normalization', 'Probability')
@@ -456,13 +456,14 @@ end
 
 
 %% Computation of the Value Function
-function [V, optIrrigation_ind, aux_farm, aux_rec] = ValueFunction(dimWL, valueFunctionMaxIter, waterLevel, dimE_Rain, ...
-                                                dimE_Evap, utilFar, utilRec, beta, valueFunctionTolerance)
[…]

### 2020-02-28 Paper/paper.tex
use->utility

@@
 \end{equation}
+
 Both distributions are visualized in \ref{fig:rain-distributions}.
+
 \begin{figure}
@@
 \end{figure}
+
 \newpage
+
+%TODO in figure+code use durch utility ersetzen
 \begin{figure}[ht]
-	\includegraphics[width=1\textwidth]{figures/aggregated-use-rainfall-default.eps}
-	\caption{Aggregated Use for lognormal distributed rainfall $(\mu=0, \sigma^2=1)$}
-	\label{fig:aggregated-use-rainfall-default}
-\end{figure}
-\begin{figure}[ht]
-\includegraphics[width=1\textwidth]{figures/aggregated-use-rainfall-more-volatile.eps}
-\caption{Aggregated Use for lognormal distributed rainfall $(\mu=-1.5, \sigma^2=4)$}
-\label{fig:aggregated-use-rainfall-more-volatile}
+	\begin{subfigure}{\textwidth}
+
+	\includegraphics[width=1\textwidth]{figures/aggregated-utility-rainfall-default.eps}
+	\caption{Lognormal distributed Rainfalls with $\mu=0$ and $\sigma=1$}
+	\label{fig:aggregated-utility-rainfall-default}
+
+	\end{subfigure}
+	\newline
+	\begin{subfigure}{\textwidth}
+
+	\includegraphics[width=1\textwidth]{figures/aggregated-utility-rainfall-more-volatile.eps}
+	\caption{Lognormal distributed Rainfalls with $\mu=-1.5$ and $\sigma=2$}
+	\label{fig:aggregated-utility-rainfall-more-volatile}
+
+	\end{subfigure}
+	\caption{Effects of more volatile Rainfalls over Time}
+	\label{fig:aggregated-utilities-rainfall}
 \end{figure}
 
-The following figures demonstrate the effect of the changed volatility.
-They show the lognormal rainfalls with the original volatility (\ref{fig:aggregated-use-rainfall-default}), and with the increased volatility (\ref{fig:aggregated-use-rainfall-more-volatile}) over 100 periods.
-When comparing both visualizations, we see that more extreme rainfalls significantly affect the water level of the reservoir: there are longer periods with a rather low amount of water in the reservoir, but some periods where the reservoir reaches it's capacity limit.
+The graphs in \ref{fig:aggregated-utilities-rainfall} demonstrate the effect of the changed volatility.
+They compare the lognormal rainfalls with the original volatility, and with the increased volatility over 100 periods.
+When comparing both visualizations, we see that more extreme rainfalls significantly affect the water level of the reservoir: there longer periods with a rather low amount of water in the reservoir, but some periods where the reservoir reaches it's capacity limit.
 %When comparing both visualizations,it is noticeable that more volatile rainfall leads to much higher peaks of rain in combination with reduced number of periods with a decent amount of rain. 
